---
title: 'SUMMARY REPORT for new IQWiG data (restricted to 2 studies)'
author:
  - 'Saverio Fontana'
date: today
format:
  html:
      toc: true
      number-sections: true
      code-fold: true
      embed-resources: true
      smooth-scroll: true
      anchor-sections: true 
  pdf:
     toc: true
     number-sections: true
     colorlinks: true
execute:
  echo: !expr "knitr::is_html_output()"
  cache: true
  freeze: auto
  warning: false
  message: false
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false

library(confMeta)
library(bayesmeta)
library(dplyr)
library(tibble)
library(tidyr)
library(ggplot2)
library(patchwork)
library(gt)
library(parallel)
library(ggbeeswarm)
library(purrr)
library(ggridges)
library(ggforce)
library(plotly)
library(irr)
library(ggh4x)


```

```{r}
#| label: load-data
#| include: false

source("00_utilities.R") 

name <- "cis.rds"
inp <- file.path("Output", name)
input <- readRDS(inp)
cis <- input$cis
df_estimates <- input$df_estimates


# FIRST DATAFRAME
# Contains all stats for every Method x MA combination

df_mas <- map_dfr(cis, function(x) {
  x$ci %>%
    as.data.frame() %>%
    mutate(
      MA = x$ma_id,
      measure = x$measure,
      # Add heterogeneity (repeated for all methods in this MA)
      Q = x$heterogeneity$Q,
      I2 = x$heterogeneity$I2,
      Tau2 = x$heterogeneity$Tau^2,

      data_skewness = x$data_skewness
    )
}) %>%
  rename(p_val = p_0) 

# 2ND
# Contains 1 row per MA with study-level info (SEs, estimates, bounds)

df_inputs <- map_dfr(cis, function(x) {
   est <- x$inputs$estimate
   se  <- x$inputs$SE
   low <- x$inputs$lower
   upp <- x$inputs$upper
   
   data.frame(
     MA = x$ma_id,
     est_1 = est[1], 
     est_2 = est[2],
     
     SE_1 = se[1],   
     SE_2 = se[2],

     min_study_lower = min(low),
     max_study_upper = max(upp),

     SE_ratio = max(se) / min(se),
     opposing_directions = (est[1] * est[2]) < 0
   )
})


# This is useful for comparison w/ FE baseline

df_fe <- df_mas %>%
  filter(method == "Fixed effect") %>%
  select(MA, width_fe = width, p_fe = p_val, est_fe = estimate)

df_comparisons <- df_mas %>%
  filter(method != "Fixed effect") %>%
  left_join(df_fe, by = "MA") %>%
  mutate(
    rel_width_increase = (width - width_fe) / width_fe,
    p_diff = p_val - p_fe,
    shrinkage_ratio = ifelse(abs(est_fe) < 1e-6, NA, estimate / est_fe),
    sign_flip = sign(estimate) != sign(est_fe)
  )


```

# Data Overview

We are inspecting a total of **`r length(unique(df_estimates$identifier))`** meta-analysis. Those are all made up of 2 studies.

## Sheet division

Here is an overview of how many meta-analyisis are there in each Excel sheet

```{r}
#| label: tbl-Excel

a <- df_estimates %>%
   group_by(sheet_name) %>%
   summarize(n_MA = n()/2)

a %>% 
  gt() %>%
  #  Relabel Columns
  cols_label(
    sheet_name = md("**Data Sheet Name**"),
    n_MA = md("**Number of Meta-Analyses (N)**")
  )
```

## Measure employed

This is an overview of the different measures employed over the file

```{r}
#| label: tbl-measures


df_estimates %>%
  count(effect.measure, .drop = FALSE)  %>%   
  mutate(n = n/2 ) %>% 
   arrange(desc(n)) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    effect.measure = md("**How measures are reported**"),
    n = md("**Number**")
  )

```

## Data formatting

Another interesting fact was the way IQWIG reported the raw data. I classified them in 15 possible ways. This also show us which are the most common ways of reporting Meta Anslysis.

-   **2x2 OR**: Binary outcome reported as a 2×2 event table (a,b,c,d); effect computed as log odds ratio from counts.

-   **2x2 RR**: Binary outcome reported as a 2×2 event table; effect computed as log risk ratio from counts.

-   **2x2 RD**: Binary outcome reported as a 2×2 event table; effect computed as risk difference from counts.

-   **HR**: Study reports the log-Hazard Ratio with its standard error directly.

-   **direct OR**: Study reports the log-Odd Ratio with its standard error directly.

-   **direct RR**: Study reports the log-Risk Ratio with its standard error directly.

-   **direct RD**: Study reports the risk difference with its standard error directly.

-   **ROM**: Study reports the ratio of means on the log scale with its standard error directly.

-   **IDR**: Study reports the incidence density rate ratio on the log scale with its standard error directly.

-   **direct SMD**: Study reports the standardized mean difference with its standard error directly.

-   **direct MD**: Study reports the mean difference with its standard error directly.

-   **MD pooled SD**: Continuous outcome: study reports mean difference (MD), pooled SD, and group sizes (n1, n2).

-   **SMD from pooled**: Continuous outcome: study reports MD, pooled SD, and group sizes; SMD (Hedges’ g) is derived from these.

-   **singular studies MD**: Continuous outcome: study reports group means and SDs (ȳ1, s1, n1; ȳ2, s2, n2); MD is computed.

-   **singular studies SMD**: Continuous outcome: study reports SDs and group means; SMD is computed.

```{r}
#| label: tbl-formatting


df_estimates %>%
  count(data_report, .drop = FALSE)  %>%   
   arrange(desc(n)) %>% 
  mutate (n = n/2) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    data_report = md("**How data is reported**"),
    n = md("**Number**")
  )
```

## Methods Employed:

-   **Edgington**

-   **Weighted Edgington** with weights $\frac{1}{SE}$ (Inverse SE weighted)

-   **Weighted Edgington** with weights $\frac{1}{SE^2}$ (Inverse Variance weighted)

-   **Fixed Effect**. When tabular data are provided, the default is Mantel-Haenszel method, otherwise we fall back to Inverse-Variance method.

-   **Hartung-Knapp** (2001a/b) and variance correction is applied if HK confidence interval is narrower than CI from classic random effects model with DerSimonian-Laird estimator (IQWiG, 2022)

This summary is computed using all the MA from above.

**Scale** All ratio-based metrics (OR, RR, HR, etc.) are analyzed on the logarithmic scale. However, to remain consistent with IQWiG reporting conventions, they are labeled using their original metric names.

# Global Performances

## Width of the confidence intervals

For each method, we summarize the distribution of **CI widths** (upper − lower) across all two-study meta-analyses.

```{r}
#| label: tbl-widths

table_width <- compute_summary(df_mas, stats = "width")
render_gt(table_width, title = "Confidence Interval Widths")

```

### Width by effect measure

Widths stratified by effect measure

```{r}
#| label: tbl-widths-bymeasure


width_summary <- compute_summary(
  df_mas, 
  stats = "width", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)


render_gt(
  width_summary,
  title = "Confidence Interval Widths by Effect Measure and Method",
  group_col = "measure"
  )


```

**OBSERVATION**: Note the "staircase" effect in both median and mean: *Unweighted Edgington \> Edgington 1/SE \> Edgington 1/SE\^2 \> Fixed Effect* FOR EACH MEASURE.

This is the explanation I suggest:

For classic **Edgington** (Equal Weighting), each study contributes equally to the combined p-value function, regardless of its sample size or precision. Consequently, *the high variance of smaller, "noisy" studies is fully integrated into the pooled result.* This means we are forcing the precise studies to "compromise" with the imprecise ones, resulting in widest CI.

But **the more aggressively we penalize large standard errors** (moving from Unweighted $\rightarrow$ $1/SE$ $\rightarrow$ $1/SE^2$), the more you force the meta-analysis to *ignore uncertainty*. Basically we are giving more influence to certain data points and reducing the influence of uncertain ones.

### Width by effect measure (plot)

For a better visualization, the HK method is removed and ROM and RD are not plotted, since they only have respectively 1 and 6 observations.

**Note** that MD and SMD are reported on the square root scale

```{r}
#| label: fig-widths-bymeasure
#| fig-width: 20
#| fig-height: 15
#| fig-dpi: 300

widths_noHK <- df_mas %>% 
  filter (method != "Hartung & Knapp") %>% 
  filter (measure != "ROM" & measure != "RD")

ggplot(widths_noHK, aes(x = method, y = width, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~ measure, scales = "free_y") +
  labs(
    x = "Method",
    y = "CI Width"
  ) +
    facetted_pos_scales(
    y = measure %in% c("SMD", "MD") ~ scale_y_sqrt()
  ) + 
  theme_bw() +
  theme(
    plot.title   = element_text(face = "bold", size = 20),
    strip.text   = element_text(face = "bold", size = 16),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y  = element_text(size = 13),
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13),
    legend.position = "bottom",
    strip.background = element_rect(fill = "lightblue")
  ) +
  scale_fill_brewer(palette = "Set2")
```

::: {.callout-note icon="false"}
**My take on why HK gives wild CI**

The HK method produces extremely wide confidence intervals. This probably because both the variance estimator and the critical value from the t-distribution become unstable.

In the HK approach, the pooled effect estimate is the weighted mean

$$
\hat{\mu} = \frac{\sum_{i=1}^{k} w_i \hat{\theta}_i}{\sum_{i=1}^{k} w_i},
\qquad
w_i = \frac{1}{s_i^2 + \hat{\tau}^2},
$$

where $\hat{\theta}_i$ is the effect estimate from study $i$, $s_i^2$ its within-study variance, and $\hat{\tau}^2$ the estimated between-study variance.

The HK estimator of the variance of $\hat{\mu}$ is based on a weighted residual sum of squares,

$$
\widehat{\mathrm{Var}}(\hat{\mu})
=
\frac{1}{k-1}
\,
\frac{
\sum_{i=1}^{k} w_i^{2} \, (\hat{\theta}_i - \hat{\mu})^{2}
}{
\left(\sum_{i=1}^{k} w_i\right)^{2}
}.
$$

For the $(1-\alpha)$-level confidence interval, HK uses the $t$-distribution with $k-1$ degrees of freedom:

$$
\hat{\mu}
\pm
t_{k-1,\,1-\alpha/2}
\sqrt{\widehat{\mathrm{Var}}(\hat{\mu})}.
$$

With $k=2$ there are two problems.

The degrees of freedom are $k-1 = 1$, so the critical value $t_{1,1-\alpha/2}$ is very large (for example, $t_{1,0.975} \approx 12.7$), which directly inflates the length of the interval. Second, the variance estimator is essentially a scaled residual based on only one degree of freedom:

$$
\widehat{\mathrm{Var}}(\hat{\mu})
=
\frac{1}{1}
\,
\frac{
\sum_{i=1}^{2} w_i^{2} \, (\hat{\theta}_i - \hat{\mu})^{2}
}{
\left(\sum_{i=1}^{2} w_i\right)^{2}
},
$$
:::

### Relative Width Increase

$$
\frac{\text{Width}_{\text{Method}} - \text{Width}_{\text{FE}}}{\text{Width}_{\text{FE}}} 
$$ {#eq-relwidth}

Basically, we are measuring **how much the interval is increasing with respect to FE.**



```{r}
#| label: tbl-relative-width-summary

summary_tab <- df_comparisons %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    SD_Increase   = sd(rel_width_increase, na.rm = TRUE),
    Q25_Increase  = quantile(rel_width_increase, 0.25, na.rm = TRUE),
    Q75_Increase  = quantile(rel_width_increase, 0.75, na.rm = TRUE),
    Min_Increase  = min(rel_width_increase, na.rm = TRUE),
    Max_Increase  = max(rel_width_increase, na.rm = TRUE),
    .groups = "drop"
  )

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Relative Width Increase distribution**")
    ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean",
    Median_Increase = "Median",
    SD_Increase = "SD",
    Q25_Increase = "Q25",
    Q75_Increase = "Q75",
    Min_Increase = "Min",
    Max_Increase = "Max"
  ) %>%
  fmt_percent(
    columns = -method, 
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

### Relative Width Increase stratified by measure

```{r}
#| label: tbl-rel-width-stratified-summary

Rel_width_increase_summary <- compute_summary(
  df_comparisons, 
  stats = "rel_width_increase", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)

Rel_width_increase_summary %>%
  filter(method != "Fixed effect") %>%
  gt(groupname_col = "measure") %>%
  cols_label(
    method = "Method",
    mean = "Mean % Increase",
    median = "Median % Increase",
    sd = "SD % Increase",
    q25 = "25th % Increase",
    q75 = "75th % Increase",
    max = "Max % Increase",
    min = "Min % Increase"
  ) %>%
  fmt_percent(
    columns = c(mean, median, sd, q25, q75, max, min),
    decimals = 2
  ) %>% opt_stylize(style = 6, color = "blue") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```

## p-values (for $H_0: \theta=0$)

### p-values across methods

Distribution of method-specific p-values for testing $H_0: \theta=0$

*Note* the dashed line marks 0.05.


```{r}
#| label: tbl-pvals


table_p0 <- compute_summary(df_mas, stats = "p_val")
render_gt(table_p0, title = "P value distribution")

 
```


```{r}
#| label: fig-pvals



df_mas %>%
  ggplot(aes(x = method, y = p_val, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.8) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Method",
    y = "p-value",
    title = "Distribution of p-values by method"
  ) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  coord_cartesian(ylim = c(0, 1))

```



This aligns with previous observations: since FE tends to produce the narrowest intervals, it is the **more "aggressive" at rejecting the null.**

### ECDF of P-values

In this graph the **ECDF** of p-values is plotted. The grey dashed line is the CDF of the Uniform distribution.

```{r}
#| label: fig-pvalue-ecdf
#| fig-height: 7



df_mas %>%
  ggplot(aes(x = p_val, color = method)) +
  stat_ecdf(geom = "step", linewidth = 1, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", 
              color = "gray40", linewidth = 0.8) +
  annotate("text", x = 0.7, y = 0.3, 
           label = "Uniform distribution\n(under global null)", 
           color = "gray40", size = 3.5) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "P-value",
    y = "Cumulative Probability",
    color = "Method"
    ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
  
```

### Distribution of difference of p values across methods


Now, we will compare each method's p-values with **Fixed Effect's p-values as baseline.**

This table is similar to @tbl-pvals but with a focus on the difference $\Delta p = p_0 (method) - p_0 (FE)$ and showing the correlation between $p_0(method)$ and $p_0 (FE)$ .

```{r}
#| label: tbl-pvals-differences

df_comparisons %>%
  group_by(method) %>%
  summarise(
    mean_diff = mean(p_diff, na.rm = TRUE),
    median_diff = median(p_diff, na.rm = TRUE),
    sd_diff = sd(p_diff, na.rm = TRUE),
    pearson = cor(p_val, p_fe, method = "pearson", use = "complete.obs"),
    spearman = cor(p_val, p_fe, method = "spearman", use = "complete.obs"),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**P-value Differences from Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    mean_diff = "Mean",
    median_diff = "Median Δ",
    sd_diff = "SD Δ",
    pearson = "Pearson",
    spearman = "Spearman"
  ) %>%
  tab_spanner(
    label = "Correlation with FE",
    columns = c(pearson, spearman)
  ) %>%
  fmt_number(
    columns = c(mean_diff, median_diff, sd_diff, pearson, spearman),
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  opt_stylize(style = 6, color = "blue")
```
Quite interestingly, the **median difference equall zero.** 

However, the slightly higher **mean difference (0.02)** suggests that in a subset of cases, likely those w/ high study-level conflict, the Edgington method provides a more conservative p-value than the Fixed Effect model.

### Distribution of difference of p values across methods (plot)

Here, each panel plots the FE p-values vs the p-value of the specific method.

*Note*: if the $p_0$ were equal, points would distribute on the diagonal.

**Each point is one meta-analysis.** Points below the diagonal indicate smaller p-values than fixed effect (less conservative), and points above indicate larger p-values (more conservative).
```{r}
#| label: fig-pvalue-vs-fe
#| fig-height: 8

df_comparisons %>% 
  ggplot (aes(x = p_fe, y = p_val, color = method)) +
  geom_point(alpha= 0.4, size = 1.5) +
  geom_abline (intercept = 0, slope = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) + 
  facet_wrap (~method) + 
  scale_x_continuous (breaks = seq(0,1, 0.2)) + 
  scale_y_continuous(breaks = seq(0,1, 0.2)) + 
  labs(
    x = "P-value (Fixed Effect)",
    y = "P-value (Method)",
    title = "Method p-values vs Fixed Effect p-values"
  ) +
  theme(
    legend.position = "none"
  )
```

### Agreement in Significance (p \< 0.05)

Here we inspect the significance agreement between the values at a $\alpha = 0.05$ level

Each meta-analysis is classified into **four categories:** both significant, both non-significant, FE-only, or method-only.

```{r}
#| label: fig-significance-agreement
#| fig-height: 6



sign_agree_p <- df_comparisons %>% 
  group_by(method) %>% 
  mutate (
    Method_sign = p_val <0.05,
    FE_sign = p_fe <0.05, 
    agreement = ifelse(FE_sign & Method_sign, "Both significant",
                       ifelse(!FE_sign & !Method_sign, "Both non-significant",
                              ifelse(FE_sign & !Method_sign, "FE sig, Method non-sig", 
                                     "FE non-sig, Method sig"))))
  
  
  
  
  

sign_agree_p %>% 
  count(method, agreement) %>% 
  ggplot(aes(x = method, y = n, fill = agreement)) + 
           geom_col (position= "stack", alpha = 0.8) + 
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3,
            color = "white", fontface = "bold") + 
  scale_fill_manual ( 
    values = c(      "Both significant" = "#2ca02c",
                     "Both non-significant" = "#d62728",
                     "FE sig, Method non-sig" = "purple",
                     "FE non-sig, Method sig" = "#7f7f7f"
    )) + 
  labs (
    x = "Methods", 
    y = "Count",
    fill = "Agreement", 
    title = "Significance agreement with Fixed Effect (α = 0.05)"
  ) + 
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
  

         
```

```{r}
#| label: tbl-significance-agreement

sign_agree_p %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    both_sig = sum(FE_sign & Method_sign, na.rm = TRUE),
    both_nonsig = sum(!FE_sign & !Method_sign, na.rm = TRUE),
    fe_only = sum(FE_sign & !Method_sign, na.rm = TRUE),
    method_only = sum(!FE_sign & Method_sign, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Significance Agreement with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "Total",
    both_sig = "Both Sig",
    both_nonsig = "Both Non-sig",
    fe_only = "FE Only",
    method_only = "Method Only"
  ) %>%
  fmt_number(
    columns = c(n_total, both_sig, both_nonsig, fe_only, method_only),
    decimals = 0
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  )  %>%
  opt_stylize(style = 6, color = "blue")
```

Thus, this shows that classic Edgington is the most independent by FE, disagreeing with it in 89 cases ($7.2\%$), whereas the weighting progressively reduces this disagreement. Notably, 25 times Edgington was significant while FE not, **this strengthen the idea that it is not only a more conservative test.**


## Point Estimates (Shrinkage Ratio)

Inspecting the Estimates (the point which has the highest p-value) provided by each method

```{r}
#| label: tbl-estimates


table_estimates <- compute_summary(df_mas, stats = "estimate")
render_gt (table_estimates, title = "Estimates")
```

### Agreement of sign of Estimates

Comparison of the **direction** of each method’s estimate with the **Fixed Effect** one.

For each method, we summarisethe proportion of meta-analyses where the method has the same sign as FE and the proportion with a flipped sign relative to FE, **among all meta-analyses**

```{r}
#| label: tbl-agreement-estimates



summ_long_vs_FE <- df_comparisons %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    # "sign_flip" is TRUE if signs differ, FALSE if they are the same
    prop_same_sign = mean(!sign_flip, na.rm = TRUE) * 100,
    prop_flip_sign = mean(sign_flip, na.rm = TRUE) * 100,
    .groups = "drop"
  )

summ_long_vs_FE %>%
  gt() %>%
  tab_header(
    title = md("**Agreement of sign with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "N studies",
    prop_same_sign = "Same sign as FE (%)",
    prop_flip_sign = "Sign flip vs FE (%)",
  ) %>%
  fmt_number(
    columns = c(prop_same_sign, prop_flip_sign),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(n_total),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")

```

### Shrinkage ratio

$$SR = \frac {\hat{\mu}_{\text{method}}}  {\hat{\mu}_{\text{FE}}}$$

Negative values of SR indicate a **sign flip** relative to FE, while positive values indicate agreement in direction. Among positive ratios, values between 0 and 1 correspond to **shrinkage toward zero** relative to FE, and values larger than 1 indicate **more extreme estimates** than FE. Finally, value equal to 1 (using the tolerance) indicate that the two methods provide the same estimate.

We **exclude** meta analysis where $|\hat{\mu}_{\text{FE}} | < tol$ (i.e., where the Fixed Effect estimate is effectively zero), as the ratio is undefined in these cases.

For each method, we report how often estimates flip sign, how often they are shrunken toward zero, how often they are more extreme than FE.

```{r}
#| label: tbl-SR-summary1
#   SHRINKAGE RATIO SR = method / FE  


ratio_summary <- df_comparisons %>%
  # Filter out undefined ratios (where FE was close to 0)
  filter(!is.na(shrinkage_ratio)) %>%
  mutate(
    SR_class = case_when(
      abs(shrinkage_ratio - 1) < 1e-6 ~ "Equal (SR=1)",
      shrinkage_ratio < 0 ~ "Flip sign (SR<0)",
      shrinkage_ratio >= 0 & shrinkage_ratio < 1 ~ "Shrink toward 0",
      shrinkage_ratio > 1 ~ "More extreme (SR>1)"
    )
  ) %>%
  group_by(method) %>%
  summarise(
    n = n(),
    prop_flip = mean(SR_class == "Flip sign (SR<0)") * 100,
    prop_shrink = mean(SR_class == "Shrink toward 0") * 100,
    prop_more_extreme = mean(SR_class == "More extreme (SR>1)") * 100,
    prop_exactly_same = mean(SR_class == "Equal (SR=1)") * 100
  )

ratio_summary %>% 
  gt() %>%
  tab_header(
    title = md("**Shrinkage ratio summary**"),
  ) %>%
  cols_label(
    method = "Method",
    n = "N MA",
    prop_flip = "Flipped signs (%)",
    prop_shrink = "Shrinked estimates(%)",
    prop_more_extreme = "More extreme estimates (%)",
    prop_exactly_same = "Same estimates (%)"
    ) %>%
  fmt_number(
    columns = c(prop_flip, prop_shrink, prop_more_extreme, prop_exactly_same),
    decimals = 2
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 

```

**NOTE**: Ratios were undefined for `r sum(is.na(df_comparisons$shrinkage_ratio)) / length(unique(df_comparisons$method))` MA, where the Fixed Effect estimate was effectively zero ($|\hat{\mu}_{FE}| < 10^{-6}$); these cases were excluded from the shrinkage summary."

```{r}
#| label: tbl-SR-summary2


sr_global <- df_comparisons %>%
  filter(!is.na(shrinkage_ratio)) %>%
  group_by(method) %>%
  summarise(
    Studies = n(),
    Mean   = mean(shrinkage_ratio),
    SD     = sd(shrinkage_ratio),
    Median = median(shrinkage_ratio),
    Q25    = quantile(shrinkage_ratio, 0.25),
    Q75    = quantile(shrinkage_ratio, 0.75),
    Min    = min(shrinkage_ratio),
    Max    = max(shrinkage_ratio)
  )

sr_global %>%
  gt() %>%
  tab_header(
    title = md("**Distribution of Shrinkage Ratios (Method / FE)**"),
    subtitle = "Values < 1 indicate shrinkage toward zero. Values < 0 indicate sign flip."
  ) %>%
  cols_label(method = "Method") %>%
  fmt_number(
    columns = c(Mean, SD, Median, Q25, Q75, Min, Max),
    decimals = 2
  ) %>%
  cols_align(align = "center", columns = -method) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

Median equal 1 is promising

### Stratified Shrinkage Ratio

```{r}
#| label: tbl-SR-stratified


sr_stratified <- compute_summary(
  df_comparisons, 
  stats = "shrinkage_ratio", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)

render_gt(
  sr_stratified,
  title = "Shrinkage Ratio Distribution by Effect Measure",
  group_col = "measure", decimals = 2
  )
```

# Conflict in Precision

Define $$SE_{ratio} = \frac{max(SE)}{min(SE)}$$

In this section we want to inspect how different methods behaves as the $SE_{ratio}$ of the **singular studies** increases.

## Width vs SE_ratio

**Does the method widen the interval when one study is much less precise than the other?**


```{r}
#| label: tbl-widths-vs-se

# Join results with inputs to get SE_ratio
merged_res_inp <- df_mas %>%
  left_join(df_inputs, by = "MA")


corr_df <- merged_res_inp %>%
  group_by(method) %>%
  summarise(
    Pearson = cor(width, SE_ratio, use = "complete.obs", method = "pearson"),
    Spearman = cor(width, SE_ratio, use = "complete.obs", method = "spearman")
  )

# Reshape the table
corr_vec_df <- corr_df %>%
  pivot_longer(
    cols = c(Pearson, Spearman), 
    names_to = "Metric", 
    values_to = "Correlation"
  ) %>%
  pivot_wider(
    names_from = method, 
    values_from = Correlation
  )

corr_vec_df %>%
  gt() %>%
  fmt_number(
    columns = -Metric,
    decimals = 2
  ) %>%
  tab_header(
    title = md("**Correlation between SE ratio and Width**"),
  ) %>% 
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_stylize(style = 6, color = "red")

```

### What about the Relative Width increase? 

```{r}
merged_comp_inp <- df_comparisons %>% 
  left_join(df_inputs, by = "MA")

corr_df <- merged_comp_inp %>%
  group_by(method) %>%
  summarise(
    Pearson = cor(rel_width_increase, SE_ratio, use = "complete.obs", method = "pearson"),
    Spearman = cor(rel_width_increase, SE_ratio, use = "complete.obs", method = "spearman")
  )

# Reshape the table
corr_vec_df <- corr_df %>%
  pivot_longer(
    cols = c(Pearson, Spearman), 
    names_to = "Metric", 
    values_to = "Correlation"
  ) %>%
  pivot_wider(
    names_from = method, 
    values_from = Correlation
  )

corr_vec_df %>%
  gt() %>%
  fmt_number(
    columns = -Metric,
    decimals = 2
  ) %>%
  tab_header(
    title = md("**Correlation between SE ratio and Relative Width Increase**"),
  ) %>% 
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_stylize(style = 6, color = "red")
```


This mean that as $SE_{ratio}$ increases, **Edgington will produces wider intervals**. This because it is "forced" to incorporate the high uncertainty (large SE) of the smaller study into the pooled p-value function.

At the same time, as $SE_{ratio}$ grows, the weigthed methods center their estimate and precision more and more exclusively on the "certain" study, reducing the contribution of the high-variance study


## p-value difference vs SE ratio 
Recall that $\Delta p = p_0 (method) - p_0 (FE)$ 


```{r}
#| label: tbl-pvals-vs-se-ratio


corr_p_se <- merged_comp_inp %>%
  group_by(method) %>%
  summarise(
    Pearson = cor(p_diff, SE_ratio, use = "complete.obs", method = "pearson"),
    Spearman = cor(p_diff, SE_ratio, use = "complete.obs", method = "spearman"),
    .groups = "drop"
  )

# Generate table
corr_p_se %>%
  pivot_longer(cols = c(Pearson, Spearman), names_to = "Metric", values_to = "Correlation") %>%
  pivot_wider(names_from = method, values_from = Correlation) %>%
  gt() %>%
  fmt_number(columns = -Metric, decimals = 2) %>%
  tab_header(
    title = md("**Correlation between SE Ratio and P-value Difference (Δp)**")
  ) %>% 
  cols_align(align = "center", columns = everything()) %>%
  opt_stylize(style = 6, color = "red")
```

All the values are extremely close to zero: this seems to be good news! The fact that the studies have very different variability, doesn't influence our conclusions on significance.


# Conflict in Direction
In this section, we stratify between MAs that show **opposite signs of the effect estimates** (e.g. study 1 has positive estimate, while study 2 has negative estimate, or the opposite) and that show the same sign.

## Sign flips vs Relative Width Increase

The idea is that if the method is sensitive to directional conflict, the relative width increase should be significantly higher for "Opposite Sign" cases.

```{r}
#| label: tbl-width-differences-inspection


conflict_width_data <- df_comparisons %>%
  filter(method %in% c("Edgington", "Edgington (1/SE)", "Edgington (1/SE^2)")) %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    Alignment = ifelse(opposing_directions, "Opposite Sign", "Same Sign")
  )


width_conflict_summary <- conflict_width_data %>%
  group_by(method, Alignment) %>%
  summarise(
    n = n(),
    mean_rel_increase = mean(rel_width_increase),
    median_rel_increase = median(rel_width_increase),
    .groups = "drop"
  )

# format table nice
tab_width_dat <- width_conflict_summary %>%
  select(method, Alignment, n, mean_rel_increase, median_rel_increase) %>%
  group_by(method) %>%
  mutate(
    method = ifelse(row_number() == 1, as.character(method), "")
  ) %>%
  ungroup()

tab_width_dat %>%
  gt() %>%
  cols_label(
    method = "Method",
    Alignment = "Sign",
    n = "N",
    mean_rel_increase = "Mean Rel. Increase",
    median_rel_increase = "Median Rel. Increase"
  ) %>%
  fmt_percent(
    columns = c(mean_rel_increase, median_rel_increase),
    decimals = 2
  ) %>%
    tab_header(
    title = md("**Sign vs Relative Width Increase**")
  ) %>% 
  cols_align(
    align = "center",
    columns = -c(method, Alignment)
  ) %>%
  opt_stylize(style = 6, color = "red") 
```


```{r}
#| label: fig-rel-width-conflict


ggplot(conflict_width_data, aes(x = method, y = rel_width_increase, fill = Alignment)) +
  geom_boxplot(alpha = 0.8, outlier.size = 1, outlier.alpha = 0.4) +
  scale_fill_manual(values = c("Same Sign" = "#34495e", "Opposite Sign" = "red")) +
  scale_y_continuous(trans = "log1p", labels = scales::percent) +
  theme_bw() +
  labs(
    y = "Relative Width Increase vs. Fixed Effect (log1p scale)",
    fill = "Study Alignment",
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```



## Sign flips p value difference


```{r}
#| label: tbl-pvals-differences-inspection


conflict_analysis <- df_comparisons %>%
  filter(method != "Hartung & Knapp") %>%
  left_join(df_inputs, by = "MA")

conflict_summary <- conflict_analysis %>%
  group_by(method, opposing_directions) %>%
  summarise(
    n = n(),
    mean_delta_p = mean(p_diff),
    median_delta_p = median(p_diff)
  )


tab_dat <- conflict_summary %>%
  mutate(
    Sign = ifelse(opposing_directions, "Opposite", "Same")
  ) %>%
  select(method, Sign, n, mean_delta_p, median_delta_p) %>%
  group_by(method) %>%
  mutate(
    method = ifelse(row_number() == 1, method, "")
  ) %>%
  ungroup()

tab_dat %>%
  gt() %>%
  cols_label(
    method = "Method",
    Sign = "Sign",
    n = "N",
    mean_delta_p = "Mean Δ",
    median_delta_p = "Median Δ"
  ) %>%
  fmt_number(
    columns = c(mean_delta_p, median_delta_p),
    decimals = 2
  ) %>%
      tab_header(
    title = md("**Sign vs p vals difference**")
  ) %>% 
  cols_align(
    align = "center",
    columns = -c(method, Sign)
  ) %>%
  opt_stylize(style = 6, color = "red")
```

```{r}
#| label: fig-directions



plot_data <- df_comparisons %>%
  filter(method == "Edgington") %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    conflict = ifelse(opposing_directions, "Directional Conflict", "Consistent Direction")
  )


ggplot(plot_data, aes(x = p_fe, y = p_val, color = conflict)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey30") + # Identity line
  scale_color_manual(values = c("Consistent Direction" = "#34495e", 
                               "Directional Conflict" = "red")) +
  theme_bw() +
  labs(
    x = "Fixed Effect P-value",
    y = "Edgington P-value",
    color = "Sign Alignment",
    title = " P values for different sign estimates"
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold")
  )
```

Here just for visual understanding I display the 10 most conflicted MAs.


```{r}
#| label: tbl-pvals-extreme-cases

extreme_conflicts <- conflict_analysis %>%
  filter(opposing_directions == TRUE) %>%
  arrange(desc(p_diff)) %>%
  slice_head(n = 10)

extreme_conflicts %>%
  select(MA, measure, p_val, p_fe, p_diff, est_1, est_2, SE_ratio) %>%
  gt() %>%
  cols_label(
    MA = "MA",
    measure = "Measure",
    p_val = "p-val Edgington",
    p_fe = "p-val FE",
    p_diff = "p-val difference",
    est_1 = "Estimate study 1",
    est_2 = "Estimate study 2",
    SE_ratio = "SE ratio"
  ) %>%
  fmt_number(
    columns = c(p_val, p_fe, p_diff,est_1, est_2, SE_ratio),
    decimals = 2
  )
```

# Heterogeneity
**Cochran's** $Q$ $$Q = \sum_{i=1}^{k} w_i (\hat{\theta}_i - \hat{\theta}_{FE})^2$$where $k$ is the number of studies (here $k=2$), and $w_i = 1/\sigma_i^2$. Under the null hypothesis of homogeneity, $Q$ follows a $\chi^2$ distribution with $k-1$ degrees of freedom.

**Inconsistency Index** ($I^2$) Describes the percentage of variation across studies that is due to heterogeneity rather than chance.$$I^2 = \max\left(0, \frac{Q - (k - 1)}{Q}\right) \times 100\%$$

**Between-study Variance** ($\tau^2$) estimated using REML

**Note** 

Under homogeneity ($\tau^2 = 0$), we have $\theta_i \sim N(\mu, v_i)$. Then $\theta_1 - \theta_2 \sim N(0, v_1 + v_2)$, so $Q = \frac{(\theta_1 - \theta_2)^2}{v_1 + v_2} \sim \chi^2_1$.
And under $H_0$ we have ${Q \sim \chi^2_1}$. 

This is a Z test... i.e. $Q = Z^2$


Critical value at $\alpha = 0.05$: $3.84$.
So you reject iff $|\theta_1 - \theta_2| > \sqrt{3.84 (v_1 + v_2)}$.

```{r}
summary_hetero <- df_mas %>% 
  summarise(
    total = n()/length(unique(df_mas$method)),
    avg_Q = mean(Q),
    avg_I2 = mean(I2),
    avg_tau = mean(Tau2),
    .groups = "drop"
  )

summary_hetero %>%
  gt() %>%
  cols_label(
    total = "Total MA",
    avg_Q = "Average Q value",
    avg_I2 = "Average I2 value",
    avg_tau = "Average tau value"
  ) %>%
  fmt_number(
    columns = c(avg_Q,avg_I2, avg_tau),
    decimals = 2
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "red")
```

## Heterogeneity summary

Full graph on the above, with a zoomed-in view below due to the outliers that ruin the visualisation. Data are reported on the square root scale.

```{r}
#| label: fig-heterogeneity-stratified
#| fig-height: 6

# We take unique MA rows to avoid counting Q multiple times
q_unique <- df_mas %>% 
  distinct(MA, .keep_all = TRUE)


#use the ggforce package --> facet_zoom allow us to zoom in


ggplot(q_unique, aes(x = Q, y = measure, color = measure)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.5) + 
  geom_jitter(height = 0.1, alpha = 0.5, size = 1) +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 500), 
    name = "Cochran's Q (Scale sqrt)"
  ) +
  
  # Note: xlim in facet_zoom must use original Q values 
  facet_zoom(xlim = c(0, 20), zoom.size = 1) + 
  
  theme_bw() +
  theme(
    legend.position = "none" 
  )

```

**NOTE**, there is a huge outlier, with id "*A20-93 \|\| Abb. 12 \|\| HbA1c*". It has a Q-statistics = 570.95.


## Correlation with Cochran's Q

```{r}
#| label: tbl-Q-vs-width-pvals

df_mas %>%
  group_by(method) %>%
  summarise(
    Cor_Q_Width = cor(Q, width, method = "spearman", use = "complete.obs"),
    Cor_Q_Width_p  = cor(Q, width, method = "pearson", use = "complete.obs"),
    Cor_Q_Pval  = cor(Q, p_val, method = "spearman", use = "complete.obs"),
    Cor_Q_Pval_p  = cor(Q, p_val, method = "pearson", use = "complete.obs"),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Correlation with Heterogeneity**"),
  ) %>%
  cols_label(
    method = "Method",
    Cor_Q_Width = "Corr(Q, Width) Spearman",
    Cor_Q_Width_p = "Corr(Q, Width) Pearson", 
    Cor_Q_Pval = "Corr(Q, P-val) Spearman", 
    Cor_Q_Pval_p  = "Corr(Q, P-val) Pearson"
  ) %>%
  fmt_number(columns = starts_with("Cor"), decimals = 2) %>%
  data_color(
    columns = starts_with("Cor"),
    method = "numeric",
    palette = "RdBu", # Red-Blue diverging palette
    domain = c(-1, 1)
  ) %>%
  opt_stylize(style = 6, color = "red")
```

The positive correlation shows that **Edgington "adapts":** When $Q$ increases (meaning studies are moving apart), the Edgington confidence interval widens to account for that conflict.

The large gap between Pearson and Spearman suggests that the relationship between heterogeneity and Edgington width is *not linear,* but it is monotonic. (e.g. Edgington doesn't widen in a straight line as $Q$ grows; but maybe stays stable for a while and then "reacts" more strongly once Q reaches a certain threshold)


The correlations between $Q$ and P-values for Edgington methods are essentially zero. This suggests that while Edgington widens its uncertainty (Width) when studies disagree, it doesn't necessarily lose its ability to find an effect (P-value) just because the studies are heterogeneous.
In contrast,  FE: has a negative Spearman correlation. This suggests that as studies move apart ($Q$ increases), the p-value actually tends to get "smaller" (more significant). 

Thus, in this context, Edgington appears more stable. 


## Relative width increase vs Q

We plot Relative width increase [@eq-relwidth] against Cochran's $Q$ to show how methods respond to heterogeneity. A positive value indicates a penalty for heterogeneity: as the study estimates diverge, the method produces a wider (more conservative) interval.

On the right, all methods are plotted together, on the right, to better inspection, the HK method is removed.

**This plot uses the LOESS smoother**. This can be misleading, it is just to detect a general trend, next plots will inspect deeper.

**IMPORTANT** I often use *log1p scale*, i.e. $log1p(x)=log(1+x)$

```{r}
#| label: fig-q-vs-rel-width
#| fig-height: 15
#| fig-width: 20

#add the panels
data_all <- df_comparisons %>% mutate(panel = "All Methods")
data_no_hk <- df_comparisons %>% 
  filter(method != "Hartung & Knapp") %>% 
  mutate(panel = "Excluding HK")

combined_data <- bind_rows(data_all, data_no_hk)


ggplot(combined_data, aes(x = Q, y = rel_width_increase, color = method)) +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) + 
  
  facet_wrap(~panel, scales = "free_y") +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 400),
    name = "Cochran's Q (sqrt scale)"
  )  + 
  scale_y_continuous(trans = "log1p", labels = scales::percent, 
      name = "Rel. Width Increase (log1p Scale)")+
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 30),
    axis.title.x = element_text(size = 25), 
    axis.title.y = element_text(size = 25), 
    axis.text.x  = element_text(size = 20), 
    axis.text.y  = element_text(size = 20),      
    legend.position = "bottom",
    legend.title = element_text(size = 25), 
    legend.text  = element_text(size = 25)  
  )
```


```{r}
#| label: tbl-relative-width-summary-Q


summary_tab <- df_comparisons%>%
 filter (method != "Fixed effect") %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean_Increase))

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Performance vs Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Increase)[Spearman]", 
    cor_pear = "Corr(Q, Increase)[Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

Edgington with a Mean % Increase of 39.7% and a Median of 11.0%, is highly sensitive to study-level conflict.
On the other hand, the weighted with $1/SE^2$ shows a near perfect Spearman correlation, but the magnitude of the actual penalty is smaller  (Mean = 15.0% Median = 0.0% )


**Note**:
Cochran’s $Q$ is fundamentally an inverse-variance weighted statistic ($w_i = 1/SE^2$). Because weighted Edgington uses this exact same weighting scheme, they are strongly correlated. 



## Stratified Heterogeneity Analysis

```{r}
#| label: fig-rel_width_stratified
#| fig-width: 16
#| fig-height: 20
#| fig-dpi: 300

#remove ROM, RD and IDR 
data_clean <- data_no_hk %>% 
  filter(measure != "ROM" & measure != "RD" & measure != "IDR")

ggplot(data_clean, aes(x = Q, y = rel_width_increase, color = method)) +
    #geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) +
    geom_point() +
    facet_wrap(~ measure, scales = "free") +
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)")+
    theme_bw() +
    theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 23),
    axis.title.x = element_text(size = 17), 
    axis.title.y = element_text(size = 17), 
    axis.text.x  = element_text(size = 12), 
    axis.text.y  = element_text(size = 12),      
    legend.position = "bottom",
    legend.title = element_text(size = 17), 
    legend.text  = element_text(size = 17) 
    )
```

```{r}
#| label: tbl-rel_width_stratified

summary_tab_strat <- df_comparisons %>%
  filter(method != "Fixed effect") %>%
  group_by(measure) %>%
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase),
    Median_Increase = median(rel_width_increase),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  mutate(measure = paste0(measure, " (N = ", n_ma, ")"))  %>%
  arrange(measure, desc(Mean_Increase))

summary_tab_strat %>%
  select (-n_ma) %>% 
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Sensitivity to Heterogeneity by Measure**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Incr) [Spearman]", 
    cor_pear = "Corr(Q, Incr) [Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "red") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```


### Plot relative width increase stratified by measure and method

Note that observation "*A20-93 \|\| Abb. 12 \|\| HbA1c*" is removed

```{r}
#| label: fig-rel-width-stratified2
#| fig-width: 16
#| fig-height: 20
#| fig-dpi: 300

data_plot <- data_clean %>% 
  filter(method != "Fixed effect") %>% 
  filter(MA != "A20-93 || Abb. 12 || HbA1c")


ggplot(data_plot, aes(x = Q, y = rel_width_increase, color = method)) +
    geom_jitter(alpha = 0.4, size = 1.5, width = 0.1) + 
    # geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) +

    facet_grid(method ~ measure, scales = "free") + 
    
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)") +
    
   #  line at 0%
      geom_hline(
    yintercept = 0,
    color = "blue",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  
    theme_bw(base_size = 25) +
    theme(
      legend.position = "none",
      strip.text = element_text(size = 10, face = "bold"),
      panel.spacing = unit(0.7, "lines"),
      panel.grid.minor = element_blank()
    )
```


For singular inspection, here is an interctive plot:

```{r}
#| label: fig-rel-width-stratified2-plotly


p <- ggplot(data_plot, aes(
    x = Q, 
    y = rel_width_increase, 
    color = method,
    # w 'text' define what shows up in the box
    text = paste("MA ID:", MA, 
                 "<br>Cochran's Q:", round(Q, 2), 
                 "<br>Increase:", scales::percent(rel_width_increase))
  )) +
  geom_jitter(alpha = 0.5, size = 1.2, width = 0.1) + 
  facet_wrap(~ method + measure, scales = "free") + 
  scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)") +
  theme_bw() +
  theme(legend.position = "none")


# 'tooltip = "text"' tells plotly to use the text we defined above
ggplotly(p, tooltip = "text", height = 1200, width = 1000)
```



# Other validations

## Confidence interval skewness

The skewness of each confidence interval is computed as:

Of course Fixed Effect mehods cannot be skewed so we don't report them. $$
\text{Skewness} = \frac{\text{upper} + \text{lower} - 2 \times \text{estimate}}{\text{upper} - \text{lower}}$$

```{r}
#| label: tbl-skewness


ci_skewness <- df_mas %>% filter (!(method %in% c("Fixed effect", "Hartung & Knapp")))
table_skewness <- compute_summary (ci_skewness, stats = "ci_skewness")


render_gt (table_skewness, title = "Confidence Interval skewness")

```




## Coherence of skewness signs (β, γ, AUCC ratio)

Agreement between different measures of skewness:

-   **β (beta)**: CI skewness
-   **γ (gamma)**: Data skewness
-   **AUCC ratio**: Area Under Confidence Curve ratio
-   **Cohen’s** $\kappa$: measure of agreement between signs

where $$\gamma = \frac{\sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^3}{\left( \sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^2 \right)^{3/2} / \sqrt{\sum_{i=1}^{n} w_i}}$$ is Fisher’s weighted skewness coefficient.

And $$
\text{AUCC Ratio} = \frac {\text{AUCC}_{\text{upper}} - \text{AUCC}_{\text{lower}}} {AUCC}
$$.

*For more details see: Held, Leonhard, Felix Hofmann, and Samuel Pawel. "A comparison of combined p-value functions for meta-analysis." Research Synthesis Methods (2025)*

```{r}
#| label: sign-agreement-preparation

# Build sign agreement dataset
sign_agreement <- df_mas %>%
  filter(
    !is.na(ci_skewness),
    !method %in% c("Fixed effect", "Hartung & Knapp", "Bayesmeta")
  ) %>%
  mutate(
    same_sign_beta_gamma = sign(data_skewness) == sign(ci_skewness),
    same_sign_beta_aucc  = sign(ci_skewness) == sign(aucc_ratio),
    same_sign_gamma_aucc = sign(data_skewness) == sign(aucc_ratio)
  )

```

```{r}
#| label: tbl-sign-agreement



# Calculate summary statistics
sign_summary <- sign_agreement %>%
  group_by(method) %>%
  summarise(
    n_MA = n(),
    agree_beta_gamma = sum(same_sign_beta_gamma, na.rm = TRUE),
    agree_beta_aucc = sum(same_sign_beta_aucc, na.rm = TRUE),
    agree_gamma_aucc = sum(same_sign_gamma_aucc, na.rm = TRUE),
    corr_beta_gamma = cor(ci_skewness, data_skewness, use = "complete.obs"),
    corr_beta_aucc = cor(ci_skewness, aucc_ratio, use = "complete.obs"),
    corr_gamma_aucc = cor(data_skewness, aucc_ratio, use = "complete.obs"),
     kappa_beta_gamma = {
      # We extract the signs and convert in factors
      s1 <- factor(sign(data_skewness), levels = c(-1, 0, 1))
      s2 <- factor(sign(ci_skewness), levels = c(-1, 0, 1))
      # kappa2 calculates the agreement:
      kappa2(data.frame(s1, s2))$value
    },
    .groups = "drop"
  )

# Create gt table
sign_summary %>%
  gt() %>%
  tab_header(
    title = md("**Sign Agreement**"),
  ) %>%
  tab_spanner(
    label = md("**Sign Agreement (Count)**"),
    columns = c(agree_beta_gamma, agree_beta_aucc, agree_gamma_aucc)
  ) %>%
  tab_spanner(
    label = md("**Pearson Correlation**"),
    columns = c(corr_beta_gamma, corr_beta_aucc, corr_gamma_aucc)
  ) %>%
  cols_label(
    method = "Method",
    n_MA = md("*MA*"),
    agree_beta_gamma = "β vs γ",
    agree_beta_aucc = "β vs AUCC",
    agree_gamma_aucc = "γ vs AUCC",
    corr_beta_gamma = "β vs γ",
    corr_beta_aucc = "β vs AUCC",
    corr_gamma_aucc = "γ vs AUCC",
    kappa_beta_gamma = "Cohen's κ (β vs γ)"

  ) %>%
  fmt_number(
    columns = c(starts_with("corr"), kappa_beta_gamma),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(n_MA, starts_with("agree")),
    decimals = 0
  ) %>%
  sub_missing(
    columns = everything(),
    missing_text = "–"
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

## Sign agreement among β, γ, and AUCC ratio

```{r}
#| label: fig-sign-agreement
#| fig-height: 6

sign_long <- sign_agreement %>%
  filter(!method %in% c("Fixed effect", "Hartung & Knapp")) %>%
  select(MA, method, starts_with("same_sign")) %>%
  pivot_longer(
    cols = starts_with("same_sign"),
    names_to = "comparison",
    values_to = "agreement"
  ) %>%
  filter(!is.na(agreement))
  
# Create bar plot showing proportions
sign_long %>%
  mutate(
    comparison = case_when(
      comparison == "same_sign_beta_gamma" ~ "β vs γ",
      comparison == "same_sign_beta_aucc" ~ "β vs AUCC",
      comparison == "same_sign_gamma_aucc" ~ "γ vs AUCC"
    )
  ) %>%
  group_by(method, comparison) %>%
  summarise(
    prop_agree = mean(agreement, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = method, y = prop_agree, fill = comparison)) +
  geom_col(position = position_dodge(0.8), width = 0.7, alpha = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey30") +
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.1)
  ) +
  scale_fill_manual(
    values = c("#1f77b4", "#ff7f0e", "#2ca02c")
  ) +
  labs(
    x = "Method",
    y = "Proportion of Agreement",
    fill = "Comparison"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 12),
    legend.position = "bottom"
  )
```

## Is the study estimate contained in the Meta-Analytic CI

How many **individual** study estimates fall outside their respective meta-analytic 95% confidence intervals?

**Definitions:**

1.  **Total Estimates:** The count of individual study estimates analyzed
2.  **Outside CI (n):** Number of estimates falling outside the CI
3.  **Outside CI (%):** Percentage of estimates falling outside the CI

```{r}
#| label: included-preparation


# Join Results with Inputs
included_data <- df_mas %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    # Check Study 1
    s1_outside = est_1 < lower | est_1 > upper,
    # Check Study 2
    s2_outside = est_2 < lower | est_2 > upper
  )

# Calculate summary statistics
included_summary <- included_data %>%
  group_by(method) %>%
  summarise(
    # Total estimates = 2 per MA
    n_estimates_total = n() * 2,
    # Count how many are outside
    n_estimates_outside = sum(s1_outside, na.rm = TRUE) + sum(s2_outside, na.rm = TRUE),
    prop_outside = n_estimates_outside / n_estimates_total,
    .groups = "drop"
  )
```

```{r}
#| label: tbl-included

included_summary %>%
  gt() %>%
  tab_header(
    title = md("**Is the study contained in the Meta-Analytic CI?**")
  ) %>%
  cols_label(
    method = "Method",
    n_estimates_total = "Total Estimates",
    n_estimates_outside = "Outside CI (n)",
    prop_outside = "Outside CI (%)"
  ) %>%
  fmt_number(
    columns = c(n_estimates_total, n_estimates_outside),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_percent(
    columns = prop_outside,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  data_color(
    columns = prop_outside,
    method = "numeric",
    palette = c("#ffffff", "#ffe5e5", "#ffcccc", "#ff9999"),
    domain = c(0, max(included_summary$prop_outside))
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

**NOTE** it can be proven that Edgington will always contain the singular study estimates in his 95% CI, so this result was predictable.

## CIs informativness

For a given meta-analysis $x$, let the 95% confidence interval for study $i$ be denoted by $[L_{i,x}, U_{i,x}]$, where $L_{i,x}$ is the lower bound and $U_{i,x}$ is the upper bound. Specifically for the two-study case, we have $L_{1,x}, U_{1,x}$ for study 1 and $L_{2,x}, U_{2,x}$ for study 2.

Define: $$
[L_{\text{min}, x}, U_{\text{max}, x}] \quad \text{s.t.} \quad L_{\text{min}, x} = \min(L_{1,x}, L_{2,x}) \quad \text{and} \quad U_{\text{max}, x} = \max(U_{1,x}, U_{2,x})
$$

Then denote the meta-analytic confidence interval by $[L_{\text{MA}, x}, U_{\text{MA}, x}]$.

We define a meta-analytic CI as **informative** if it is contained entirely within the range of the individual study limits. $$
L_{\text{MA}, x} > L_{\text{min}, x} \quad \text{and} \quad U_{\text{MA}, x} < U_{\text{max}, x}
$$

As a contrary, an interval is defined as *non-informative* if it fails to meet this criterion (i.e., it is not strictly narrower than the study range).

```{r}
#| label: informativness-preparation

info_data <- df_mas %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    informative = (lower > min_study_lower & upper < max_study_upper)
  )

summ <- info_data %>%
  group_by(method) %>%
  summarise(
    total = n(),
    n_informative = sum(informative, na.rm = TRUE),
    prop_informative = mean(informative, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
#| label: tbl-informativness

summ %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Informativness**")
  ) %>%
  cols_label(
    method = "Method",
    total = "Total Estimates",
    n_informative = "Informative CI (n)",
    prop_informative = "Informative CI (%)"
  ) %>%
  fmt_percent(
    columns = prop_informative,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

# Lemma 4.3.1 (Two-study case)

::: {.callout-note icon="false"}

Consider a meta-analysis with exactly two studies, with point estimates $\hat\theta_1, \hat\theta_2$ and standard errors $\sigma_1, \sigma_2 > 0$. Let the confidence level be $1-\alpha$ with $0 < \alpha < 1/4$, and define:

$$
z_{\alpha/2} := \Phi^{-1}(\alpha/2), \qquad z_{\sqrt{\alpha/4}} := \Phi^{-1}\!\left(\sqrt{\alpha/4}\right)
$$

where $\Phi^{-1}$ denotes the quantile function of the standard normal distribution. Let $\text{width}_E$ be the width of the $(1-\alpha)$-confidence interval based on the unweighted Edgington method, and $\text{width}_{FE}$ the width of the fixed-effect confidence interval.

Set $\Delta\theta := |\hat\theta_1 - \hat\theta_2|$. If at least one of the following two conditions holds, then $\text{width}_E < \text{width}_{FE}$.

**Condition 1:** $$
\frac{\max(\sigma_1,\sigma_2)}{\min(\sigma_1,\sigma_2)} \;<\; \sqrt{\left(\frac{z_{\alpha/2}}{z_{\sqrt{\alpha/4}}}\right)^2 - 1}
$$ and $$
\Delta\theta \;<\; -\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big|
$$

**Condition 2:** $$
-\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big| \;\le\; \Delta\theta \;<\; \frac{\sigma_1 \sigma_2}{\sqrt{\sigma_1^2 + \sigma_2^2}}\, \big(-2 z_{\alpha/2}\big) \;+\; (\sigma_1 + \sigma_2)\,z_{\sqrt{\alpha/4}}
$$
:::

```{r}
#| label: fig-lemma
#| fig-height: 8

alpha <- 0.05

# Check lemma conditions (commented out since I already compute it in processing)
#cis_lemma <- check_lemma_conditions(cis, alpha = alpha)

# Extract and summarize results
lemma_detailed <- extract_lemma_results(cis_lemma)


lemma_applicable <- lemma_detailed %>%
  filter(applicable == TRUE)

# Plot results
plot_lemma_results(cis_lemma)
```

```{r}
#| label: tbl-lemma-summary

# Create simple summary table
lemma_summary <- lemma_applicable %>%
  summarise(
    "Total 2-Study MAs" = n(),
    "Condition 1 Holds" = sum(condition1_holds, na.rm = TRUE),
    "Condition 2 Holds" = sum(condition2_holds, na.rm = TRUE),
    "Both Conditions" = sum(condition1_holds & condition2_holds, na.rm = TRUE),
    "Neither Condition" = sum(!condition1_holds & !condition2_holds, na.rm = TRUE),
    "Edgington < FE" = sum(edgington_smaller, na.rm = TRUE),
    "Lemma Verified" = sum(lemma_verified == TRUE, na.rm = TRUE),
    "Lemma Failed" = sum(lemma_verified == FALSE, na.rm = TRUE)
  ) %>%
  pivot_longer(
    everything(),
    names_to = "Metric",
    values_to = "Count"
  ) %>%
  mutate(
    Percentage = ifelse(
      Metric == "Total 2-Study MAs",
      NA_real_,
      Count / first(Count) * 100
    )
  )

# Create table
lemma_summary %>%
  gt() %>%
  tab_header(
    title = md("**Lemma 4.3.1 Summary Statistics**"),
  ) %>%
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_number(
    columns = Percentage,
    decimals = 1,
    suffix = "%"
  ) %>%
  sub_missing(
    columns = Percentage,
    missing_text = "—"
  ) %>%
  cols_align(
    align = "left",
    columns = Metric
  ) %>%
  cols_align(
    align = "center",
    columns = c(Count, Percentage)
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  opt_stylize(style = 6, color = "blue")

```

{{< pagebreak >}}

```{r}
#| label: prep-p1-p2



extract_individual_p <- function(ma_obj) {
  
  # Calculate directional ONE-SIDED p-values (Alternative: "Greater")
  studies <- ma_obj$inputs
  z_scores <- studies$estimate / studies$SE
  
  # probability of getting an effect GREATER than observed.
  p_vals_studies <- 1- stats::pnorm(z_scores)
  
  p1 <- p_vals_studies[1]
  p2 <- p_vals_studies[2]
  
  # ma_obj$p_0 contains the TWO-SIDED doubled p-values 
  data.frame(
    MA = ma_obj$ma_id,
    p1 = p1,
    p2 = p2,
    method = names(ma_obj$p_0),
    p_combined = as.numeric(ma_obj$p_0),
    measure = ma_obj$measure,
    stringsAsFactors = FALSE
  )
}

# Bind
df_p <- do.call(rbind, lapply(cis, extract_individual_p))


target_methods <- c("Edgington", "Fixed effect", "Edgington (1/SE)", "Edgington (1/SE^2)")
df_plot <- df_p %>% 
  filter(method %in% target_methods)
```
# APPENDIX
## Additional stuff on two trials p values

In the paper *Beyond the two-trials rule* there is a nice plot showing the property of the Edgington combination.

Here I fix ($\alpha = 0.05$) and compute **one-sided** p-values for the individual studies under ($H_1:\theta \ge 0$). For study i:

$$
z_i=\frac{\hat\theta_i}{\sigma_i},
\qquad
p_i = 1-\Phi(z_i),
$$ Under the null, the p-values ($p_1,p_2$) are i.i.d. Uniforms (assuming independent studies). Define the Edgington statistic:

$$
S = p_1+p_2 \in [0,2].
$$

Since S is the sum of two independent Uniform((0,1)) variables, it follows the Irwin–Hall distribution with (n=2), i.e. a triangular distribution. Its CDF is

$$
F_S(s)=
\begin{cases}
\dfrac{s^2}{2}, & 0\le s\le 1,\\[6pt]
1-\dfrac{(2-s)^2}{2}, & 1<s\le 2.
\end{cases}
$$

The **two-sided** Edgington combined p-value computed in confMeta is:

$$
p_{\mathrm{Edg}}(S_{\mathrm{obs}})
=
2\min\!\Bigl(F_S(S_{\mathrm{obs}}),\,1-F_S(S_{\mathrm{obs}})\Bigr).
$$

In the special case (n=2), this simplifies to

$$
p_{\mathrm{Edg}}(S_{\mathrm{obs}})
=
\min\!\Bigl(S_{\mathrm{obs}}^2,\,(2-S_{\mathrm{obs}})^2\Bigr).
$$

To reject (H_0) at level ($\alpha$), we require ($p_{\mathrm{Edg}}(S\_{\mathrm{obs}})\le \alpha$). This yields two disjoint rejection regions:

-   **Lower (bottom-left) corner:**

    $$
    2F_S(s)\le\alpha
    \;\Longleftrightarrow\;
    2\cdot\frac{s^2}{2}\le\alpha
    \;\Longleftrightarrow\;
    s\le\sqrt{\alpha},
    $$ hence

    $$
    \mathcal R_{\mathrm{lower}}
    =
    \{(p_1,p_2): p_1+p_2 \le \sqrt{\alpha}\}.
    $$

-   **Upper (top-right) corner:**

    $$
    2(1-F_S(s))\le\alpha
    \;\Longleftrightarrow\;
    2\cdot\frac{(2-s)^2}{2}\le\alpha
    \;\Longleftrightarrow\;
    s\ge 2-\sqrt{\alpha},
    $$ hence

    $$
    \mathcal R_{\mathrm{upper}}
    =
    \{(p_1,p_2): p_1+p_2 \ge 2-\sqrt{\alpha}\}.
    $$

Basically, ($\mathcal R_{lower}$) corresponds to both studies providing strong evidence in the ($H_1:\theta \ge 0$) direction (small one-sided p-values), while ($\mathcal R_{upper}$) corresponds to both studies providing strong evidence in the opposite direction (one-sided p-values close to 1).

```{r}
#| label: fig-pvalue-final
#| fig-width: 10
#| fig-height: 8

# Define Alpha 
alpha <- 0.05

# For Edgington (n=2), the critical is sqrt(alpha)
# this comes from triangular distribution, E = p1 + p2 and to get p val we do E^2 /2
# 2 * (E^2 / 2) = 0.05  -> E = sqrt(0.05)
#Note there is also the upper corner to take into account

critical_value <- sqrt(alpha) 

# Add significance data
df_plot <- df_plot %>%
  mutate(
    is_significant = p_combined < alpha
  )

ggplot(df_plot, aes(x = p1, y = p2)) +
  geom_point(aes(color = is_significant, shape = is_significant), 
             alpha = 0.6, size = 2) +
  # Lower corner (Positive agrrement)
  # Condition: p1 + p2 < 0.2236
  geom_abline(intercept = critical_value, slope = -1, 
              linetype = "dashed", color = "black") +
  
  # Upper corner (Negative agreement)
  # Condition: (1 - p1) + (1 - p2) < 0.2236
  # this because: 2 - (p1 + p2) < 0.2236  =>  p1 + p2 > 2 - 0.2236
  geom_abline(intercept = 2 - critical_value, slope = -1, 
              linetype = "dashed", color = "black") +
  
  facet_wrap(~ method) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 1 (One sided)") +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 2 (One sided)") +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey80"), name = "Significant") +
  scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1), name = "Significant") +
  coord_fixed() + 
  theme_bw() 
```

## let's map on the lower corner

This is the exact same as before, but now we map all singular study p values to two sided before plotting them, so we only need the bottom left triangle.

Even if it seems a good idea at first, there are cases in which we have $p_1 \approx 1$ and $p_2 \approx 0$ (or the inverse). Those cases get mapped in the bottom left corner, but they are not significant. That is why there are points in the Edgington graph that are inside the triangle but are not significant, so I guess the previous plot is better.

```{r}
#| label: fig-pvalue-twosdied
#| fig-width: 10
#| fig-height: 8

df_two_sided <- df_plot %>%
  mutate(
    p1_two_sided = pmin(p1, 1 - p1),
    p2_two_sided = pmin(p2, 1 - p2),
    is_significant = p_combined < 0.05
  )

s_crit <- sqrt(0.05)

ggplot(df_two_sided, aes(x = p1_two_sided, y = p2_two_sided)) +
    geom_point(aes(color = is_significant, shape = is_significant), 
             alpha = 0.6, size = 2) +
    geom_abline(intercept = s_crit, slope = -1, 
              linetype = "dashed", color = "black") +
    facet_wrap(~ method) +
  
  # Scales: Zoom in to the relevant area (0 to 0.5)
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 1 (two sided") +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 2 (two sided)") +
  
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey80")) +
  scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1)) +
  coord_fixed() + 
  theme_bw() 
```

## Inspection of absolute difference of singular p values

Here, I plot the absolute difference of singular p values against the combined p value. This may be viewed as a similar analysis as relative width increase, but adding the interactive part, we can easily spot those meta analysis with very different individual p values.

```{r}


df_abs_diff <- df_p %>%
  select(MA, method, p1, p2, p_combined) %>%
  mutate(
    p_diff_abs = abs(p1 - p2),
    min_p = pmin(p1, p2),
    max_p = pmax(p1, p2)
  )

#add text for plotly 
df_abs_diff <- df_abs_diff %>% 
  mutate (
    text = paste0 ("<b>MA ID:</b> ", MA, "<br>",
      "<b>p1:</b> ", round(p1, 4), "<br>",
      "<b>p2:</b> ", round(p2, 4), "<br>",
      "<b>Combined p:</b> ", round(p_combined, 4), "<br>")
  )

p <- ggplot(df_abs_diff, aes(x = p_diff_abs, y = p_combined, color = method)) +
  geom_point(alpha = 0.3, aes(text = text)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Absolute Difference |p1 - p2|",
    y = "Combined P-value"
  ) +
  theme_bw() +
  facet_wrap(~method)

ggplotly(p, tooltip = "text")

```

