---
title: 'SUMMARY REPORT for new IQWiG data (restricted to 2 studies)'
author:
  - 'Saverio Fontana'
date: today
format:
  html:
      toc: true
      number-sections: true
      code-fold: true
      embed-resources: true
      smooth-scroll: true
      anchor-sections: true 
  pdf:
     toc: true
     number-sections: true
     colorlinks: true
execute:
  echo: !expr "knitr::is_html_output()"
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false

load_libraries <- function(...) {
  pkgs <- as.character(match.call(expand.dots = FALSE)$...)
  
  for (pkg in pkgs) {
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      
      # If not there (FALSE), install and then load it
      install.packages(pkg)
      library(pkg, character.only = TRUE, quietly = TRUE)
    }
  }
}

# Install and load required packages
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
if (!requireNamespace("confMeta", quietly = TRUE)) {
  remotes::install_github("SaveFonta/confMeta", upgrade = "never")
}

# Load all libraries
load_libraries(
  "confMeta", "bayesmeta", "dplyr", "tibble", "tidyr", 
  "ggplot2", "patchwork", "gt", "parallel", 
  "ggbeeswarm", "purrr", "ggridges", "ggforce", "plotly", "irr",
  "ggh4x"
)

# Set global theme for ggplot
theme_set(theme_minimal(base_size = 11))

```

```{r}
#| label: load-data
#| include: false

source("00_utilities.R") 

name <- "cis.rds"
inp <- file.path("Output", name)
input <- readRDS(inp)
cis <- input$cis
df_estimates <- input$df_estimates

```

# Data Overview

We are inspecting a total of **`r length(unique(df_estimates$identifier))`** meta-analysis. Those are all made up of 2 studies.

## Sheet division

Here is an overview of how many meta-analyisis are there in each Excel sheet

```{r}
#| label: tbl-Excel

a <- df_estimates %>%
   group_by(sheet_name) %>%
   summarize(n_MA = n()/2)

a %>% 
  gt() %>%
  #  Relabel Columns
  cols_label(
    sheet_name = md("**Data Sheet Name**"),
    n_MA = md("**Number of Meta-Analyses (N)**")
  )
```

## Measure employed

This is an overview of the different measures employed over the file

```{r}
#| label: tbl-measures


df_estimates %>%
  count(effect.measure, .drop = FALSE)  %>%   
  mutate(n = n/2 ) %>% 
   arrange(desc(n)) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    effect.measure = md("**How measures are reported**"),
    n = md("**Number**")
  )

```

## Data formatting

Another interesting fact was the way IQWIG reported the raw data. I classified them in 15 possible ways. This also show us which are the most common ways of reporting Meta Anslysis.

-   **2x2 OR**: Binary outcome reported as a 2×2 event table (a,b,c,d); effect computed as log odds ratio from counts.

-   **2x2 RR**: Binary outcome reported as a 2×2 event table; effect computed as log risk ratio from counts.

-   **2x2 RD**: Binary outcome reported as a 2×2 event table; effect computed as risk difference from counts.

-   **HR**: Study reports the log-Hazard Ratio with its standard error directly.

-   **direct OR**: Study reports the log-Odd Ratio with its standard error directly.

-   **direct RR**: Study reports the log-Risk Ratio with its standard error directly.

-   **direct RD**: Study reports the risk difference with its standard error directly.

-   **ROM**: Study reports the ratio of means on the log scale with its standard error directly.

-   **IDR**: Study reports the incidence density rate ratio on the log scale with its standard error directly.

-   **direct SMD**: Study reports the standardized mean difference with its standard error directly.

-   **direct MD**: Study reports the mean difference with its standard error directly.

-   **MD pooled SD**: Continuous outcome: study reports mean difference (MD), pooled SD, and group sizes (n1, n2).

-   **SMD from pooled**: Continuous outcome: study reports MD, pooled SD, and group sizes; SMD (Hedges’ g) is derived from these.

-   **singular studies MD**: Continuous outcome: study reports group means and SDs (ȳ1, s1, n1; ȳ2, s2, n2); MD is computed.

-   **singular studies SMD**: Continuous outcome: study reports SDs and group means; SMD is computed.

```{r}
#| label: tbl-formatting


df_estimates %>%
  count(data_report, .drop = FALSE)  %>%   
   arrange(desc(n)) %>% 
  mutate (n = n/2) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    data_report = md("**How data is reported**"),
    n = md("**Number**")
  )
```

## Methods Employed:

-   **Edgington**

-   **Weighted Edgington** with weights $\frac{1}{SE}$ (Inverse SE weighted)

-   **Weighted Edgington** with weights $\frac{1}{SE^2}$ (Inverse Variance weighted)

-   **Fixed Effect**. When tabular data are provided, the default is Mantel-Haenszel method, otherwise we fall back to Inverse-Variance method.

-   **Hartung-Knapp** (2001a/b) and variance correction is applied if HK confidence interval is narrower than CI from classic random effects model with DerSimonian-Laird estimator (IQWiG, 2022)



# Summary measures

This summary is computed using all the MA from above.

**Scale** All ratio-based metrics (OR, RR, HR, etc.) are analyzed on the logarithmic scale. However, to remain consistent with IQWiG reporting conventions, they are labeled using their original metric names.


## Width of the confidence intervals

For each method, we summarize the distribution of **CI widths** (upper − lower) across all two-study meta-analyses.

```{r}
#| label: tbl-widths

widths <- extract_ma_metric(cis, "width")

table_width  <- compute_summary (widths, stats = "width")


render_gt (table_width, title = "Confidence Interval Widths")

```


### Width by effect measure

Widths stratified by effect measure

```{r}
#| label: tbl-widths-bymeasure


width_summary <- compute_summary(
  widths, 
  stats = "width", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)


render_gt(
  width_summary,
  title = "Confidence Interval Widths by Effect Measure and Method",
  group_col = "measure"
  )


```

**OBSERVATION**: Note the "staircase" effect in both median and mean: *Unweighted Edgington \> Edgington 1/SE \> Edgington 1/SE\^2 \> Fixed Effect* FOR EACH MEASURE.

This is the explanation I suggest:

For classic **Edgington** (Equal Weighting), each study contributes equally to the combined p-value function, regardless of its sample size or precision. Consequently, *the high variance of smaller, "noisy" studies is fully integrated into the pooled result.* This means we are forcing the precise studies to "compromise" with the imprecise ones, resulting in widest CI.

But **the more aggressively we penalize large standard errors** (moving from Unweighted $\rightarrow$ $1/SE$ $\rightarrow$ $1/SE^2$), the more you force the meta-analysis to *ignore uncertainty*. Basically we are giving more influence to certain data points and reducing the influence of uncertain ones.

---

For a better visualization, the HK method is removed and ROM and RD are not plotted, since they only have respectively 1 and 6 observations.

**Note** that MD and SMD are reported on the square root scale

```{r}
#| label: fig-widths-bymeasure
#| fig-width: 20
#| fig-height: 15
#| fig-dpi: 300

widths_noHK <- widths %>% 
  filter (method != "Hartung & Knapp") %>% 
  filter (measure != "ROM" & measure != "RD")

ggplot(widths_noHK, aes(x = method, y = width, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~ measure, scales = "free_y") +
  labs(
    title = "Distribution of Confidence Interval Widths Stratified by measure",
    x = "Method",
    y = "CI Width"
  ) +
    facetted_pos_scales(
    y = measure %in% c("SMD", "MD") ~ scale_y_sqrt()
  ) + 
  theme_bw() +
  theme(
    plot.title   = element_text(face = "bold", size = 20),
    strip.text   = element_text(face = "bold", size = 16),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y  = element_text(size = 13),
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13),
    legend.position = "bottom",
    strip.background = element_rect(fill = "lightblue")
  ) +
  scale_fill_brewer(palette = "Set2")
```


::: {.callout-note icon="false"}
**My take on why HK gives wild CI**

The HK method produces extremely wide confidence intervals. This probably because both the variance estimator and the critical value from the t-distribution become unstable.

In the HK approach, the pooled effect estimate is the weighted mean

$$
\hat{\mu} = \frac{\sum_{i=1}^{k} w_i \hat{\theta}_i}{\sum_{i=1}^{k} w_i},
\qquad
w_i = \frac{1}{s_i^2 + \hat{\tau}^2},
$$

where $\hat{\theta}_i$ is the effect estimate from study $i$, $s_i^2$ its within-study variance, and $\hat{\tau}^2$ the estimated between-study variance.

The HK estimator of the variance of $\hat{\mu}$ is based on a weighted residual sum of squares,

$$
\widehat{\mathrm{Var}}(\hat{\mu})
=
\frac{1}{k-1}
\,
\frac{
\sum_{i=1}^{k} w_i^{2} \, (\hat{\theta}_i - \hat{\mu})^{2}
}{
\left(\sum_{i=1}^{k} w_i\right)^{2}
}.
$$

For the $(1-\alpha)$-level confidence interval, HK uses the $t$-distribution with $k-1$ degrees of freedom:

$$
\hat{\mu}
\pm
t_{k-1,\,1-\alpha/2}
\sqrt{\widehat{\mathrm{Var}}(\hat{\mu})}.
$$

With $k=2$ there are two problems.

The degrees of freedom are $k-1 = 1$, so the critical value $t_{1,1-\alpha/2}$ is very large (for example, $t_{1,0.975} \approx 12.7$), which directly inflates the length of the interval. Second, the variance estimator is essentially a scaled residual based on only one degree of freedom:

$$
\widehat{\mathrm{Var}}(\hat{\mu})
=
\frac{1}{1}
\,
\frac{
\sum_{i=1}^{2} w_i^{2} \, (\hat{\theta}_i - \hat{\mu})^{2}
}{
\left(\sum_{i=1}^{2} w_i\right)^{2}
},
$$

so it depends on a single random quantity and can vary wildly even under small changes in the observed effects.
:::

---

{{< pagebreak >}}

## Width of the CI compared to individual SE

Following my observation on the "staircase", I want to empirically prove that **using the weighting schemes, we tend to ignore the studies that are largely uncertain.** This mean that as the ratio of Standard Errors $\frac{max(SE)}{min(SE)}$ increases, Edgington will produces **wider intervals.** This because It is *"forced" to incorporate the high uncertainty (large SE) of the smaller study into the pooled p-value function.*

At the same time, as $SE_{ratio}$ grows, the weigthed methods center their estimate and precision more and more exclusively on the "certain" study, **reducing the contribution of the high-variance study**.

So we compute the **Pearson correlation between $SE_{ratio}$ and width** per method.

```{r}
#| label: tbl-widths-vs-se


# prepare two Dfs taht will also be used later
ma_ci_df <- bind_rows(
  lapply(cis, function(x) {
    x$ci %>% 
      as.data.frame() %>% 
      mutate (
        estimate_ma = estimate,
        lower_MA = lower,
        upper_MA = upper,
        width_MA = width,
        significant_MA = significant,
        MA = x$ma_id
      ) %>% 
      select (c(MA, method, estimate_ma,lower_MA, upper_MA, width_MA, significant_MA))
  }) )


input_df <- bind_rows(
  lapply(cis, function(x) {
    x$inputs %>%  
      as.data.frame() %>% 
      mutate (
        MA = x$ma_id,
        study = rownames(x$inputs),
        input_estimate = estimate,
        input_SE = SE,
        input_lower = lower,
        input_upper = upper,
        input_width = width,
        input_significant = significant 
      ) %>% 
      select (-c(estimate, SE, lower, upper, width, significant))
  }))





SEs_df <- input_df %>% 
  group_by (MA) %>% 
  summarise (
    SE_1 = input_SE[1],
    SE_2 = input_SE[2],
    SE_ratio = max(input_SE) / min(input_SE),
    .groups= "drop"
  )

merged_se_df <- merge(ma_ci_df, SEs_df, by = "MA") 

corr_df <- merged_se_df %>%
  group_by(method) %>%
  summarise(
    Pearson = cor(width_MA, SE_ratio, use = "complete.obs", method = "pearson"),
    Spearman = cor(width_MA, SE_ratio, use = "complete.obs", method = "spearman"),
    .groups = "drop"
  )

# Reshape the table
corr_vec_df <- corr_df %>%
  pivot_longer(
    cols = c(Pearson, Spearman), 
    names_to = "Metric", 
    values_to = "Correlation"
  ) %>%
  pivot_wider(
    names_from = method, 
    values_from = Correlation
  )

# 3. Generate the formatted gt table
corr_vec_df %>%
  gt() %>%
  fmt_number(
    columns = -Metric,
    decimals = 3
  ) %>%
  tab_header(
    title = md("**Correlation between SE ratio and Width**"),
  ) %>% 
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_stylize(style = 6, color = "blue")

```




{{< pagebreak >}}

## Confidence interval skewness

The skewness of each confidence interval is computed as:

Of course Fixed Effect mehods cannot be skewed so we don't report them. $$
\text{Skewness} = \frac{\text{upper} + \text{lower} - 2 \times \text{estimate}}{\text{upper} - \text{lower}}
$$

```{r}
#| label: tbl-skewness

ci_skewness <- extract_ma_metric(cis, "ci_skewness")

ci_skewness <- ci_skewness %>% filter (!(method %in% c("Fixed effect", "Hartung & Knapp")))
table_skewness <- compute_summary (ci_skewness, stats = "ci_skewness")


render_gt (table_skewness, title = "Confidence Interval skewness")

```

{{< pagebreak >}}

## Is the study estimate contained in the Meta-Analytic CI

How many **individual** study estimates fall outside their respective meta-analytic 95% confidence intervals?

**Definitions:**

1.  **Total Estimates:** The count of individual study estimates analyzed
2.  **Outside CI (n):** Number of estimates falling outside the CI
3.  **Outside CI (%):** Percentage of estimates falling outside the CI

```{r}
#| label: included-preparation



# Merge
merged_df <- merge(input_df, ma_ci_df, by = "MA") 

merged_df <- merged_df %>%
  mutate(
    outside_CI = input_estimate < lower_MA | input_estimate > upper_MA
  )

# Calculate summary statistics
included_summary <- merged_df %>%
  group_by(method) %>%
  summarise(
    n_estimates_total = n(),
    n_estimates_outside = sum(outside_CI, na.rm = TRUE),
    prop_outside = n_estimates_outside / n_estimates_total,
    .groups = "drop"
  )
```

```{r}
#| label: tbl-included

included_summary %>%
  gt() %>%
  tab_header(
    title = md("**Is the study contained in the Meta-Analytic CI?**")
  ) %>%
  cols_label(
    method = "Method",
    n_estimates_total = "Total Estimates",
    n_estimates_outside = "Outside CI (n)",
    prop_outside = "Outside CI (%)"
  ) %>%
  fmt_number(
    columns = c(n_estimates_total, n_estimates_outside),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_percent(
    columns = prop_outside,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  data_color(
    columns = prop_outside,
    method = "numeric",
    palette = c("#ffffff", "#ffe5e5", "#ffcccc", "#ff9999"),
    domain = c(0, max(included_summary$prop_outside))
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

**NOTE** it can be proven that Edgington will always contain the singular study estimates in his 95% CI, so this result was predictable.

{{< pagebreak >}}

## CIs informativness

For a given meta-analysis $x$, let the 95% confidence interval for study $i$ be denoted by $[L_{i,x}, U_{i,x}]$, where $L_{i,x}$ is the lower bound and $U_{i,x}$ is the upper bound. Specifically for the two-study case, we have $L_{1,x}, U_{1,x}$ for study 1 and $L_{2,x}, U_{2,x}$ for study 2.

Define: $$
[L_{\text{min}, x}, U_{\text{max}, x}] \quad \text{s.t.} \quad L_{\text{min}, x} = \min(L_{1,x}, L_{2,x}) \quad \text{and} \quad U_{\text{max}, x} = \max(U_{1,x}, U_{2,x})
$$

Then denote the meta-analytic confidence interval by $[L_{\text{MA}, x}, U_{\text{MA}, x}]$.

We define a meta-analytic CI as **informative** if it is contained entirely within the range of the individual study limits. $$
L_{\text{MA}, x} > L_{\text{min}, x} \quad \text{and} \quad U_{\text{MA}, x} < U_{\text{max}, x}
$$

As a contrary, an interval is defined as *non-informative* if it fails to meet this criterion (i.e., it is not strictly narrower than the study range).

```{r}
#| label: informativness-preparation


bounds_df <- input_df %>% 
  group_by(MA) %>% 
  summarise(
    studies_lower = min(input_lower),
    studies_upper = max(input_upper),
    .groups = "drop"
  )

# Merge
merged_df <- merge(ma_ci_df, bounds_df, by = "MA")

merged_df <- merged_df %>% 
  mutate(
    # MA CI is informative if it is NARROWER than the studies range
    informative = (lower_MA > studies_lower & upper_MA < studies_upper)
  )

summ <- merged_df %>% 
  group_by(method) %>% 
  summarise(
    total = n(),
    n_informative = sum(informative),
    prop_informative = mean(informative),
    .groups = "drop"
  )
```

```{r}
#| label: tbl-informativness

summ %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Informativness**")
  ) %>%
  cols_label(
    method = "Method",
    total = "Total Estimates",
    n_informative = "Informative CI (n)",
    prop_informative = "Informative CI (%)"
  ) %>%
  fmt_percent(
    columns = prop_informative,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```


{{< pagebreak >}}

## p-values (for $H_0: \theta=0$)

Distribution of method-specific p-values for testing $H_0: \theta=0$

*Note* the dashed line marks 0.05.

```{r}
#| label: pvals-preparation

pvals <- extract_ma_metric(cis, "p_0")


pvals %>%
  ggplot(aes(x = method, y = p_0, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.8) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Method",
    y = "p-value",
    title = "Distribution of p-values by method"
  ) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  coord_cartesian(ylim = c(0, 1))

```

```{r}
#| label: tbl-pvals



table_p0 <- compute_summary(pvals, stats = "p_0")


render_gt (table_p0, title = "P value distribution")

 
```

This aligns with previous observations: since FE tends to produce the narrowest intervals, it is the **more "aggressive" at rejecting the null.**

### Cumulative Distribution of P-values

In this graph the **ECDF** of p-values is plotted. The grey dashed line is the CDF of the Uniform distribution.

```{r}
#| label: fig-pvalue-ecdf
#| fig-height: 7



pvals %>%
  ggplot(aes(x = p_0, color = method)) +
  stat_ecdf(geom = "step", linewidth = 1, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", 
              color = "gray40", linewidth = 0.8) +
  annotate("text", x = 0.7, y = 0.3, 
           label = "Uniform distribution\n(under global null)", 
           color = "gray40", size = 3.5) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "P-value",
    y = "Cumulative Probability",
    color = "Method"
    ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
  
```

{{< pagebreak >}}

### P-value Comparison to Fixed Effect

Now, we will compare each method's p-values with **Fixed Effect's p-values as baseline.**

```{r}
#| label: pvals-vs-FE-preparation

pvals_wide <- pvals %>%
  select(MA, method, p_0) %>%
  pivot_wider(names_from = method, values_from = p_0)

# Calculate differences from Fixed Effect
only_fe <- pvals %>% 
  filter(method == "Fixed effect") %>% 
  select (MA, p_0_fe = p_0)

pvals_diff <- pvals %>% 
  left_join(
    only_fe,
    by = "MA"
  ) %>% 
  mutate (p_diff = p_0 - p_0_fe) %>% 
  filter (method != "Fixed effect")
```

Here, each panel plots the FE p-values vs the p-value of the specific method.

*Note*: if the $p_0$ were equal, points would distribute on the diagonal.

**Each point is one meta-analysis.** Points below the diagonal indicate smaller p-values than fixed effect (less conservative), and points above indicate larger p-values (more conservative).

```{r}
#| label: fig-pvalue-vs-fe
#| fig-height: 8

pvals_diff %>% 
  ggplot (aes(x = p_0_fe, y = p_0, color = method)) +
  geom_point(alpha= 0.4, size = 1.5) +
  geom_abline (intercept = 0, slope = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) + 
  facet_wrap (~method) + 
  scale_x_continuous (breaks = seq(0,1, 0.2)) + 
  scale_y_continuous(breaks = seq(0,1, 0.2)) + 
  labs(
    x = "P-value (Fixed Effect)",
    y = "P-value (Method)",
    title = "Method p-values vs Fixed Effect p-values"
  ) +
  theme(
    legend.position = "none"
  )
```

{{< pagebreak >}}

### P-value Differences from Fixed Effect

This table is similar to @tbl-pvals but with a focus on the difference $\Delta p = p_0 (method) - p_0 (FE)$ and showing the correlation between $p_0(method)$ and $p_0 (FE)$ .

```{r}
#| label: tbl-pvals-differences

pvals_diff %>%
  group_by(method) %>%
  summarise(
    mean_diff = mean(p_diff, na.rm = TRUE),
    median_diff = median(p_diff, na.rm = TRUE),
    sd_diff = sd(p_diff, na.rm = TRUE),
    pearson = cor(p_0, p_0_fe, method = "pearson", use = "complete.obs"),
    spearman = cor(p_0, p_0_fe, method = "spearman", use = "complete.obs"),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**P-value Differences from Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    mean_diff = "Mean Δ",
    median_diff = "Median Δ",
    sd_diff = "SD Δ",
    pearson = "Pearson",
    spearman = "Spearman"
  ) %>%
  tab_spanner(
    label = "Correlation with FE",
    columns = c(pearson, spearman)
  ) %>%
  fmt_number(
    columns = c(mean_diff, median_diff, sd_diff, pearson, spearman),
    decimals = 4
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

Quite interestingly, the **median difference equall zero.** (at least half of observation have 0 difference with FE method)

However, the slightly higher **mean difference (0.02)** suggests that in a subset of cases, likely those w/ high study-level conflict, the Edgington method provides a more conservative p-value than the Fixed Effect model.


### P values for different sign estimates

To better inspect the problem, I statify between MAs that show **opposite signs of the effect estimates** (e.g. study 1 has positive estimate, while study 2 has negative estimate, or the opposite) and that show the same sign.

The idea is that in those scenarios, the Fixed Effect model remains dominated by the more precise study, maintaining statistical significance despite the contradictory evidence. While Edgington method will provide extremely high p values since the combination of two contrasting estimates will tend to not reject the null.

```{r}
#| label: tbl-pvals-differences-inspection


# Join p-value differences with the raw study inputs
conflict_analysis <- pvals_diff %>%
  filter(method != "Hartung & Knapp") %>%
  left_join(input_df %>% 
              group_by(MA) %>% 
              summarise(
                est_1 = input_estimate[1],
                est_2 = input_estimate[2],
                opposing_directions = (est_1 * est_2 < 0), # True if one is + and one is -
                abs_diff = abs(est_1 - est_2)
              ), by = "MA")

# Look at the Mean p_diff when studies conflict vs when they agree
conflict_summary <- conflict_analysis %>%
  group_by(method, opposing_directions) %>%
  summarise(
    n = n(),
    mean_delta_p = mean(p_diff),
    median_delta_p = median(p_diff),
    .groups = "drop"
  )


tab_dat <- conflict_summary %>%
  mutate(
    Sign = ifelse(opposing_directions, "Opposite", "Same")
  ) %>%
  select(method, Sign, n, mean_delta_p, median_delta_p) %>%
  group_by(method) %>%
  mutate(
    method = ifelse(row_number() == 1, method, "")
  ) %>%
  ungroup()

tab_dat %>%
  gt() %>%
  cols_label(
    method = "Method",
    Sign = "Sign",
    n = "N",
    mean_delta_p = "Mean Δ",
    median_delta_p = "Median Δ"
  ) %>%
  fmt_number(
    columns = c(mean_delta_p, median_delta_p),
    decimals = 3
  ) %>%
  cols_align(
    align = "center",
    columns = -c(method, Sign)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```



This proves empirically that Edgington is a safer method (provides larger p values) when the sings of estimates are opposite. Interestingly, when the signs agree, Edgington p values tend to be **slighly** smaller.

Note also that the magnitude of the divergence decreases as the weighting scheme becomes more aggressive.

---


```{r}
#| label: fig-directions


plot_data <- pvals_diff %>%
  filter(method == "Edgington") %>%
  left_join(input_df %>% 
              group_by(MA) %>% 
              summarise(
                conflict = ifelse(input_estimate[1] * input_estimate[2] < 0, 
                                 "Directional Conflict", "Consistent Direction"),
                .groups = "drop"
              ), by = "MA")

# 2. Generate the Scatter Plot
ggplot(plot_data, aes(x = p_0_fe, y = p_0, color = conflict)) +
  geom_point(alpha = 0.4, size = 1.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey30") + # Identity line
  scale_color_manual(values = c("Consistent Direction" = "#34495e", 
                               "Directional Conflict" = "#e74c3c")) +
  theme_minimal() +
  labs(
    x = "Fixed Effect P-value",
    y = "Edgington P-value",
    color = "Sign Alignment",
    title = " P values for different sign estimates"
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold")
  )
```

---

Here just for visual understanding I display the 10 most conflicted MAs.

```{r}
#| label: tbl-pvals-extreme-cases


# Extract the Top 10 most "conflicted" MAs to showcase in a table
extreme_conflicts <- conflict_analysis %>%
  filter(opposing_directions == TRUE) %>%
  arrange(desc(p_diff)) %>%
  slice_head(n = 10)

extreme_conflicts %>%
  select(MA, measure, `p_0`, p_0_fe, p_diff, est_1, est_2, abs_diff) %>%
  gt() %>%
  cols_label(
    MA = "MA",
    measure = "Measure",
    p_0 = "p-val Edgington",
    p_0_fe = "p-val FE",
    p_diff = "p-val difference",
    est_1 = "Estimate study 1",
    est_2 = "Estimate study 2",
    abs_diff = "|Est 1 − Est 2|"
  ) %>%
  fmt_number(
    columns = c(p_0, p_0_fe, p_diff,est_1, est_2, abs_diff),
    decimals = 3
  )


```

{{< pagebreak >}}

### Agreement in Significance (p \< 0.05)

Here we inspect the significance agreement between the values at a $\alpha = 0.05$ level

Each meta-analysis is classified into **four categories:** both significant, both non-significant, FE-only, or method-only.

```{r}
#| label: fig-significance-agreement
#| fig-height: 6



sign_agree_p <- pvals_diff %>% 
  group_by(method) %>% 
  mutate (
    Method_sign = p_0 <0.05,
    FE_sign = p_0_fe <0.05, 
    agreement = ifelse(FE_sign & Method_sign, "Both significant",
                       ifelse(!FE_sign & !Method_sign, "Both non-significant",
                              ifelse(FE_sign & !Method_sign, "FE sig, Method non-sig", 
                                     "FE non-sig, Method sig"))))
  
  
  
  
  

sign_agree_p %>% 
  count(method, agreement) %>% 
  ggplot(aes(x = method, y = n, fill = agreement)) + 
           geom_col (position= "stack", alpha = 0.8) + 
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3,
            color = "white", fontface = "bold") + 
  scale_fill_manual ( 
    values = c(      "Both significant" = "#2ca02c",
                     "Both non-significant" = "#d62728",
                     "FE sig, Method non-sig" = "purple",
                     "FE non-sig, Method sig" = "#7f7f7f"
    )) + 
  labs (
    x = "Methods", 
    y = "Count",
    fill = "Agreement", 
    title = "Significance agreement with Fixed Effect (α = 0.05)"
  ) + 
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
  

         
```

```{r}
#| label: tbl-significance-agreement

sign_agree_p %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    both_sig = sum(FE_sign & Method_sign, na.rm = TRUE),
    both_nonsig = sum(!FE_sign & !Method_sign, na.rm = TRUE),
    fe_only = sum(FE_sign & !Method_sign, na.rm = TRUE),
    method_only = sum(!FE_sign & Method_sign, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Significance Agreement with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "Total",
    both_sig = "Both Sig",
    both_nonsig = "Both Non-sig",
    fe_only = "FE Only",
    method_only = "Method Only"
  ) %>%
  fmt_number(
    columns = c(n_total, both_sig, both_nonsig, fe_only, method_only),
    decimals = 0
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  )  %>%
  opt_stylize(style = 6, color = "blue")
```

Thus, this shows that classic Edgington is the most independent by FE, disagreeing with it in 89 cases ($7.2\%$), whereas the weighting progressively reduces this disagreement. Notably, 25 times Edgington was significant while FE not, **this strengthen the idea that it is not only a more conservative test.**  

{{< pagebreak >}}

## Coherence of skewness signs (β, γ, AUCC ratio)

Agreement between different measures of skewness:

-   **β (beta)**: CI skewness
-   **γ (gamma)**: Data skewness
-   **AUCC ratio**: Area Under Confidence Curve ratio
-   **Cohen’s** $\kappa$: measure of agreement between signs

where $$\gamma = \frac{\sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^3}{\left( \sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^2 \right)^{3/2} / \sqrt{\sum_{i=1}^{n} w_i}}$$ is Fisher’s weighted skewness coefficient.

And $$
\text{AUCC Ratio} = \frac {\text{AUCC}_{\text{upper}} - \text{AUCC}_{\text{lower}}} {AUCC}
$$.

*For more details see: Held, Leonhard, Felix Hofmann, and Samuel Pawel. "A comparison of combined p-value functions for meta-analysis." Research Synthesis Methods (2025)*

```{r}
#| label: sign-agreement-preparation

# Build sign agreement dataset
sign_agreement <- do.call(rbind, lapply(cis, function(x) {
  df <- x$ci 
  df <- subset(df, !is.na(ci_skewness) & !method %in% c("Fixed effect", "Hartung & Knapp", "Bayesmeta"))
  data.frame(
    MA = x$ma_id,
    method = df$method,
    data_skewness = x$data_skewness,
    ci_skewness = df$ci_skewness,
    aucc_ratio = df$aucc_ratio
  )
})) %>%
  mutate(
    same_sign_beta_gamma = sign(data_skewness) == sign(ci_skewness),
    same_sign_beta_aucc = sign(ci_skewness) == sign(aucc_ratio),
    same_sign_gamma_aucc = sign(data_skewness) == sign(aucc_ratio)
  )

```

```{r}
#| label: tbl-sign-agreement



# Calculate summary statistics
sign_summary <- sign_agreement %>%
  group_by(method) %>%
  summarise(
    n_MA = n(),
    agree_beta_gamma = sum(same_sign_beta_gamma, na.rm = TRUE),
    agree_beta_aucc = sum(same_sign_beta_aucc, na.rm = TRUE),
    agree_gamma_aucc = sum(same_sign_gamma_aucc, na.rm = TRUE),
    corr_beta_gamma = cor(ci_skewness, data_skewness, use = "complete.obs"),
    corr_beta_aucc = cor(ci_skewness, aucc_ratio, use = "complete.obs"),
    corr_gamma_aucc = cor(data_skewness, aucc_ratio, use = "complete.obs"),
     kappa_beta_gamma = {
      # We extract the signs and convert in factors
      s1 <- factor(sign(data_skewness), levels = c(-1, 0, 1))
      s2 <- factor(sign(ci_skewness), levels = c(-1, 0, 1))
      # kappa2 calculates the agreement:
      kappa2(data.frame(s1, s2))$value
    },
    .groups = "drop"
  )

# Create gt table
sign_summary %>%
  gt() %>%
  tab_header(
    title = md("**Sign Agreement**"),
  ) %>%
  tab_spanner(
    label = md("**Sign Agreement (Count)**"),
    columns = c(agree_beta_gamma, agree_beta_aucc, agree_gamma_aucc)
  ) %>%
  tab_spanner(
    label = md("**Pearson Correlation**"),
    columns = c(corr_beta_gamma, corr_beta_aucc, corr_gamma_aucc)
  ) %>%
  cols_label(
    method = "Method",
    n_MA = md("*MA*"),
    agree_beta_gamma = "β vs γ",
    agree_beta_aucc = "β vs AUCC",
    agree_gamma_aucc = "γ vs AUCC",
    corr_beta_gamma = "β vs γ",
    corr_beta_aucc = "β vs AUCC",
    corr_gamma_aucc = "γ vs AUCC",
    kappa_beta_gamma = "Cohen's κ (β vs γ)"

  ) %>%
  fmt_number(
    columns = c(starts_with("corr"), kappa_beta_gamma),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(n_MA, starts_with("agree")),
    decimals = 0
  ) %>%
  sub_missing(
    columns = everything(),
    missing_text = "–"
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

## Sign agreement among β, γ, and AUCC ratio

```{r}
#| label: fig-sign-agreement
#| fig-height: 6

sign_long <- sign_agreement %>%
  filter(!method %in% c("Fixed effect", "Hartung & Knapp")) %>%
  select(MA, method, starts_with("same_sign")) %>%
  pivot_longer(
    cols = starts_with("same_sign"),
    names_to = "comparison",
    values_to = "agreement"
  ) %>%
  filter(!is.na(agreement))
  
# Create bar plot showing proportions
sign_long %>%
  mutate(
    comparison = case_when(
      comparison == "same_sign_beta_gamma" ~ "β vs γ",
      comparison == "same_sign_beta_aucc" ~ "β vs AUCC",
      comparison == "same_sign_gamma_aucc" ~ "γ vs AUCC"
    )
  ) %>%
  group_by(method, comparison) %>%
  summarise(
    prop_agree = mean(agreement, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = method, y = prop_agree, fill = comparison)) +
  geom_col(position = position_dodge(0.8), width = 0.7, alpha = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey30") +
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.1)
  ) +
  scale_fill_manual(
    values = c("#1f77b4", "#ff7f0e", "#2ca02c")
  ) +
  labs(
    x = "Method",
    y = "Proportion of Agreement",
    fill = "Comparison"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 12),
    legend.position = "bottom"
  )
```

{{< pagebreak >}}

## Heterogeneity summary

**Cochran's** $Q$ $$Q = \sum_{i=1}^{k} w_i (\hat{\theta}_i - \hat{\theta}_{FE})^2$$where $k$ is the number of studies (here $k=2$), and $w_i = 1/\sigma_i^2$. Under the null hypothesis of homogeneity, $Q$ follows a $\chi^2$ distribution with $k-1$ degrees of freedom.

**Inconsistency Index** ($I^2$) Describes the percentage of variation across studies that is due to heterogeneity rather than chance.$$I^2 = \max\left(0, \frac{Q - (k - 1)}{Q}\right) \times 100\%$$

**Between-study Variance** ($\tau^2$) estimated using REML

**Note** 

Under homogeneity ($\tau^2 = 0$), we have $\theta_i \sim N(\mu, v_i)$. Then $\theta_1 - \theta_2 \sim N(0, v_1 + v_2)$, so $Q = \frac{(\theta_1 - \theta_2)^2}{v_1 + v_2} \sim \chi^2_1$.
And under $H_0$ we have ${Q \sim \chi^2_1}$.
Critical value at $\alpha = 0.05$: $3.84$.
So you reject iff $|\theta_1 - \theta_2| > \sqrt{3.84 (v_1 + v_2)}$.



```{r}
#| label: heterogeneity-preparation


hetero_df <- bind_rows(
  lapply(cis, function(x) {
    x$heterogeneity %>%  
      as.data.frame() %>% 
      mutate (
        MA = x$ma_id,
        Tau2 = Tau^2
      ) }
  )
) 

summary_hetero <- hetero_df %>% 
  summarise(
    total = n(),
    avg_Q = mean(Q),
    avg_pQ = mean(p_Q),
    avg_I2 = mean(I2),
    avg_tau = mean(Tau2),
    .groups = "drop"
  )

summary_hetero %>%
  gt() %>%
  cols_label(
    total = "Total MA",
    avg_Q = "Average Q value",
    avg_pQ = "Average value of  p value of Q",
    avg_I2 = "Average I2 value",
    avg_tau = "Average tau value"
  ) %>%
  fmt_number(
    columns = c(avg_Q,avg_pQ,  avg_I2, avg_tau),
    decimals = 2
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

## Stratified Q statistics

```{r}
#| label: heterogeneity-preparation-stratified

# Create a long dataframe with one row per Method per MA
q_all_methods <- bind_rows(lapply(cis, function(x) {
  
  # 1. Extract heterogeneity (Common to all methods in this MA)
  Q_val <- x$heterogeneity$Q             # -->   would it be better to use I2???
  
  # 2. Extract Method-specific metrics
  methods <- names(x$width)
  
  data.frame(
    MA = x$ma_id,
    measure = x$measure,
    Q = Q_val,
    method = methods,
    width = as.numeric(x$width),
    p_value = as.numeric(x$p_0), # P-value for H0: theta=0
    stringsAsFactors = FALSE
  )
}))

```

### Q stratified by measure

Full graph on the above, with a zoomed-in view below due to the outliers that ruin the visualisation. Data are reported on the square root scale.

```{r}
#| label: fig-heterogeneity-stratified
#| fig-height: 6

# We take unique MA rows to avoid counting Q multiple times
q_unique <- q_all_methods %>% 
  distinct(MA, .keep_all = TRUE)


#use the ggforce package --> facet_zoom allow us to zoom in


ggplot(q_unique, aes(x = Q, y = measure, color = measure)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.5) + 
  geom_jitter(height = 0.1, alpha = 0.5, size = 1) +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 500), 
    name = "Cochran's Q (Scale sqrt)"
  ) +
  
  # Note: xlim in facet_zoom must use original Q values 
  facet_zoom(xlim = c(0, 20), zoom.size = 1) + 
  
  theme_bw() +
  theme(
    legend.position = "none" 
  )

```

**NOTE**, there is a huge outlier, with id "*A20-93 \|\| Abb. 12 \|\| HbA1c*". It has a Q-statistics = 570.95.

### Summary table

```{r}
#| label: tbl-Q-vs-width-pvals

q_all_methods %>%
  group_by(method) %>%
  summarise(
    Cor_Q_Width = cor(Q, width, method = "spearman", use = "complete.obs"),
    Cor_Q_Width_p = cor(Q, width, method = "pearson", use = "complete.obs"),
    Cor_Q_Pval  = cor(Q, p_value, method = "spearman", use = "complete.obs"),
    Cor_Q_Pval_p  = cor(Q, p_value, method = "pearson", use = "complete.obs"),
    .groups = "drop"
  ) %>%
  arrange(desc(Cor_Q_Width)) %>%
  gt() %>%
  tab_header(
    title = md("**Correlation with Heterogeneity**"),
  ) %>%
  cols_label(
    method = "Method",
    Cor_Q_Width = "Corr(Q, Width) Spearman",
    Cor_Q_Width_p = "Corr(Q, Width) Pearson", 
    Cor_Q_Pval = "Corr(Q, P-val) Spearman", 
    Cor_Q_Pval_p  = "Corr(Q, P-val) Pearson"
  ) %>%
  fmt_number(columns = starts_with("Cor"), decimals = 3) %>%
  data_color(
    columns = starts_with("Cor"),
    method = "numeric",
    palette = "RdBu", # Red-Blue diverging palette
    domain = c(-1, 1)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```
The positive correlation shows that **Edgington "adapts":** When $Q$ increases (meaning studies are moving apart), the Edgington confidence interval widens to account for that conflict.

The large gap between Pearson and Spearman suggests that the relationship between heterogeneity and Edgington width is *not linear,* but it is monotonic. (e.g. Edgington doesn't widen in a straight line as $Q$ grows; but maybe stays stable for a while and then "reacts" more strongly once Q reaches a certain threshold, *which I suppose should be Q>1*).


The correlations between $Q$ and P-values for Edgington methods are essentially zero. This suggests that while Edgington widens its uncertainty (Width) when studies disagree, it doesn't necessarily lose its ability to find an effect (P-value) just because the studies are heterogeneous.
In contrast,  FE: has a negative Spearman correlation. This suggests that as studies move apart ($Q$ increases), the p-value actually tends to get "smaller" (more significant). 

Thus, in this context, Edgington appears more stable. 




## Relative Width Increase

$$
\frac{\text{Width}_{\text{Method}} - \text{Width}_{\text{FE}}}{\text{Width}_{\text{FE}}} 
$$

Basically, we are measuring **how much the interval is increasing with respect to FE.**

We plot this increase against Cochran's $Q$ to show how methods respond to heterogeneity. A positive value indicates a penalty for heterogeneity: as the study estimates diverge, the method produces a wider (more conservative) interval.

On the right, all methods are plotted together, on the right, to better inspection, the HK method is removed.

**This plot uses the LOESS smoother**. This can be misleading, it is just to detect a general trend, next plots will inspect deeper.

**IMPORTANT** I often use *log1p scale*, i.e. $log1p(x)=log(1+x)$

```{r}
#| label: q-vs-rel-width-preparation

fe_widths <- q_all_methods %>%
  filter(method == "Fixed effect") %>%
  select(MA, width_fe = width)

q_relative <- q_all_methods %>%
  left_join(fe_widths, by = "MA") %>%
  mutate(
    rel_width_increase = (width - width_fe) / width_fe
  )
```

```{r}
#| label: fig-q-vs-rel-width
#| fig-height: 15
#| fig-width: 20

#add the panels
data_all <- q_relative %>% 
  mutate(panel = "All Methods (Including HK)")

data_no_hk <- q_relative %>% 
  filter(method != "Hartung & Knapp") %>% 
  mutate(panel = "Excluding Hartung & Knapp")

#same as above
combined_data <- bind_rows(data_all, data_no_hk) %>%
  mutate(panel = factor(panel, levels = c("All Methods (Including HK)", "Excluding Hartung & Knapp")))


ggplot(combined_data, aes(x = Q, y = rel_width_increase, color = method)) +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) + 
  
  facet_wrap(~panel, scales = "free_y") +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 400),
    name = "Cochran's Q (sqrt scale)"
  )  + 
  scale_y_continuous(trans = "log1p", labels = scales::percent, 
      name = "Rel. Width Increase (log1p Scale)")+
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 30),
    axis.title.x = element_text(size = 25), 
    axis.title.y = element_text(size = 25), 
    axis.text.x  = element_text(size = 20), 
    axis.text.y  = element_text(size = 20),     
    legend.position = "bottom",
    legend.title = element_text(size = 25), 
    legend.text  = element_text(size = 25)  
  )
```

```{r}
#| label: tbl-relative-width-summary


summary_tab <- q_relative %>%
 filter (method != "Fixed effect") %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean_Increase))

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Performance vs Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Increase)[Spearman]", 
    cor_pear = "Corr(Q, Increase)[Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 3
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

Edgington with a Mean % Increase of 39.7% and a Median of 11.0%, is highly sensitive to study-level conflict.
On the other hand, the weighted with $1/SE^2$ shows a near perfect Spearman correlation, but the magnitude of the actual penalty is smaller  (Mean = 15.0%) The Median increase = 0.0% indicated that it behaves like FE in at least half of all cases. 


**Note**:
Cochran’s $Q$ is fundamentally an inverse-variance weighted statistic ($w_i = 1/SE^2$). Because weighted Edgington uses this exact same weighting scheme, they are strongly correlated. 

---

Remember @tbl-pvals-differences-inspection ? Now we use again wheter the signs of the individual studies agree or not. This shows clearly that **when studies have different signs, the width will tend to increase wrt FE.**

```{r}
#| label: fig-rel-width-conflict

conflict_width_data <- q_relative %>%
  filter(method %in% c("Edgington", "Edgington (1/SE)", "Edgington (1/SE^2)")) %>%
  left_join(
    input_df %>%
      group_by(MA) %>%
      summarise(
        opposing_directions = (input_estimate[1] * input_estimate[2] < 0),
        .groups = "drop"
      ),
    by = "MA"
  ) %>%
  mutate(
    Alignment = ifelse(opposing_directions, "Opposite Sign", "Same Sign"),
    method = factor(method, levels = c("Edgington", "Edgington (1/SE)", "Edgington (1/SE^2)"))
  )

ggplot(conflict_width_data, aes(x = method, y = rel_width_increase, fill = Alignment)) +
  geom_boxplot(alpha = 0.8, outlier.size = 1, outlier.alpha = 0.4) +
  scale_fill_manual(values = c("Same Sign" = "#34495e", "Opposite Sign" = "#e74c3c")) +
  scale_y_continuous(trans = "log1p", labels = scales::percent) +
  theme_bw() +
  labs(
    y = "Relative Width Increase vs. Fixed Effect (log1p scale)",
    fill = "Study Alignment",
  ) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 16),
    axis.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 12)
  )
```

### Relative width increase stratified by measure

Now let's do the same inspection stratified by measure. 

*Note*: we don't plot ROM, RD and IDR since respectively only 1,6,8 Meta-analysis use those measures and we exclude HK method.

```{r}
#| label: fig-rel_width_stratified
#| fig-width: 16
#| fig-height: 20
#| fig-dpi: 300

#remove ROM, RD and IDR 
data_clean <- data_no_hk %>% 
  filter(measure != "ROM" & measure != "RD" & measure != "IDR")

ggplot(data_clean, aes(x = Q, y = rel_width_increase, color = method)) +
    #geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) +
    geom_point() +
    facet_wrap(~ measure, scales = "free") +
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)")+
    theme_bw() +
    theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 23),
    axis.title.x = element_text(size = 17), 
    axis.title.y = element_text(size = 17), 
    axis.text.x  = element_text(size = 12), 
    axis.text.y  = element_text(size = 12),     
    legend.position = "bottom",
    legend.title = element_text(size = 17), 
    legend.text  = element_text(size = 17) 
    )
```

```{r}
#| label: tbl-rel_width_stratified

summary_tab_strat <- q_relative %>%
  filter(method != "Fixed effect") %>%
  group_by(measure) %>%
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase),
    Median_Increase = median(rel_width_increase),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  mutate(measure = paste0(measure, " (N = ", n_ma, ")"))  %>%
  arrange(measure, desc(Mean_Increase))

summary_tab_strat %>%
  select (-n_ma) %>% 
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Sensitivity to Heterogeneity by Measure**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Incr) [Spearman]", 
    cor_pear = "Corr(Q, Incr) [Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 3
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```



### Relative width increase stratified by measure and method

Now stratify by measure and method.
```{r}
#| label: tbl-rel-width-stratified-summary


summary_skewness <- q_relative %>%
  filter(method != "Fixed effect") %>%
  group_by(measure) %>%
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase),
    Median_Increase = median(rel_width_increase),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) 

q_relative_nofe <- q_relative  %>%  filter(method != "Fixed effect")

Rel_width_increase_summary <- compute_summary(
  q_relative, 
  stats = "rel_width_increase", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)


Rel_width_increase_summary %>% 
  gt(groupname_col = "measure") %>%
  cols_label(
    method = "Method",
    mean = "Mean % Increase",
    median = "Median % Increase",
    sd = "SD % Increase",
    q25 = "25th % Increase",
    q75 = "75th % Increase",
    max = "Max % Increase",
    min = "Min % Increase"
  ) %>%
  fmt_percent(
    columns = c(mean, median, sd, q25, q75, max, min),
    decimals = 1
  ) %>%  opt_stylize(style = 6, color = "blue") %>%
  tab_options(row_group.background.color = "#E6E6E6")



```



### Plot relative width increase stratified by measure and method

Note that observation "*A20-93 \|\| Abb. 12 \|\| HbA1c*" is removed

```{r}
#| label: fig-rel-width-stratified2
#| fig-width: 16
#| fig-height: 20
#| fig-dpi: 300

data_plot <- data_clean %>% 
  filter(method != "Fixed effect") %>% 
  filter(MA != "A20-93 || Abb. 12 || HbA1c")


ggplot(data_plot, aes(x = Q, y = rel_width_increase, color = method)) +
    geom_jitter(alpha = 0.4, size = 1.5, width = 0.1) + 
    # geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) +

    facet_grid(method ~ measure, scales = "free") + 
    
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)") +
    
   #  line at 0%
      geom_hline(
    yintercept = 0,
    color = "blue",
    linewidth = 0.8,
    linetype = "dashed"
  ) +
  
    theme_bw(base_size = 25) +
    theme(
      legend.position = "none",
      strip.text = element_text(size = 10, face = "bold"),
      panel.spacing = unit(0.7, "lines"),
      panel.grid.minor = element_blank()
    )
```

```{r}
#| label: fig-rel-width-stratified2-plotly


p <- ggplot(data_plot, aes(
    x = Q, 
    y = rel_width_increase, 
    color = method,
    # w 'text' define what shows up in the box
    text = paste("MA ID:", MA, 
                 "<br>Cochran's Q:", round(Q, 2), 
                 "<br>Increase:", scales::percent(rel_width_increase))
  )) +
  geom_jitter(alpha = 0.5, size = 1.2, width = 0.1) + 
  facet_wrap(~ method + measure, scales = "free") + 
  scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "log1p",
      labels = scales::percent,
      name = "Rel. Width Increase (log1p Scale)") +
  theme_bw() +
  theme(legend.position = "none")


# 'tooltip = "text"' tells plotly to use the text we defined above
ggplotly(p, tooltip = "text", height = 1200, width = 1000)
```

{{< pagebreak >}}

## Checking Estimates

Inspecting the Estimates (the point which has the highest p-value) provided by each method

```{r}
#| label: tbl-estimates

estimates_df <- extract_ma_metric(cis, "estimates")

table_estimates <- compute_summary(estimates_df, stats = "estimates")

render_gt (table_estimates, title = "Estimates")
```

```{r}

# wide: one row per MA, one column per method
wide_estimates <- estimates_df %>%
  select(MA, method, estimates, measure) %>%
  pivot_wider(names_from = method, values_from = estimates)

# long comparison vs FE
long_vs_FE <- wide_estimates %>%
  pivot_longer(
    cols = -c(MA, `Fixed effect`, measure), #keep FE as column so that I can compare
    names_to = "method",
    values_to = "method_est"
  ) %>%
  mutate(
    fe = `Fixed effect`,
    delta = method_est - fe,
    method_bigger_FE = method_est > `Fixed effect`,
    method_smaller_FE = method_est < `Fixed effect`,
    equal = method_est == `Fixed effect`
  ) %>% 
  select (-`Fixed effect`)
```

### Agreement of sign of Estimates

Comparison of the **direction** of each method’s estimate with the **Fixed Effect** one.

For each method, we summarise:

The proportion of meta-analyses for which at least one of the FE or method estimates has no well-defined sign (\|estimate\| ≤ tol).

The proportion of meta-analyses where the method has the same sign as FE and the proportion with a flipped sign relative to FE, **among all meta-analyses**

```{r}

# Sign treshold ->  If a value is smaller than this, it is not considered as a sign. VALUE CAN BE CHANGED 
tol <- 1e-6

#Better sign function w tolerance
sign3 <- function(x, tol) {
       ifelse(abs(x) <= tol, 0L, ifelse(x > 0, 1L, -1L)) 
}

long_vs_FE <- long_vs_FE %>%
  mutate(
    sign_fe = sign3(fe, tol),
    sign_method = sign3(method_est, tol),
    sign_defined = (sign_fe != 0L) & (sign_method != 0L), #both different from 0
    sign_agree = ifelse(sign_defined, sign_fe == sign_method, NA), # same sign, if sign not defined put NA
    sign_flip = ifelse(sign_defined, sign_fe != sign_method, NA)
  )




summ_long_vs_FE <- long_vs_FE %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    prop_sign_non.defined = mean(!sign_defined, na.rm = TRUE) *100,
    prop_same_sign = sum(sign_agree, na.rm = TRUE) / n_total *100,
    prop_flip_sign = sum(sign_flip, na.rm = TRUE) / n_total * 100,
    .groups = "drop"
  )


summ_long_vs_FE %>%
  gt() %>%
  tab_header(
    title = md("**Agreement of sign with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "N studies",
    prop_sign_non.defined = "Sign not defined (%)",
    prop_same_sign = "Same sign as FE (%)",
    prop_flip_sign = "Sign flip vs FE (%)",
  ) %>%
  fmt_number(
    columns = c(prop_sign_non.defined, prop_same_sign, prop_flip_sign),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(n_total),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")

```

### Shrinkage ratio

$SR = \frac {\hat{\mu}_{\text{method}}}  {\hat{\mu}_{\text{FE}}}$

Negative values of SR indicate a **sign flip** relative to FE, while positive values indicate agreement in direction. Among positive ratios, values between 0 and 1 correspond to **shrinkage toward zero** relative to FE, and values larger than 1 indicate **more extreme estimates** than FE. Finally, value equal to 1 indicate that the two methods provide the same estimate.

For each method, we report how often estimates flip sign, how often they are shrunken toward zero, how often they are more extreme than FE.

```{r}

# ---- SHRINKAGE RATIO SR = method / FE ----

#use a small tolerance to avoid division by zero..
tol <- 1e-6


ratio_df <- long_vs_FE %>%
  mutate(
    SR = ifelse(abs(fe) < tol, NA, method_est / fe),   #  shrinkage ratio
    SR_class = case_when(
      SR < 0 ~ "Flip sign (SR<0)",
      SR >= 0 & SR < 1 ~ "Shrink toward 0 (0<=SR<1)",
      SR == 1 ~ "Equal (SR=1)",
      SR > 1 ~ "More extreme (SR>1)"
      )
  ) %>% 
  filter(is.finite(SR)) #just keep defined ratios


ratio_summary <- ratio_df %>%
  group_by(method) %>%
  summarise(
    n = n(),
    prop_flip = mean(SR < 0 ) *100,
    prop_shrink = mean(SR >= 0 & SR < 1 ) *100,
    prop_more_extreme = mean(SR > 1 ) * 100,
    prop_exactly_same = mean(SR == 1) * 100,
    .groups = "drop"
  )

ratio_summary %>% 
  gt() %>%
  tab_header(
    title = md("**Shrinkage ratio summary**"),
  ) %>%
  cols_label(
    method = "Method",
    n = "N studies",
    prop_flip = "Flipped signs (%)",
    prop_shrink = "Shrinked estimates(%)",
    prop_more_extreme = "More extreme estimates (%)",
    prop_exactly_same = "Exactly same estimates (%)"
    ) %>%
  fmt_number(
    columns = c(prop_flip, prop_shrink, prop_more_extreme, prop_exactly_same),
    decimals = 2
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 

```

```{r}
sr_global <- ratio_df %>%
  group_by(method) %>%
  summarise(
    Mean   = mean(SR),
    SD     = sd(SR),
    Median = median(SR),
    Q25    = quantile(SR, 0.25),
    Q75    = quantile(SR, 0.75),
    Min    = min(SR),
    Max    = max(SR),
    .groups = "drop"
  )


sr_global %>%
  gt() %>%
  tab_header(
    title = md("**Distribution of Shrinkage Ratios (Method / FE)**"),
    subtitle = "Values < 1 indicate shrinkage toward zero. Values < 0 indicate sign flip."
  ) %>%
  cols_label(method = "Method") %>%
  fmt_number(
    columns = c(Mean, SD, Median, Q25, Q75, Min, Max),
    decimals = 3
  ) %>%
  cols_align(align = "center", columns = -method) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")

# -------------------------------------------------------
# 3. Stratified Distribution Table (by Measure)
# -------------------------------------------------------

sr_stratified <- ratio_df %>%
  group_by(measure, method) %>%
  summarise(
    Mean   = mean(SR),
    SD     = sd(SR),
    Median = median(SR),
    Q25    = quantile(SR, 0.25),
    Q75    = quantile(SR, 0.75),
    Min    = min(SR),
    Max    = max(SR),
    .groups = "drop"
  ) %>%
  arrange(measure, method)

sr_stratified %>%
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Shrinkage Ratio Distribution by Effect Measure**")
  ) %>%
  cols_label(method = "Method") %>%
  fmt_number(
    columns = c(Mean, SD, Median, Q25, Q75, Min, Max),
    decimals = 3
  ) %>%
  cols_align(align = "center", columns = -method) %>%
  tab_options(
    row_group.background.color = "#E6E6E6",
    row_group.font.weight = "bold"
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

# Lemma 4.3.1

::: {.callout-note icon="false"}
## Lemma 4.3.1 (Two-study case)

Consider a meta-analysis with exactly two studies, with point estimates $\hat\theta_1, \hat\theta_2$ and standard errors $\sigma_1, \sigma_2 > 0$. Let the confidence level be $1-\alpha$ with $0 < \alpha < 1/4$, and define:

$$
z_{\alpha/2} := \Phi^{-1}(\alpha/2), \qquad z_{\sqrt{\alpha/4}} := \Phi^{-1}\!\left(\sqrt{\alpha/4}\right)
$$

where $\Phi^{-1}$ denotes the quantile function of the standard normal distribution. Let $\text{width}_E$ be the width of the $(1-\alpha)$-confidence interval based on the unweighted Edgington method, and $\text{width}_{FE}$ the width of the fixed-effect confidence interval.

Set $\Delta\theta := |\hat\theta_1 - \hat\theta_2|$. If at least one of the following two conditions holds, then $\text{width}_E < \text{width}_{FE}$.

**Condition 1:** $$
\frac{\max(\sigma_1,\sigma_2)}{\min(\sigma_1,\sigma_2)} \;<\; \sqrt{\left(\frac{z_{\alpha/2}}{z_{\sqrt{\alpha/4}}}\right)^2 - 1}
$$ and $$
\Delta\theta \;<\; -\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big|
$$

**Condition 2:** $$
-\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big| \;\le\; \Delta\theta \;<\; \frac{\sigma_1 \sigma_2}{\sqrt{\sigma_1^2 + \sigma_2^2}}\, \big(-2 z_{\alpha/2}\big) \;+\; (\sigma_1 + \sigma_2)\,z_{\sqrt{\alpha/4}}
$$
:::

```{r}
#| label: fig-lemma
#| fig-height: 8

alpha <- 0.05

# Check lemma conditions
cis_lemma <- check_lemma_conditions(cis, alpha = alpha)

# Extract and summarize results
lemma_detailed <- extract_lemma_results(cis_lemma)


lemma_applicable <- lemma_detailed %>%
  filter(applicable == TRUE)

# Plot results
plot_lemma_results(cis_lemma)
```

```{r}
#| label: tbl-lemma-summary

# Create simple summary table
lemma_summary <- lemma_applicable %>%
  summarise(
    "Total 2-Study MAs" = n(),
    "Condition 1 Holds" = sum(condition1_holds, na.rm = TRUE),
    "Condition 2 Holds" = sum(condition2_holds, na.rm = TRUE),
    "Both Conditions" = sum(condition1_holds & condition2_holds, na.rm = TRUE),
    "Neither Condition" = sum(!condition1_holds & !condition2_holds, na.rm = TRUE),
    "Edgington < FE" = sum(edgington_smaller, na.rm = TRUE),
    "Lemma Verified" = sum(lemma_verified == TRUE, na.rm = TRUE),
    "Lemma Failed" = sum(lemma_verified == FALSE, na.rm = TRUE)
  ) %>%
  pivot_longer(
    everything(),
    names_to = "Metric",
    values_to = "Count"
  ) %>%
  mutate(
    Percentage = ifelse(
      Metric == "Total 2-Study MAs",
      NA_real_,
      Count / first(Count) * 100
    )
  )

# Create table
lemma_summary %>%
  gt() %>%
  tab_header(
    title = md("**Lemma 4.3.1 Summary Statistics**"),
  ) %>%
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_number(
    columns = Percentage,
    decimals = 1,
    suffix = "%"
  ) %>%
  sub_missing(
    columns = Percentage,
    missing_text = "—"
  ) %>%
  cols_align(
    align = "left",
    columns = Metric
  ) %>%
  cols_align(
    align = "center",
    columns = c(Count, Percentage)
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  opt_stylize(style = 6, color = "blue")

```

{{< pagebreak >}}

```{r}
#| label: prep-p1-p2



extract_individual_p <- function(ma_obj) {
  
  # Calculate directional ONE-SIDED p-values (Alternative: "Greater")
  studies <- ma_obj$inputs
  z_scores <- studies$estimate / studies$SE
  
  # probability of getting an effect GREATER than observed.
  p_vals_studies <- 1- stats::pnorm(z_scores)
  
  p1 <- p_vals_studies[1]
  p2 <- p_vals_studies[2]
  
  # ma_obj$p_0 contains the TWO-SIDED doubled p-values 
  data.frame(
    MA = ma_obj$ma_id,
    p1 = p1,
    p2 = p2,
    method = names(ma_obj$p_0),
    p_combined = as.numeric(ma_obj$p_0),
    measure = ma_obj$measure,
    stringsAsFactors = FALSE
  )
}

# Bind
df_p <- do.call(rbind, lapply(cis, extract_individual_p))


target_methods <- c("Edgington", "Fixed effect", "Edgington (1/SE)", "Edgington (1/SE^2)")
df_plot <- df_p %>% 
  filter(method %in% target_methods)
```

# Additional stuff on two trials p values

In the paper *Beyond the two-trials rule* there is a nice plot showing the property of the Edgington combination.

Here I fix ($\alpha = 0.05$) and compute **one-sided** p-values for the individual studies under ($H_1:\theta \ge 0$). For study i:

$$
z_i=\frac{\hat\theta_i}{\sigma_i},
\qquad
p_i = 1-\Phi(z_i),
$$ Under the null, the p-values ($p_1,p_2$) are i.i.d. Uniforms (assuming independent studies). Define the Edgington statistic:

$$
S = p_1+p_2 \in [0,2].
$$

Since S is the sum of two independent Uniform((0,1)) variables, it follows the Irwin–Hall distribution with (n=2), i.e. a triangular distribution. Its CDF is

$$
F_S(s)=
\begin{cases}
\dfrac{s^2}{2}, & 0\le s\le 1,\\[6pt]
1-\dfrac{(2-s)^2}{2}, & 1<s\le 2.
\end{cases}
$$

The **two-sided** Edgington combined p-value computed in confMeta is:

$$
p_{\mathrm{Edg}}(S_{\mathrm{obs}})
=
2\min\!\Bigl(F_S(S_{\mathrm{obs}}),\,1-F_S(S_{\mathrm{obs}})\Bigr).
$$

In the special case (n=2), this simplifies to

$$
p_{\mathrm{Edg}}(S_{\mathrm{obs}})
=
\min\!\Bigl(S_{\mathrm{obs}}^2,\,(2-S_{\mathrm{obs}})^2\Bigr).
$$

To reject (H_0) at level ($\alpha$), we require ($p_{\mathrm{Edg}}(S\_{\mathrm{obs}})\le \alpha$). This yields two disjoint rejection regions:

-   **Lower (bottom-left) corner:**

    $$
    2F_S(s)\le\alpha
    \;\Longleftrightarrow\;
    2\cdot\frac{s^2}{2}\le\alpha
    \;\Longleftrightarrow\;
    s\le\sqrt{\alpha},
    $$ hence

    $$
    \mathcal R_{\mathrm{lower}}
    =
    \{(p_1,p_2): p_1+p_2 \le \sqrt{\alpha}\}.
    $$

-   **Upper (top-right) corner:**

    $$
    2(1-F_S(s))\le\alpha
    \;\Longleftrightarrow\;
    2\cdot\frac{(2-s)^2}{2}\le\alpha
    \;\Longleftrightarrow\;
    s\ge 2-\sqrt{\alpha},
    $$ hence

    $$
    \mathcal R_{\mathrm{upper}}
    =
    \{(p_1,p_2): p_1+p_2 \ge 2-\sqrt{\alpha}\}.
    $$

Basically, ($\mathcal R_{lower}$) corresponds to both studies providing strong evidence in the ($H_1:\theta \ge 0$) direction (small one-sided p-values), while ($\mathcal R_{upper}$) corresponds to both studies providing strong evidence in the opposite direction (one-sided p-values close to 1).

```{r}
#| label: fig-pvalue-final
#| fig-width: 10
#| fig-height: 8

# Define Alpha 
alpha <- 0.05

# For Edgington (n=2), the critical is sqrt(alpha)
# this comes from triangular distribution, E = p1 + p2 and to get p val we do E^2 /2
# 2 * (E^2 / 2) = 0.05  -> E = sqrt(0.05)
#Note there is also the upper corner to take into account

critical_value <- sqrt(alpha) 

# Add significance data
df_plot <- df_plot %>%
  mutate(
    is_significant = p_combined < alpha
  )

ggplot(df_plot, aes(x = p1, y = p2)) +
  geom_point(aes(color = is_significant, shape = is_significant), 
             alpha = 0.6, size = 2) +
  # Lower corner (Positive agrrement)
  # Condition: p1 + p2 < 0.2236
  geom_abline(intercept = critical_value, slope = -1, 
              linetype = "dashed", color = "black") +
  
  # Upper corner (Negative agreement)
  # Condition: (1 - p1) + (1 - p2) < 0.2236
  # this because: 2 - (p1 + p2) < 0.2236  =>  p1 + p2 > 2 - 0.2236
  geom_abline(intercept = 2 - critical_value, slope = -1, 
              linetype = "dashed", color = "black") +
  
  facet_wrap(~ method) +
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 1 (One sided)") +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 2 (One sided)") +
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey80"), name = "Significant") +
  scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1), name = "Significant") +
  coord_fixed() + 
  theme_bw() 
```

## let's map on the lower corner

This is the exact same as before, but now we map all singular study p values to two sided before plotting them, so we only need the bottom left triangle.

Even if it seems a good idea at first, there are cases in which we have $p_1 \approx 1$ and $p_2 \approx 0$ (or the inverse). Those cases get mapped in the bottom left corner, but they are not significant. That is why there are points in the Edgington graph that are inside the triangle but are not significant, so I guess the previous plot is better.

```{r}
#| label: fig-pvalue-twosdied
#| fig-width: 10
#| fig-height: 8

df_two_sided <- df_plot %>%
  mutate(
    p1_two_sided = pmin(p1, 1 - p1),
    p2_two_sided = pmin(p2, 1 - p2),
    is_significant = p_combined < 0.05
  )

s_crit <- sqrt(0.05)

ggplot(df_two_sided, aes(x = p1_two_sided, y = p2_two_sided)) +
    geom_point(aes(color = is_significant, shape = is_significant), 
             alpha = 0.6, size = 2) +
    geom_abline(intercept = s_crit, slope = -1, 
              linetype = "dashed", color = "black") +
    facet_wrap(~ method) +
  
  # Scales: Zoom in to the relevant area (0 to 0.5)
  scale_x_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 1 (two sided") +
  scale_y_continuous(limits = c(0, 1), expand = c(0, 0), name = "P-value Study 2 (two sided)") +
  
  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "grey80")) +
  scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1)) +
  coord_fixed() + 
  theme_bw() 
```

# Inspection of absolute difference of singular p values

Here, I plot the absolute difference of singular p values against the combined p value. This may be viewed as a similar analysis as relative width increase, but adding the interactive part, we can easily spot those meta analysis with very different individual p values.

```{r}


df_abs_diff <- df_p %>%
  select(MA, method, p1, p2, p_combined) %>%
  mutate(
    p_diff_abs = abs(p1 - p2),
    min_p = pmin(p1, p2),
    max_p = pmax(p1, p2)
  )

#add text for plotly 
df_abs_diff <- df_abs_diff %>% 
  mutate (
    text = paste0 ("<b>MA ID:</b> ", MA, "<br>",
      "<b>p1:</b> ", round(p1, 4), "<br>",
      "<b>p2:</b> ", round(p2, 4), "<br>",
      "<b>Combined p:</b> ", round(p_combined, 4), "<br>")
  )

p <- ggplot(df_abs_diff, aes(x = p_diff_abs, y = p_combined, color = method)) +
  geom_point(alpha = 0.3, aes(text = text)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Absolute Difference |p1 - p2|",
    y = "Combined P-value"
  ) +
  theme_bw() +
  facet_wrap(~method)

ggplotly(p, tooltip = "text")

```
