---
title: "Meta-Analysis Reports Index"
format: 
  dashboard:
    theme: cosmo
    embed-resources: true
---

```{css}
/*navbar is the principal title bar */
.navbar {
    background-color: #008bff !important;
}

/* text of navbar */
.navbar-title, .navbar-brand {
    color: white !important;
}

/* "Showing X entries" */
.dataTables_info {
    background-color: #008bff;
    color: white !important
}

/* Header of table */
table.dataTable thead th {
    background-color: #f8f9fa;
    border-bottom: 2px solid #008bff !important;
}
```

```{r}
#| label: setup
#| include: false
load_libraries <- function(...) {
  pkgs <- as.character(match.call(expand.dots = FALSE)$...)
  
  for (pkg in pkgs) {
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      
      # If not there (FALSE), install and then load it
      install.packages(pkg)
      library(pkg, character.only = TRUE, quietly = TRUE)
    }
  }
}

# Install and load required packages
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
#remotes::install_github("SaveFonta/confMeta", quiet = TRUE, upgrade = "never")

# Load all libraries
load_libraries(
  "dplyr", "DT", "tidyr")


# input data from
path <- file.path("Output", "cis.rds")
input <- readRDS(path)
df_estimates <- input$df_estimates

```

```{r}
#| label: setup-data
#| message: false
#| echo: false




ma_info <- df_estimates %>%
  select(no, identifier, sheet_name, effect.measure) %>%
  arrange(no) %>%
  distinct() %>% #fundamental, otw there are two rows each MA
  mutate(
    safe_id = gsub("[^A-Za-z0-9_-]", "_", identifier), #i checked and still unique id
  Report = paste0('<a href="Output/Reports/', safe_id, '.html" target="_blank" class="btn btn-primary btn-sm" style="text-decoration:none;">View Report</a>')
  )

# GOAL: add results of the methods to each MA
cis <- input$cis
cis_list <- lapply(cis, function(x) x[["ci"]])
cis_long <- bind_rows(cis_list, .id = "identifier")

# Wide with estimate
df_estimates_wide <- cis_long %>%
  select(identifier, method, estimate) %>%
  pivot_wider(
    names_from = method, 
    values_from = estimate,
    names_prefix = "est_"
  )

# Wide with significant
df_significant_wide <- cis_long %>%
  select(identifier, method, p_0) %>%
  pivot_wider(
    names_from = method, 
    values_from = p_0,
    names_prefix = "p0_"
  )

# containing all the singular studies
df_studies <- df_estimates %>% 
  group_by(identifier) %>% 
  summarise(
    est1 = logEst[1],
    est2 = logEst[2],
    SE1  = selogEst[1],
    SE2  = selogEst[2],

    opposing_directions = est1 * est2 < 0,   # opp signs
    abs_diff = abs(est1 - est2),
    SE_ratio = max(c(SE1, SE2)) / min(c(SE1, SE2)),

    .groups = "drop"
  )


#now let's also add the significance of Q
hetero_list <- lapply(cis, function(x) x[["heterogeneity"]])
hetero_df <- bind_rows(hetero_list, .id = "identifier") %>% 
  select(identifier, Q, I2)


# Merge (I use left join here, can also use merge)
ma_info <- ma_info %>%
  left_join(df_estimates_wide, by = "identifier") %>%
  left_join(df_significant_wide, by = "identifier") %>% 
  left_join(hetero_df, by = "identifier") %>% 
  left_join(df_studies, by = "identifier")


#convert in factor what needed
ma_info <- ma_info %>% 
  mutate(
    sheet_name = as.factor(sheet_name),
    effect.measure = as.factor (effect.measure),
    p0_Edgington = p0_Edgington, 
    opposing_directions = as.factor(opposing_directions)
         )

# 
ma_info <- ma_info  %>% 
  rowwise() %>%  #to process rowise 
  mutate(
    est_range = {
      methods_to_check <- c(`est_Edgington`, `est_Edgington (1/SE)`, 
                           `est_Edgington (1/SE^2)`, `est_Fixed effect`)
      abs(max(methods_to_check, na.rm = TRUE) - min(methods_to_check, na.rm = TRUE))
    },
    
    pval_range = {
      p_values <- c(`p0_Edgington`, `p0_Edgington (1/SE)`, 
                      `p0_Edgington (1/SE^2)`, `p0_Fixed effect`)
      max(p_values, na.rm = TRUE) - min(p_values, na.rm = TRUE)
    }
  ) %>%
  ungroup()
```


## Row {height=20%}
### Brief explanation

This dashboard lists all meta-analyses included in the project.  
You can filter each column and click **View Report** to open the full analysis.

**Variables**

- **Identifier**: unique meta-analysis ID  
- **Effect Measure**: effect size metric (e.g., OR, RR, MD, SMD)  
- **p value Range ($H_0: \theta = 0$)**: difference between the maximum and minimum p-values across the methods, computed as $\max(p) - \min(p)$  
- **Estimate Range**: absolute difference between point estimates across methods, computed as $|\max(\hat{\theta}) - \min(\hat{\theta})|$  
- **Q value**: Cochran’s heterogeneity statistic  
- **I² value**: heterogeneity percentage (higher values indicate more heterogeneity)  
- **Opposite directions**: *TRUE* if the singular study estimates have opposite signs
- **SE ratio **: ratio of the singular studies standard errors (with the convention that the larger is at the numerator) i.e. $SE_{ratio} = \frac{max(SE)}{min(SE)}$
- **Link**: opens the HTML report for the meta-analysis



## Row {height=80%}
```{r}
#| title: "Select a Meta Analysis"
#| echo: false

#check: https://rstudio.github.io/DT/#https://rstudio.github.io/DT/#

ma_info %>%
  select(
    "Identifier" = identifier, 
    "Effect Measure" = effect.measure,
    "p value Range (H<sub>0</sub>: &theta; = 0)" = pval_range,
    "Estimate Range" = est_range,
    "Q value" = Q,
    "I2 value" = I2,
    "Edgington p value" = p0_Edgington,
    "Opposite directions" = opposing_directions,
    "SE ratio" = SE_ratio,
    "Link" = Report
  ) %>%
  datatable(
    escape = FALSE, # Absolutely necessary to trasform those HTML in clickable links 
    filter = "top",   # Adds filters on top of each column
    options = list(
      pageLength = 50, #50 MA per page
      ordering = TRUE, ## allows ordering
      searchHighlight = TRUE,
      autoWidth = TRUE
    )
  ) %>%
  formatRound(columns = c("Edgington p value", "Estimate Range", "Q value", "I2 value", "p value Range (H<sub>0</sub>: &theta; = 0)", "SE ratio"), digits = 3)
```

