---
title: "Meta-Analysis Reports Index"
format: 
  dashboard:
    theme: cosmo
    embed-resources: true
---

```{css}
/*navbar is the principal title bar */
.navbar {
    background-color: #008bff !important;
}

/* text of navbar */
.navbar-title, .navbar-brand {
    color: white !important;
}

/* "Showing X entries" */
.dataTables_info {
    background-color: #008bff;
    color: white !important
}

/* Header of table */
table.dataTable thead th {
    background-color: #f8f9fa;
    border-bottom: 2px solid #008bff !important;
}
```

```{r}
#| label: setup
#| include: false

library("dplyr")
library("DT")
library("tidyr")
library("purrr")


# input data from
path <- file.path("Output", "cis.rds")
input <- readRDS(path)
cis <- input$cis
df_estimates <- input$df_estimates
```

```{r}
#| label: setup-data
#| message: false
#| echo: false




# FIRST DATAFRAME
# Contains all stats for every Method x MA combination

df_mas <- map_dfr(cis, function(x) {
  x$ci %>%
    as.data.frame() %>%
    mutate(
      MA = x$ma_id,
      measure = x$measure,
      # Add heterogeneity (repeated for all methods in this MA)
      Q = x$heterogeneity$Q,
      I2 = x$heterogeneity$I2,
      Tau2 = x$heterogeneity$Tau^2,

      data_skewness = x$data_skewness
    )
}) %>%
  rename(p_val = p_0) 

# 2ND
# Contains 1 row per MA with study-level info (SEs, estimates, bounds)

df_inputs <- map_dfr(cis, function(x) {
   est <- x$inputs$estimate
   se  <- x$inputs$SE
   low <- x$inputs$lower
   upp <- x$inputs$upper
   
   data.frame(
     MA = x$ma_id,
     est_1 = est[1], 
     est_2 = est[2],
     
     SE_1 = se[1],   
     SE_2 = se[2],

     min_study_lower = min(low),
     max_study_upper = max(upp),

     SE_ratio = max(se) / min(se),
     opposing_directions = (est[1] * est[2]) < 0
   )
})


# This is useful for comparison w/ FE baseline

df_fe <- df_mas %>%
  filter(method == "Fixed effect") %>%
  select(MA, width_fe = width, p_fe = p_val, est_fe = estimate)

df_comparisons <- df_mas %>%
  filter(method != "Fixed effect") %>%
  left_join(df_fe, by = "MA") %>%
  mutate(
    rel_width_increase = (width - width_fe) / width_fe,
    p_diff = p_val - p_fe,
    shrinkage_ratio = ifelse(abs(est_fe) < 1e-6, NA, estimate / est_fe),
    sign_flip = sign(estimate) != sign(est_fe)
  )






df_ranges <- df_mas %>%
  group_by(MA) %>%
  summarise(
    measure = measure[1],
    Q = Q[1],

    p_range = max(p_val) - min(p_val),
    est_range = abs(max(estimate) - min(estimate)),

    p_Edgington = p_val[method == "Edgington"][1]
  )

#just rel width increase for Edg
df_width_edg <- df_comparisons %>%
  filter(method == "Edgington") %>%
  select(MA, rel_width_increase)



# FINAL REQUEST:
# Calculate informativeness for every Method/MA combination
info_data_wide <- df_mas %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    # Calculating study widths from SEs (assuming 95% CI)
    width_s1 = SE_1 * 1.96 * 2,
    width_s2 = SE_2 * 1.96 * 2,
    
    # Logic for informativeness
    smaller_than_s1 = width < width_s1,
    smaller_than_s2 = width < width_s2,
    
    status = case_when(
      smaller_than_s1 & smaller_than_s2 ~ "Both",
      smaller_than_s1 | smaller_than_s2 ~ "One",
      TRUE                               ~ "None"
    )
  ) %>%
  # Pivot the table so each Method gets its own Informativeness column
  select(MA, method, status) %>%
  pivot_wider(
    names_from = method, 
    values_from = status,
    names_prefix = "Info_"
  )



# merge
final_table <- df_inputs %>%
  left_join(df_ranges, by = "MA") %>%
  left_join(df_width_edg, by = "MA") %>%
  left_join(info_data_wide, by = "MA") %>% 
  mutate(
    safe_id = gsub("[^A-Za-z0-9_-]", "_", MA),
    Report = paste0('<a href="Output/Reports/', safe_id, '.html" target="_blank" class="btn btn-primary btn-sm" style="text-decoration:none;">View Report</a>')
  )





# convert in factor what needed
final_table <- final_table %>% 
  mutate(
    measure = as.factor (measure),
    opposing_directions = as.factor(opposing_directions),
    `Info_Edgington`= factor(`Info_Edgington`, levels = c("None", "One", "Both")),
    `Info_Edgington (1/SE)` = factor(`Info_Edgington (1/SE)`, levels = c("None", "One", "Both")),
    `Info_Edgington (1/SE^2)` = factor(`Info_Edgington (1/SE^2)`, levels = c("None", "One", "Both"))
  )
```


## Row {height=20%}
### Brief explanation

This dashboard lists all meta-analyses included in the project.  
You can filter each column and click **View Report** to open the full analysis.

**Variables**

- **Identifier**: unique meta-analysis ID  
- **Effect Measure**: effect size metric (e.g., OR, RR, MD, SMD)  
- **p value Range ($H_0: \theta = 0$)**: difference between the maximum and minimum p-values across the methods, computed as $\max(p) - \min(p)$  
- **Estimate Range**: absolute difference between point estimates across methods, computed as $|\max(\hat{\theta}) - \min(\hat{\theta})|$  
- **Q value**: Cochran’s heterogeneity statistic  
- **Rel Width Incr.**: Relative increase in confidence interval width of the **Edgington** method compared to Fixed Effect ($\frac{W_{Edg} - W_{FE}}{W_{FE}}$).
- **Opposite directions**: *TRUE* if the singular study estimates have opposite signs
- **SE ratio **: ratio of the singular studies standard errors (with the convention that the larger is at the numerator) i.e. $SE_{ratio} = \frac{max(SE)}{min(SE)}$
- **Link**: opens the HTML report for the meta-analysis



## Row {height=80%}
```{r}
#| title: "Select a Meta Analysis"
#| echo: false

#check: https://rstudio.github.io/DT/#https://rstudio.github.io/DT/#

final_table %>%
  select(
    "Identifier" = MA, 
    "Effect Measure" = measure,
    "p value Range" = p_range,
    "Estimate Range" = est_range,
    "Rel Width Incr." = rel_width_increase,
    "Q value" = Q,
    "Edgington p" = p_Edgington,
    "Opposite directions" = opposing_directions,
    "Info Edgington" = `Info_Edgington`, 
    "Info Edgington (1/SE)" = `Info_Edgington (1/SE)`, 
    "Info Edgington (1/SE^2)" = `Info_Edgington (1/SE^2)`,
    "SE ratio" = SE_ratio,
    "Link" = Report
  ) %>%
  datatable(
  escape = FALSE, # Absolutely necessary to trasform those HTML in clickable links 
    filter = "top",   # Adds filters on top of each column
    options = list(
      pageLength = 50, #50 MA per page
      ordering = TRUE, ## allows ordering
      searchHighlight = TRUE,
      autoWidth = TRUE
    )
  ) %>%
  formatRound(columns = c("p value Range", "Estimate Range", "Rel Width Incr.", "Q value", "Edgington p", "SE ratio"), digits = 3) %>%
  formatPercentage(columns = "Rel Width Incr.", digits = 1)



```

