---
title: "Summary Report: IQWiG 17_02_2026"
subtitle: "Restricted to 2 studies"
author: "Saverio Fontana"
date: today
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    embed-resources: true
    grid:
      sidebar-width: 250px
      body-width: 900px
  pdf:
    documentclass: article
    geometry: [margin=1in]
    toc: true
execute:
  echo: !expr "knitr::is_html_output()"
  warning: false
  message: false
  fig-width: 10
  fig-height: 7
  fig-dpi: 300
---

```{r}
#| label: setup
#| include: false

load_libraries <- function(...) {
  pkgs <- as.character(match.call(expand.dots = FALSE)$...)
  
  for (pkg in pkgs) {
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      
      # If not there (FALSE), install and then load it
      install.packages(pkg)
      library(pkg, character.only = TRUE, quietly = TRUE)
    }
  }
}

# Install and load required packages
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
if (!requireNamespace("confMeta", quietly = TRUE)) {
  remotes::install_github("SaveFonta/confMeta", upgrade = "never")
}

# Load all libraries
load_libraries(
  "confMeta", "bayesmeta", "dplyr", "tibble", "tidyr", 
  "ggplot2", "patchwork", "gt", "parallel", 
  "ggbeeswarm", "purrr", "ggridges", "ggforce", "plotly", "irr",
  "ggh4x"
)



```

```{r}
#| label: load-data
#| include: false

source("00_utilities.R") 
   
name <- "cis.rds"
inp <- file.path("Output", name)
input <- readRDS(inp)
cis <- input$cis
df_estimates <- input$df_estimates


# FIRST DATAFRAME
# Contains all stats for every Method x MA combination

df_mas <- map_dfr(cis, function(x) {
  x$ci %>%
    as.data.frame() %>%
    mutate(
      MA = x$ma_id,
      measure = x$measure,
      # Add heterogeneity (repeated for all methods in this MA)
      Q = x$heterogeneity$Q,
      I2 = x$heterogeneity$I2,
      Tau2 = x$heterogeneity$Tau^2,

      data_skewness = x$data_skewness
    )
}) %>%
  rename(p_val = p_0) 

# 2ND
# Contains 1 row per MA with study-level info (SEs, estimates, bounds)

df_inputs <- map_dfr(cis, function(x) {
   est <- x$inputs$estimate
   se  <- x$inputs$SE
   low <- x$inputs$lower
   upp <- x$inputs$upper
   
   data.frame(
     MA = x$ma_id,
     est_1 = est[1], 
     est_2 = est[2],
     
     SE_1 = se[1],   
     SE_2 = se[2],
    
     
     low_1 = low[1],
     low_2 = low[2],
     upp_1 = upp[1],
     upp_2 = upp[2],
     min_study_lower = min(low),
     max_study_upper = max(upp),

     SE_ratio = max(se) / min(se),
     opposing_directions = (est[1] * est[2]) < 0
   )
})


# This is useful for comparison w/ FE baseline

df_fe <- df_mas %>%
  filter(method == "Fixed effect") %>%
  select(MA, width_fe = width, p_fe = p_val, est_fe = estimate)

df_comparisons <- df_mas %>%
  filter(method != "Fixed effect") %>%
  left_join(df_fe, by = "MA") %>%
  mutate(
    rel_width_increase = (width - width_fe) / width_fe,
    p_diff = p_val - p_fe,
    shrinkage_ratio = ifelse(abs(est_fe) < 1e-6, NA, estimate / est_fe),
    sign_flip = sign(estimate) != sign(est_fe)
  )


```

# Data Overview

We are inspecting a total of **`r length(unique(df_estimates$identifier))`** meta-analysis. Those are all made up of 2 studies.

::: {layout-ncol=2}

## Sheet division

Here is an overview of how many meta-analyisis are there in each Excel sheet

```{r}
#| label: tbl-Excel

a <- df_estimates %>%
   group_by(sheet_name) %>%
   summarize(n_MA = n()/2)

a %>% 
  gt() %>%
  cols_label(
    sheet_name = md("**Data Sheet Name**"),
    n_MA = md("**Number of Meta-Analyses (N)**")
  )
```

## Measure employed

This is an overview of the different measures employed over the file

```{r}
#| label: tbl-measures


df_estimates %>%
  count(effect.measure, .drop = FALSE)  %>%   
  mutate(n = n/2 ) %>% 
   arrange(desc(n)) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    effect.measure = md("**How measures are reported**"),
    n = md("**Number**")
  )

```
:::

## Data formatting

Another interesting fact was the way IQWIG reported the raw data. I classified them in 15 possible ways. This also show us which are the most common ways of reporting Meta Anslysis.

-   **2x2 OR**: Binary outcome reported as a 2×2 event table (a,b,c,d); effect computed as log odds ratio from counts.

-   **2x2 RR**: Binary outcome reported as a 2×2 event table; effect computed as log risk ratio from counts.

-   **2x2 RD**: Binary outcome reported as a 2×2 event table; effect computed as risk difference from counts.

-   **HR**: Study reports the log-Hazard Ratio with its standard error directly.

-   **direct OR**: Study reports the log-Odd Ratio with its standard error directly.

-   **direct RR**: Study reports the log-Risk Ratio with its standard error directly.

-   **direct RD**: Study reports the risk difference with its standard error directly.

-   **ROM**: Study reports the ratio of means on the log scale with its standard error directly.

-   **IDR**: Study reports the incidence density rate ratio on the log scale with its standard error directly.

-   **direct SMD**: Study reports the standardized mean difference with its standard error directly.

-   **direct MD**: Study reports the mean difference with its standard error directly.

-   **MD pooled SD**: Continuous outcome: study reports mean difference (MD), pooled SD, and group sizes (n1, n2).

-   **SMD from pooled**: Continuous outcome: study reports MD, pooled SD, and group sizes; SMD (Hedges’ g) is derived from these.

-   **singular studies MD**: Continuous outcome: study reports group means and SDs (ȳ1, s1, n1; ȳ2, s2, n2); MD is computed.

-   **singular studies SMD**: Continuous outcome: study reports SDs and group means; SMD is computed.

```{r}
#| label: tbl-formatting


df_estimates %>%
  count(data_report, .drop = FALSE)  %>%   
   arrange(desc(n)) %>% 
  mutate (n = n/2) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    data_report = md("**How data is reported**"),
    n = md("**Number**")
  )
```

## Methods Employed:

-   **Edgington**

-   **Weighted Edgington** with weights $\frac{1}{SE}$ (Inverse SE weighted)

-   **Weighted Edgington** with weights $\frac{1}{SE^2}$ (Inverse Variance weighted)

-   **Fixed Effect**. When tabular data are provided, the default is **Mantel-Haenszel** method, otherwise we fall back to Inverse-Variance method.

-   **Hartung-Knapp** (2001a/b) and variance correction is applied if HK confidence interval is narrower than CI from classic random effects model with DerSimonian-Laird estimator (IQWiG, 2022)

This summary is computed using all the MA from above.

**Scale** All ratio-based metrics (OR, RR, HR, etc.) are analyzed on the logarithmic scale. However, to remain consistent with IQWiG reporting conventions, they are labeled using their original metric names.




::: {.callout-warning collapse="true"}
**A20-93 || Abb. 12 || HbA1c** was reported with estimates:

* **PIONEER6** = $-0.660$
* **SUSTAIN6** = $0.870$

However, according to [IQWiG Project A20-93](https://www.iqwig.de/en/projects/a20-93.html), it should be:

* **PIONEER6** = $-0.660$
* **SUSTAIN6** = $-0.870$
:::


# Width Overview

 **CI width** = (upper − lower) of the 95% CI of the specific method. 





## Relative Width Increase

$$
\frac{\text{Width}_{\text{Method}} - \text{Width}_{\text{FE}}}{\text{Width}_{\text{FE}}} 
$$ {#eq-relwidth}

Basically, we are measuring **how much the interval is increasing with respect to FE.**



```{r}
#| label: tbl-relative-width-summary

summary_tab <- df_comparisons %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    SD_Increase   = sd(rel_width_increase, na.rm = TRUE),
    Q25_Increase  = quantile(rel_width_increase, 0.25, na.rm = TRUE),
    Q75_Increase  = quantile(rel_width_increase, 0.75, na.rm = TRUE),
    Min_Increase  = min(rel_width_increase, na.rm = TRUE),
    Max_Increase  = max(rel_width_increase, na.rm = TRUE),
    .groups = "drop"
  )

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Relative Width Increase distribution**")
    ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean",
    Median_Increase = "Median",
    SD_Increase = "SD",
    Q25_Increase = "Q25",
    Q75_Increase = "Q75",
    Min_Increase = "Min",
    Max_Increase = "Max"
  ) %>%
  fmt_percent(
    columns = -method, 
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```



### Relative Width Increase stratified by measure

```{r}
#| label: tbl-rel-width-stratified-summary


df_no_hk <- df_comparisons %>% 
  filter (method != "Hartung & Knapp")
  
  
  
Rel_width_increase_summary <- compute_summary(
  df_no_hk, 
  stats = "rel_width_increase", 
  grouping = c("measure", "method"), 
  count_id = "MA"
)

Rel_width_increase_summary %>%
  filter(method != "Fixed effect") %>%
  gt(groupname_col = "measure") %>%
  cols_label(
    method = "Method",
    mean = "Mean % Increase",
    median = "Median % Increase",
    sd = "SD % Increase",
    q25 = "25th % Increase",
    q75 = "75th % Increase",
    max = "Max % Increase",
    min = "Min % Increase"
  ) %>%
  fmt_percent(
    columns = c(mean, median, sd, q25, q75, max, min),
    decimals = 2
  ) %>% opt_stylize(style = 6, color = "blue") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```

**Note** the **staircase** effect in both median and mean: *Unweighted Edgington \> Edgington 1/SE \> Edgington 1/SE\^2 \> Fixed Effect* FOR EACH MEASURE.


### Relative Width Increase stratified by measure (plot)
Here we don't plot ROM since it only have 1 observation, and also HK is not reported

```{r}
#| fig-width: 9
#| fig-height: 6
#| fig-dpi: 300


plot_df <- df_no_hk %>% 
  filter (measure != "ROM")

ggplot(plot_df, aes(x = method, y = rel_width_increase, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.size = 1, outlier.alpha = 0.3, width = 0.6) +
  facet_wrap(~measure, scales = "free_y") + 
  scale_y_continuous(
      trans = "pseudo_log", 
    labels = scales::label_percent(),
    name = "Width Increase (%) (Pseudo-log scale) "
  ) +
  scale_fill_brewer(palette = "Blues") +
  theme_bw(base_size = 12) + 
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1), 
    strip.background = element_rect(fill = "aliceblue"),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    panel.spacing = unit(1.2, "lines") 
  )
```






# Heterogeneity
**Cochran's** $Q$ $$Q = \sum_{i=1}^{k} w_i (\hat{\theta}_i - \hat{\theta}_{FE})^2$$where $k$ is the number of studies (here $k=2$), and $w_i = 1/\sigma_i^2$. Under the null hypothesis of homogeneity, $Q$ follows a $\chi^2$ distribution with $k-1$ degrees of freedom.

**Inconsistency Index** ($I^2$) Describes the percentage of variation across studies that is due to heterogeneity rather than chance.$$I^2 = \max\left(0, \frac{Q - (k - 1)}{Q}\right) \times 100\%$$

**Between-study Variance** ($\tau^2$) estimated using REML

**Note** 

Under homogeneity ($\tau^2 = 0$), we have $\theta_i \sim N(\mu, v_i)$. Then $\theta_1 - \theta_2 \sim N(0, v_1 + v_2)$, so $Q = \frac{(\theta_1 - \theta_2)^2}{v_1 + v_2} \sim \chi^2_1$.
And under $H_0$ we have ${Q \sim \chi^2_1}$. 

This is a Z test... i.e. $Q = Z^2$


Critical value at $\alpha = 0.05$: $3.84$.
So you reject iff $|\theta_1 - \theta_2| > \sqrt{3.84 (v_1 + v_2)}$.

```{r}
summary_hetero <- df_mas %>% 
  summarise(
    total = n()/length(unique(df_mas$method)),
    avg_Q = mean(Q),
    avg_I2 = mean(I2),
    avg_tau = mean(Tau2),
    .groups = "drop"
  )

summary_hetero %>%
  gt() %>%
  cols_label(
    total = "Total MA",
    avg_Q = "Average Q value",
    avg_I2 = "Average I2 value",
    avg_tau = "Average tau value"
  ) %>%
  fmt_number(
    columns = c(avg_Q,avg_I2, avg_tau),
    decimals = 2
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "red")
```

## Heterogeneity across Measures

```{r}
#| label: fig-heterogeneity-stratified
#| fig-height: 6

# We take unique MA rows to avoid counting Q multiple times
q_unique <- df_mas %>% 
  distinct(MA, .keep_all = TRUE)


#use the ggforce package --> facet_zoom allow us to zoom in


ggplot(q_unique, aes(x = Q, y = measure, color = measure)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.5) + 
  geom_jitter(height = 0.1, alpha = 0.5, size = 1) +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 500), 
    name = "Cochran's Q (Scale sqrt)"
  ) +
  
  # Note: xlim in facet_zoom must use original Q values 
  facet_zoom(xlim = c(0, 20), zoom.size = 1) + 
  
  theme_bw() +
  theme(
    legend.position = "none" 
  )

```



## Relative width increase vs Q

We plot Relative width increase [@eq-relwidth] against Cochran's $Q$ to show how methods respond to heterogeneity. A positive value indicates a penalty for heterogeneity: as the study estimates diverge, the method produces a wider (more conservative) interval.

On the left, all methods are plotted together, on the right, to better inspection, the HK method is removed.

**This plot uses the LOESS smoother**. This can be misleading, it is just to detect a general trend, next plots will inspect deeper.

```{r}
#| label: fig-q-vs-rel-width
#| fig-width: 12
#| fig-height: 6
#| fig-dpi: 300

#add the panels
data_all <- df_comparisons %>% mutate(panel = "All Methods")
data_no_hk <- df_comparisons %>% 
  filter(method != "Hartung & Knapp") %>% 
  mutate(panel = "Excluding HK")

combined_data <- bind_rows(data_all, data_no_hk)


ggplot(combined_data, aes(x = Q, y = rel_width_increase, color = method)) +
  # Using loess to show the trend
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2, alpha = 0.8) + 
  facet_wrap(~panel, scales = "free_y") +
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 400),
    name = "Cochran's Q (sqrt scale)"
  ) + 
  scale_y_continuous(
      trans = "pseudo_log", 
    labels = scales::label_percent(),
    name = "Rel. Width Increase (%) (Pseudo-log scale)"
  ) +
  theme_bw(base_size = 14) + 
  theme(
    strip.background = element_rect(fill = "aliceblue"),
    strip.text = element_text(face = "bold", size = rel(1.2)), 
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box = "vertical", 
    panel.grid.minor = element_blank()
  ) 
```


```{r}
#| label: tbl-relative-width-summary-Q


summary_tab <- df_comparisons%>%
 filter (method != "Fixed effect") %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean_Increase))

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Performance vs Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Increase)[Spearman]", 
    cor_pear = "Corr(Q, Increase)[Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

Edgington seems to be highly sensitive to study-level conflict.
On the other hand, the weighted with $1/SE^2$ shows a near perfect Spearman correlation, but the magnitude of the actual penalty is smaller.


**Note**:
Cochran’s $Q$ is fundamentally an inverse-variance weighted statistic ($w_i = 1/SE^2$). Because weighted Edgington uses this exact same weighting scheme, they are strongly correlated. 



## Stratified Heterogeneity Analysis

```{r}
#| label: fig-rel-width-stratified
#| fig-width: 10
#| fig-height: 8  
#| fig-dpi: 300

#remove ROM, RD and IDR 
data_clean <- data_no_hk %>% 
  filter(measure != "ROM" & measure != "RD" & measure != "IDR") 

ggplot(data_clean, aes(x = Q, y = rel_width_increase, color = method)) +
    geom_point(alpha = 0.5, size = 1.2) + 
    facet_wrap(~ measure, scales = "free", ncol = 2) + 
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "pseudo_log", 
      labels = scales::label_percent(),
      name = "Rel. Width Increase (%) (Pseudo-log scale)"
    ) +
    theme_bw(base_size = 13) +
    theme(
      strip.background = element_rect(fill = "aliceblue"),
      strip.text = element_text(face = "bold", size = rel(1.1)),
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      panel.spacing = unit(1, "lines") 
    ) 
```

```{r}
#| label: tbl-rel_width_stratified

summary_tab_strat <- data_no_hk %>%
  group_by(measure) %>%
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase),
    Median_Increase = median(rel_width_increase),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  mutate(measure = paste0(measure, " (N = ", n_ma, ")"))  %>%
  arrange(measure, desc(Mean_Increase))

summary_tab_strat %>%
  select (-n_ma) %>% 
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Sensitivity to Heterogeneity by Measure**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Incr) [Spearman]", 
    cor_pear = "Corr(Q, Incr) [Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 2
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "red") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```


### Plot relative width increase stratified by measure and method

```{r}
#| label: fig-rel-width-stratified-grid
#| fig-width: 10
#| fig-height: 12
#| fig-dpi: 300




ggplot(data_clean, aes(x = Q, y = rel_width_increase, color = method)) +
    geom_jitter(alpha = 0.4, size = 1, width = 0.1) + 
        facet_grid(method ~ measure) + 
    
    geom_hline(yintercept = 0, color = "blue", linewidth = 0.6, linetype = "dashed") +    # reference line at 0
  
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    
    scale_y_continuous(
      trans = "pseudo_log", 
      labels = scales::label_percent(),
      name = "Rel. Width Increase (Pseudo-Log Scale)" 
    ) +
    theme_bw(base_size = 12) + 
      theme( legend.position = "none",
      strip.text = element_text(size = rel(0.8), face = "bold"),
      strip.background = element_rect(fill = "aliceblue"),
      panel.spacing = unit(0.5, "lines"),
      panel.grid.minor = element_blank(),
      axis.text = element_text(size = rel(0.7))
    )
```


For singular inspection, here is an interactive plot:

```{r}
#| label: fig-rel-width-stratified2-plotly


p <- ggplot(data_clean, aes(
    x = Q, 
    y = rel_width_increase, 
    color = method,
    # w 'text' define what shows up in the box
    text = paste("MA ID:", MA, 
                 "<br>Cochran's Q:", round(Q, 2), 
                 "<br>Increase:", scales::percent(rel_width_increase))
  )) +
  geom_jitter(alpha = 0.5, size = 1.2, width = 0.1) + 
  facet_grid(method ~ measure, scales = "free") + 
  scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(
      trans = "pseudo_log",
      labels = scales::percent,
      name = "Rel. Width Increase (Pseudo-log Scale)") +
  theme_bw() +
  theme(legend.position = "none",
    strip.text = element_text(face = "bold", size = 10), # Make headers readable
    panel.spacing = unit(1, "lines") # Add breathing room between plots
  )


# 'tooltip = "text"' tells plotly to use the text we defined above
ggplotly(p, tooltip = "text", height = 1200, width = 1000)
```



# CIs informativness

For a given meta-analysis $x$, let the 95% confidence interval for study $i$ be denoted by $[L_{i,x}, U_{i,x}]$, where $L_{i,x}$ is the lower bound and $U_{i,x}$ is the upper bound. Specifically for the two-study case, we have $L_{1,x}, U_{1,x}$ for study 1 and $L_{2,x}, U_{2,x}$ for study 2.

Define: $$
[L_{\text{min}, x}, U_{\text{max}, x}] \quad \text{s.t.} \quad L_{\text{min}, x} = \min(L_{1,x}, L_{2,x}) \quad \text{and} \quad U_{\text{max}, x} = \max(U_{1,x}, U_{2,x})
$$

Then denote the meta-analytic confidence interval by $[L_{\text{MA}, x}, U_{\text{MA}, x}]$.

We define a meta-analytic CI as **informative** if it is contained entirely within the range of the individual study limits. $$
L_{\text{MA}, x} > L_{\text{min}, x} \quad \text{and} \quad U_{\text{MA}, x} < U_{\text{max}, x}
$$

As a contrary, an interval is defined as *non-informative* if it fails to meet this criterion (i.e., it is not strictly narrower than the study range).

```{r}
#| label: informativness-preparation

info_data <- df_mas %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    informative = (lower > min_study_lower & upper < max_study_upper)
  )

summ <- info_data %>%
  group_by(method) %>%
  summarise(
    total = n(),
    n_informative = sum(informative, na.rm = TRUE),
    prop_informative = mean(informative, na.rm = TRUE),
    .groups = "drop"
  )
```

```{r}
#| label: tbl-informativness

summ %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Informativness**")
  ) %>%
  cols_label(
    method = "Method",
    total = "Total Estimates",
    n_informative = "Informative CI (n)",
    prop_informative = "Informative CI (%)"
  ) %>%
  fmt_percent(
    columns = prop_informative,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

## Informativeness 2 ($\alpha = 0.05$)

Here we check the relationship of $width_{method}$ with $width_{study1}$ $width_{study2}$ (the confidence intervals of the **singular studies**). 

The main question is: 
**Is the width of the method smaller than the widths of both singular studies?**

We classify each Meta Analysis in:

-  **Both:** $width_{method} < min(width_{study1}, width_{study2})$ (i.e. the method is narrower wrt both studies) 
-  **One** There is one study s.t. $width_{method} > (width_{study})$ (i.e. the method is narrower than one study, but wider than the other)
-  **None**  $width_{method} > (width_{study})$ (i.e. the method is wider than both studies)


```{r}
#| label: informativ-calculation

info_data <- df_mas %>%
  left_join(df_inputs, by = "MA") %>%
  mutate(
    # Widths of the individual studies
    width_s1 = upp_1 - low_1,
    width_s2 = upp_2 - low_2,
    
    # Is MA width smaller than study width?
    smaller_than_s1 = width < width_s1,
    smaller_than_s2 = width < width_s2,
    
    # Categorize
    informativ = case_when(
      smaller_than_s1 & smaller_than_s2 ~ "Both",
      smaller_than_s1 | smaller_than_s2 ~ "One",
      TRUE                              ~ "None"
    )
  ) %>%
  # create factor
  mutate(informativ = factor(informativ, levels = c("Both", "One", "None")))
```


```{r}
#| label: tbl-informativeness-new

summary_info <- info_data %>%
  group_by(method, informativ) %>%
  summarise(n = n(), 
            .groups = "drop") %>%
  group_by(method) %>%
  mutate(
    prop = n / sum(n)) %>%
  ungroup() %>%
  pivot_wider(
    names_from = informativ, 
    values_from = c(n, prop),
    values_fill = 0
  ) 

summary_info %>%
  gt() %>%
  tab_header(title = md("**CI Informativeness 2**")) %>%
  cols_label(
    method = "Method",
    n_Both = "N Both", prop_Both = "Both (%)",
    n_One = "N One", prop_One = "One (%)",
    n_None = "N None", prop_None = "None (%)"
  ) %>%
  fmt_percent(columns = starts_with("prop"), decimals = 1) %>%
  data_color(
    columns = prop_Both,       
    method = "numeric",        
    palette = "Blues"         
  ) %>% 
  opt_stylize(style = 6, color = "blue")

```
## Distribution of Q by Informativness 2 category 
```{r}
#| label: fig-Q-dist-informativeness 
#| fig-width: 10 
#| fig-height: 8 
#| fig-dpi: 300 #


toplt <- info_data %>% 
  filter(method != "Hartung & Knapp" )
  
ggplot(toplt, aes(x = informativ, y = Q, fill = informativ)) +
  geom_boxplot(alpha = 0.6, outlier.shape = NA) + 
  geom_jitter(width = 0.2, alpha = 0.3, size = 0.8) + 
  facet_wrap(~ method) + 
  scale_y_sqrt(breaks = c(0, 5, 20, 50, 100, 200)) +
  scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "Distribution of Q by Informativeness Category", 
    y = "Cochran's Q (sqrt scale)", 
    x = "Informativeness Category" 
  ) + 
  theme_bw(base_size = 12) + 
  theme(
    legend.position = "none", 
    strip.background = element_rect(fill = "aliceblue"), 
    strip.text = element_text(face = "bold") 
  )
```
## Inspection of Informativness categories


```{r}
table_3 <- info_data %>%
  select (MA, method, informativ) %>% 
 pivot_wider(
   id_cols = "MA",
   names_from = "method",
   values_from= "informativ"
 )

metrics <- info_data %>% 
  distinct(MA, Q, SE_ratio)


df_categ <- table_3 %>%
  left_join(metrics, by = "MA") %>% 
  mutate( #map to the single letter
    code_1 = case_when(`Edgington` == "Both" ~ "B", `Edgington` == "One" ~ "O", TRUE ~ "N"),
    code_2 = case_when(`Edgington (1/SE)` == "Both" ~ "B", `Edgington (1/SE)` == "One" ~ "O", TRUE ~ "N"),
    code_3 = case_when(`Edgington (1/SE^2)` == "Both" ~ "B", `Edgington (1/SE^2)` == "One" ~ "O", TRUE ~ "N"),
    
    # paste to create category 
    category = paste0(code_1, code_2, code_3)
  ) %>%
  mutate(category = factor(category, levels = c("BBB", "OBB", "OOB", "OOO", "NOO", "NNO", "NNN", "total")))

```

```{r}
#| label: tbl-categories


summ_categ <- df_categ %>% 
  group_by(category) %>% 
  summarise (
    n= n(),
    .groups = "drop"
  )

summ_categ[8, 2] <- sum(summ_categ$n) 
summ_categ[8, 1] <- as.factor("total" )

summ_categ %>% 
  gt()


```

```{r}
#| label: tbl-categories-QvsSE

tab_Q <- compute_summary(df_categ, stats=  "Q", grouping = "category")
tab_SE <- compute_summary(df_categ, stats=  "SE_ratio", grouping = "category")

tab_combined <- tab_Q %>%
  left_join(tab_SE, by = "category", suffix = c("_Q", "_SE"))

tab_combined %>%
  gt() %>%
  tab_spanner(
    label = md("**Q**"),
    columns = ends_with("_Q")
  ) %>%
  tab_spanner(
    label = md("**SE ratio**"),
    columns = ends_with("_SE")
  ) %>%
    cols_label(
    # Q cols
    mean_Q = "Mean",
    median_Q = "Median",
    sd_Q = "SD",
    q25_Q = "Q1",
    q75_Q = "Q3",
    min_Q = "Min",
    max_Q = "Max",
    
    # SE cols
    mean_SE = "Mean",
    median_SE = "Median",
    sd_SE = "SD",
    q25_SE = "Q1",
    q75_SE = "Q3",
    min_SE = "Min",
    max_SE = "Max"
  ) %>%
    fmt_number(
    columns = where(is.numeric),
    decimals = 2
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(columns = category)
  ) %>%
  tab_style(
    style = cell_borders(sides = "right", color = "gray", weight = px(2)),
    locations = cells_body(columns = max_Q) # last column of the Q group
  ) %>% 
  data_color(
    columns = c(mean_Q, median_Q),
    method = "numeric",
    palette = "Greys" 
  ) %>%
  data_color(
    columns = c(mean_SE, median_SE),
    method = "numeric",
    palette = "Greys"
  )
```

```{r}
#| label: fig-categories-QvsSE

ggplot(df_categ, aes( x = SE_ratio, y = Q)) + 
  geom_point (alpha = 0.5, size = 1) + 
  facet_wrap (~category) + 
  scale_x_continuous(
    trans = "pseudo_log",
    name = "SE ratio (Pseudo-log scale)"
  ) + 
  scale_y_continuous(
    trans = "pseudo_log",
    name = "Q (Pseudo-log scale)"
  )+ 
  theme_bw(base_size = 12) + 
  theme(
    legend.position = "none", 
    strip.background = element_rect(fill = "aliceblue"), 
    strip.text = element_text(face = "bold") 
  )



```








# Conflict in Precision

Define $$SE_{ratio} = \frac{\max(SE_{study1}, SE_{study2})}{\min(SE_{study1}, SE_{study2})}$$

In this section we want to inspect how different methods behaves as the $SE_{ratio}$ of the **singular studies** increases.

## Relative Width Increase vs SE_ratio

**Does the method widen the interval wrt FE when one study is much less precise than the other?**


```{r}
merged_comp_inp <- df_comparisons %>% 
  left_join(df_inputs, by = "MA")

corr_df <- merged_comp_inp %>%
  group_by(method) %>%
  summarise(
    Pearson = cor(rel_width_increase, SE_ratio, use = "complete.obs", method = "pearson"),
    Spearman = cor(rel_width_increase, SE_ratio, use = "complete.obs", method = "spearman")
  )

# Reshape the table
corr_vec_df <- corr_df %>%
  pivot_longer(
    cols = c(Pearson, Spearman), 
    names_to = "Metric", 
    values_to = "Correlation"
  ) %>%
  pivot_wider(
    names_from = method, 
    values_from = Correlation
  )

corr_vec_df %>%
  gt() %>%
  fmt_number(
    columns = -Metric,
    decimals = 2
  ) %>%
  tab_header(
    title = md("**Correlation between SE ratio and Relative Width Increase**"),
  ) %>% 
  cols_align(
    align = "center",
    columns = everything()
  ) %>%
  opt_stylize(style = 6, color = "red")
```


This mean that as $SE_{ratio}$ increases, **Edgington will produces wider intervals**. This because it is "forced" to incorporate the high uncertainty (large SE) of the smaller study into the pooled p-value function.

At the same time, as $SE_{ratio}$ grows, the weigthed methods center their estimate and precision more and more exclusively on the "certain" study, reducing the contribution of the high-variance study


```{r}
#| label: fig-SE_ratio 
#| fig-width: 10
#| fig-height: 7
#| fig-dpi: 300


to_plot <- merged_comp_inp %>% 
  filter(method != "Hartung & Knapp") %>% 
  filter(measure != "ROM" & measure != "RD" & measure != "IDR")

ggplot(to_plot, aes(x = SE_ratio, y = rel_width_increase, color = method)) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 1, alpha = 0.8) +
    geom_point(alpha = 0.4, size = 1) + 
    facet_wrap(~ measure, scales = "free") +
    scale_y_continuous(
      trans = "pseudo_log", 
      labels = scales::label_percent(),
      name = "Rel. Width Increase (Pseudo-Log Scale)"
    ) + 
  scale_x_continuous(
    trans = "sqrt", 
    name = "SE ratio (sqrt scale)"
  ) + 
    theme_bw(base_size = 14) + 
    theme(
      strip.background = element_rect(fill = "aliceblue"),
      strip.text = element_text(face = "bold"),
      legend.position = "bottom"
    )
```










