---
title: 'SUMMARY REPORT for new IQWiG data (restricted to 2 studies)'
author:
  - 'Saverio Fontana'
date: today
format:
  html:
      toc: true
      number-sections: true
      code-fold: true
      embed-resources: true
      smooth-scroll: true
      anchor-sections: true 
  pdf:
     toc: true
     number-sections: true
     colorlinks: true
execute:
  echo: !expr "knitr::is_html_output()"
editor: 
  markdown: 
    wrap: 72
---

```{r}
#| label: setup
#| include: false

load_libraries <- function(...) {
  pkgs <- as.character(match.call(expand.dots = FALSE)$...)
  
  for (pkg in pkgs) {
    if (!require(pkg, character.only = TRUE, quietly = TRUE)) {
      
      # If not there (FALSE), install and then load it
      install.packages(pkg)
      library(pkg, character.only = TRUE, quietly = TRUE)
    }
  }
}

# Install and load required packages
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
if (!requireNamespace("confMeta", quietly = TRUE)) {
  remotes::install_github("SaveFonta/confMeta", upgrade = "never")
}

# Load all libraries
load_libraries(
  "confMeta", "bayesmeta", "dplyr", "tibble", "tidyr", 
  "ggplot2", "patchwork", "gt", "parallel", 
  "ggbeeswarm", "purrr", "ggridges", "ggforce"
)

# Set global theme for ggplot
theme_set(theme_minimal(base_size = 11))

```

```{r}
#| label: load-data
#| include: false

source("00_utilities.R") 

name <- "cis.rds"
inp <- file.path("Output", name)


input <- readRDS(inp)
cis <- input$cis
df_estimates <- input$df_estimates

```

# Data Overview

We are inspecting a total of **`r length(unique(df_estimates$identifier))`** meta-analysis. Those are all made up of 2 studies.

## Sheet division

Here is an overview of how many meta-analyisis are there in each Excel sheet

```{r}
a <- df_estimates %>%
   group_by(sheet_name) %>%
   summarize(n_MA = n()/2)

a %>% 
  gt() %>%
  #  Relabel Columns
  cols_label(
    sheet_name = md("**Data Sheet Name**"),
    n_MA = md("**Number of Meta-Analyses (N)**")
  )
```

## Measure employed

This is an overview of the different measures employed over the file

```{r}
df_estimates %>%
  count(effect.measure, .drop = FALSE)  %>%   
  mutate(n = n/2 ) %>% 
   arrange(desc(n)) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    effect.measure = md("**How measures are reported**"),
    n = md("**Number**")
  )

```

## Data formatting

Another interesting fact was the way IQWIG reported the raw data. I classified them in 15 possible ways. This also show us which are the most common ways of reporting Meta Anslysis.

-   **2x2 OR**: Binary outcome reported as a 2×2 event table (a,b,c,d); effect computed as log odds ratio from counts.

-   **2x2 RR**: Binary outcome reported as a 2×2 event table; effect computed as log risk ratio from counts.

-   **2x2 RD**: Binary outcome reported as a 2×2 event table; effect computed as risk difference from counts.

-   **HR**: Study reports the log-Hazard Ratio with its standard error directly.

-   **direct OR**: Study reports the log-Odd Ratio with its standard error directly.

-   **direct RR**: Study reports the  log-Risk Ratio with its standard error directly.

-   **direct RD**: Study reports the risk difference with its standard error directly.

-   **ROM**: Study reports the ratio of means on the log scale with its standard error directly.

-   **IDR**: Study reports the incidence density rate ratio on the log scale with its standard error directly.

-   **direct SMD**: Study reports the standardized mean difference with its standard error directly.

-   **direct MD**: Study reports the mean difference with its standard error directly.

-   **MD pooled SD**: Continuous outcome: study reports mean difference (MD), pooled SD, and group sizes (n1, n2).

-   **SMD from pooled**: Continuous outcome: study reports MD, pooled SD, and group sizes; SMD (Hedges’ g) is derived from these.

-   **singular studies MD**: Continuous outcome: study reports group means and SDs (ȳ1, s1, n1; ȳ2, s2, n2); MD is computed.

-   **singular studies SMD**: Continuous outcome: study reports SDs and group means; SMD is computed.

```{r}
df_estimates %>%
  count(data_report, .drop = FALSE)  %>%   
   arrange(desc(n)) %>% 
  mutate (n = n/2) %>% 
   gt() %>%
  # Relabel Columns
  cols_label(
    data_report = md("**How data is reported**"),
    n = md("**Number**")
  )
```

# Summary measures

This summary is computed using all the MA from above.   

**Scale** All ratio-based metrics (OR, RR, HR, etc.) are analyzed on the logarithmic scale. However, to remain consistent with IQWiG reporting conventions, they are labeled using their original metric names.


## Width of the confidence intervals

For each method, we summarize the distribution of **CI widths** (upper − lower) across all two-study meta-analyses.

```{r}
#| label: fig-widths
#| fig-height: 8

widths <- extract_ma_metric(cis, "width")

# Calculate summary statistics
width_summary <- widths %>%
  group_by(method) %>%
  summarise(
    mean = mean(width, na.rm = TRUE),
    median = median(width, na.rm = TRUE),
    q25 = quantile(width, 0.25, na.rm = TRUE),
    q75 = quantile(width, 0.75, na.rm = TRUE),
    min = min(width, na.rm = TRUE),
    max = max(width, na.rm = TRUE),
    .groups = "drop"
  )


width_summary %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Widths**"),
  ) %>%
  cols_label(
    method = "Method",
    mean = "Mean",
    median = "Median",
    q25 = "Q25",
    q75 = "Q75",
    min = "Min",
    max = "Max"
  ) %>%
  fmt_number(
    columns = c(mean, median, q25, q75, min, max),
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 
```

{{< pagebreak >}}

### Width by effect measure 

Widths stratified by effect measure
```{r}

width_summary <- widths %>%
  group_by(measure) %>%
  # Count unique meta-analyses in each measure group
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    mean = mean(width),
    median = median(width),
    q25 = quantile(width, 0.25),
    q75 = quantile(width, 0.75),
    min = min(width),
    max = max(width),
    .groups = "drop"
  ) %>%
  mutate(measure = paste0(measure, " (N = ", n_ma, ")")) #Append the count to the measure name for the gt group header

#gt table
width_summary %>%
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Confidence Interval Widths by Effect Measure and Method**"),
  ) %>%
  cols_hide(columns = n_ma) %>% 
  cols_label(
    method = "Method",
    mean = "Mean",
    median = "Median",
    q25 = "Q25",
    q75 = "Q75",
    min = "Min",
    max = "Max"
  ) %>%
  fmt_number(
    columns = c(mean, median, q25, q75, min, max),
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

For a better visualization, the HK method is removed and ROM and RD are not plotted, since they only have respectively 1 and 6 observations.

**Note** that MD and SMD are reported on the square root scale

```{r}
#| label: fig-ci-widths_facet
#| fig-width: 20
#| fig-height: 15
#| fig-dpi: 300

widths_noHK <- widths %>% 
  filter (method != "Hartung & Knapp") %>% 
  filter (measure != "ROM" & measure != "RD")

ggplot(widths_noHK, aes(x = method, y = width, fill = method)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~ measure, scales = "free_y") +
  labs(
    title = "Distribution of Confidence Interval Widths Stratified by measure",
    x = "Method",
    y = "CI Width"
  ) +
    ggh4x::facetted_pos_scales(
    y = measure %in% c("SMD", "MD") ~ scale_y_sqrt()
  ) + 
  theme_bw() +
  theme(
    plot.title   = element_text(face = "bold", size = 20),
    strip.text   = element_text(face = "bold", size = 16),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text.x  = element_text(angle = 45, hjust = 1, size = 13),
    axis.text.y  = element_text(size = 13),
    legend.title = element_text(size = 14),
    legend.text  = element_text(size = 13),
    legend.position = "bottom",
    strip.background = element_rect(fill = "lightblue")
  ) +
  scale_fill_brewer(palette = "Set2")
```

## Confidence interval skewness

The skewness of each confidence interval is computed as:

Of course Fixed Effect mehods cannot be skewed so we don't report them. 
$$
\text{Skewness} = \frac{\text{upper} + \text{lower} - 2 \times \text{estimate}}{\text{upper} - \text{lower}}
$$

```{r}
#| label: fig-skewness
#| fig-height: 7

ci_skewness <- extract_ma_metric(cis, "ci_skewness")

ci_skewness %>%
  filter(!method %in% c("Fixed effect", "Hartung & Knapp")) %>%
  group_by(method) %>%
  summarise(
    mean = mean(ci_skewness, na.rm = TRUE),
    median = median(ci_skewness, na.rm = TRUE),
    sd = sd(ci_skewness, na.rm = TRUE),
    q25 = quantile(ci_skewness, 0.25, na.rm = TRUE),
    q75 = quantile(ci_skewness, 0.75, na.rm = TRUE),
    min = min(ci_skewness, na.rm = TRUE),
    max = max(ci_skewness, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**CI Skewness**"),
  ) %>%
  cols_label(
    method = "Method",
    mean = "Mean",
    median = "Median",
    sd = "SD",
    q25 = "Q25",
    q75 = "Q75",
    min = "Min",
    max = "Max"
  ) %>%
  fmt_number(
    columns = where(is.numeric),
    decimals = 3
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

## Summary of Confidence Interval

How many **individual** study estimates fall outside their respective meta-analytic confidence intervals?

**Definitions:**

1.  **Total Estimates:** The count of individual study estimates analyzed
2.  **Outside CI (n):** Number of estimates falling outside the CI
3.  **Outside CI (%):** Percentage of estimates falling outside the CI

```{r}
#| label: included-analysis

# Prepare data frames
ma_ci_df <- bind_rows(
  lapply(cis, function(x) {
    x$ci %>% 
      as.data.frame() %>% 
      mutate (
        estimate_ma = estimate,
        lower_MA = lower,
        upper_MA = upper,
        width_MA = width,
        significant_MA = significant,
        MA = x$ma_id
      ) %>% 
      select (c(MA, method, estimate_ma,lower_MA, upper_MA, width_MA, significant_MA))
  }) )


input_df <- bind_rows(
  lapply(cis, function(x) {
    x$inputs %>%  
      as.data.frame() %>% 
      mutate (
        MA = x$ma_id,
        study = rownames(x$inputs),
        input_estimate = estimate,
        input_SE = SE,
        input_lower = lower,
        input_upper = upper,
        input_width = width,
        input_significant = significant 
      ) %>% 
      select (-c(estimate, SE, lower, upper, width, significant))
  }))

# Merge
merged_df <- merge(input_df, ma_ci_df, by = "MA") 

merged_df <- merged_df %>%
  mutate(
    outside_CI = input_estimate < lower_MA | input_estimate > upper_MA
  )

# Calculate summary statistics
included_summary <- merged_df %>%
  group_by(method) %>%
  summarise(
    n_estimates_total = n(),
    n_estimates_outside = sum(outside_CI, na.rm = TRUE),
    prop_outside = n_estimates_outside / n_estimates_total,
    .groups = "drop"
  )
```

```{r}
#| label: tbl-included

included_summary %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Analysis**")
  ) %>%
  cols_label(
    method = "Method",
    n_estimates_total = "Total Estimates",
    n_estimates_outside = "Outside CI (n)",
    prop_outside = "Outside CI (%)"
  ) %>%
  fmt_number(
    columns = c(n_estimates_total, n_estimates_outside),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_percent(
    columns = prop_outside,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  data_color(
    columns = prop_outside,
    method = "numeric",
    palette = c("#ffffff", "#ffe5e5", "#ffcccc", "#ff9999"),
    domain = c(0, max(included_summary$prop_outside))
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}


## CIs informativness


For a given meta-analysis $x$, let the 95% confidence interval for study $i$ be denoted by $[L_{i,x}, U_{i,x}]$, where $L_{i,x}$ is the lower bound and $U_{i,x}$ is the upper bound. Specifically for the two-study case, we have $L_{1,x}, U_{1,x}$ for study 1 and $L_{2,x}, U_{2,x}$ for study 2.

Define:
$$
[L_{\text{min}, x}, U_{\text{max}, x}] \quad \text{s.t.} \quad L_{\text{min}, x} = \min(L_{1,x}, L_{2,x}) \quad \text{and} \quad U_{\text{max}, x} = \max(U_{1,x}, U_{2,x})
$$

Then denote the meta-analytic confidence interval by $[L_{\text{MA}, x}, U_{\text{MA}, x}]$.


We define a meta-analytic CI as **informative** if it is contained entirely within the range of the individual study limits.
$$
L_{\text{MA}, x} > L_{\text{min}, x} \quad \text{and} \quad U_{\text{MA}, x} < U_{\text{max}, x}
$$

As a contrary, an interval is defined as *non-informative* if it fails to meet this criterion (i.e., it is not strictly narrower than the study range).


```{r}

bounds_df <- input_df %>% 
  group_by(MA) %>% 
  summarise(
    studies_lower = min(input_lower),
    studies_upper = max(input_upper),
    .groups = "drop"
  )

# Merge
merged_df <- merge(ma_ci_df, bounds_df, by = "MA")

merged_df <- merged_df %>% 
  mutate(
    # MA CI is informative if it is NARROWER than the studies range
    informative = (lower_MA > studies_lower & upper_MA < studies_upper)
  )

summ <- merged_df %>% 
  group_by(method) %>% 
  summarise(
    total = n(),
    n_informative = sum(informative),
    prop_informative = mean(informative),
    .groups = "drop"
  )
```
```{r}
summ %>%
  gt() %>%
  tab_header(
    title = md("**Confidence Interval Informativness**")
  ) %>%
  cols_label(
    method = "Method",
    total = "Total Estimates",
    n_informative = "Informative CI (n)",
    prop_informative = "Informative CI (%)"
  ) %>%
  fmt_percent(
    columns = prop_informative,
    decimals = 2
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    cells_body(columns = method)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```


## p-values (for $H_0: \theta=0$)

Distribution of method-specific p-values for testing $H_0: \theta=0$

*Note* the dashed line marks 0.05.

```{r}
pvals <- extract_ma_metric(cis, "p_0")


pvals %>%
  ggplot(aes(x = method, y = p_0, fill = method)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0.05, linetype = "dashed", color = "red", linewidth = 0.8) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "Method",
    y = "p-value",
    title = "Distribution of p-values by method"
  ) +
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  coord_cartesian(ylim = c(0, 1))

```

```{r}
#| label: tbl-pvalue

pvals %>%
  group_by(method) %>%
  summarise(
    mean = mean(p_0, na.rm = TRUE),
    median = median(p_0, na.rm = TRUE),
    sd = sd(p_0, na.rm = TRUE),
    q25 = quantile(p_0, 0.25, na.rm = TRUE),
    q75 = quantile(p_0, 0.75, na.rm = TRUE),
    min = min(p_0, na.rm = TRUE),
    max = max(p_0, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**p-value Summary Statistics**")
  ) %>%
  cols_label(
    method = "Method",
    mean = "Mean",
    median = "Median",
    sd = "SD",
    q25 = "Q25",
    q75 = "Q75",
    min = "Min",
    max = "Max"
) %>%
  fmt_number(
    columns = c(mean, median, sd, q25, q75, min, max),
    decimals = 4
  )  %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

### Cumulative Distribution of P-values

In this graph the **ECDF** of p-values is plotted. The grey dashed line is the CDF of the Uniform distribution.

```{r}
#| label: fig-pvalue-ecdf
#| fig-height: 7

pvals %>%
  ggplot(aes(x = p_0, color = method)) +
  stat_ecdf(geom = "step", linewidth = 1, alpha = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", 
              color = "gray40", linewidth = 0.8) +
  annotate("text", x = 0.7, y = 0.3, 
           label = "Uniform distribution\n(under global null)", 
           color = "gray40", size = 3.5) +
  scale_x_continuous(breaks = seq(0, 1, 0.1)) +
  scale_y_continuous(breaks = seq(0, 1, 0.1)) +
  labs(
    x = "P-value",
    y = "Cumulative Probability",
    color = "Method"
    ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    legend.direction = "horizontal"
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1))
  
```

{{< pagebreak >}}

### P-value Comparison to Fixed Effect

Now, we will compare each method's p-values with **Fixed Effect's p-values as baseline.**

```{r}
#| label: prepare-pvalue-comparison

pvals_wide <- pvals %>%
  select(MA, method, p_0) %>%
  pivot_wider(names_from = method, values_from = p_0)

# Calculate differences from Fixed Effect
only_fe <- pvals %>% 
  filter(method == "Fixed effect") %>% 
  select (MA, p_0_fe = p_0)

pvals_diff <- pvals %>% 
  left_join(
    only_fe,
    by = "MA"
  ) %>% 
  mutate (p_diff = p_0 - p_0_fe) %>% 
  filter (method != "Fixed effect")
```

Here, each panel plots the FE p-values vs the p-value of the specific method.

*Note*: if the $p_0$ were equal, points would distribute on the diagonal.

**Each point is one meta-analysis.** Points below the diagonal indicate smaller p-values than fixed effect (less conservative), and points above indicate larger p-values (more conservative).

```{r}
#| label: fig-pvalue-vs-fe
#| fig-height: 8

pvals_diff %>% 
  ggplot (aes(x = p_0_fe, y = p_0, color = method)) +
  geom_point(alpha= 0.4, size = 1.5) +
  geom_abline (intercept = 0, slope = 1, linetype = "dashed", color = "gray50", linewidth = 0.8) + 
  facet_wrap (~method) + 
  scale_x_continuous (breaks = seq(0,1, 0.2)) + 
  scale_y_continuous(breaks = seq(0,1, 0.2)) + 
  labs(
    x = "P-value (Fixed Effect)",
    y = "P-value (Method)",
    title = "Method p-values vs Fixed Effect p-values"
  ) +
  theme(
    legend.position = "none"
  )
```

{{< pagebreak >}}

### P-value Differences from Fixed Effect

This table is similar to @tbl-pvalue but with a focus on the difference $\Delta p = p_0 (method) - p_0 (FE)$ and showing the correlation between $p_0(method)$ and $p_0 (FE)$ .

```{r}
#| label: tbl-pvalue-differences

pvals_diff %>%
  group_by(method) %>%
  summarise(
    mean_diff = mean(p_diff, na.rm = TRUE),
    median_diff = median(p_diff, na.rm = TRUE),
    sd_diff = sd(p_diff, na.rm = TRUE),
    cor_with_fe = cor(p_0, p_0_fe, use = "complete.obs"),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**P-value Differences from Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    mean_diff = "Mean Δ",
    median_diff = "Median Δ",
    sd_diff = "SD Δ",
    cor_with_fe = "Correlation"
  ) %>%
  fmt_number(
    columns = c(mean_diff, median_diff, sd_diff, cor_with_fe),
    decimals = 4
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

### Agreement in Significance (p \< 0.05)

Here we inspect the significance agreement between the values at a $\alpha = 0.05$ level

Each meta-analysis is classified into **four categories:** both significant, both non-significant, FE-only, or method-only.

```{r}
#| label: fig-significance-agreement
#| fig-height: 6





sign_agree_p <- pvals_diff %>% 
  group_by(method) %>% 
  mutate (
    Method_sign = p_0 <0.05,
    FE_sign = p_0_fe <0.05, 
    agreement = ifelse(FE_sign & Method_sign, "Both significant",
                       ifelse(!FE_sign & !Method_sign, "Both non-significant",
                              ifelse(FE_sign & !Method_sign, "FE sig, Method non-sig", 
                                     "FE non-sig, Method sig"))))
  
  
  
  
  

sign_agree_p %>% 
  count(method, agreement) %>% 
  ggplot(aes(x = method, y = n, fill = agreement)) + 
           geom_col (position= "stack", alpha = 0.8) + 
  geom_text(aes(label = n), position = position_stack(vjust = 0.5), size = 3,
            color = "white", fontface = "bold") + 
  scale_fill_manual ( 
    values = c(      "Both significant" = "#2ca02c",
                     "Both non-significant" = "#d62728",
                     "FE sig, Method non-sig" = "purple",
                     "FE non-sig, Method sig" = "#7f7f7f"
    )) + 
  labs (
    x = "Methods", 
    y = "Count",
    fill = "Agreement", 
    title = "Significance agreement with Fixed Effect (α = 0.05)"
  ) + 
  theme_bw() + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "right"
  )
  

         
```

```{r}
#| label: tbl-significance-agreement

sign_agree_p %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    both_sig = sum(FE_sign & Method_sign, na.rm = TRUE),
    both_nonsig = sum(!FE_sign & !Method_sign, na.rm = TRUE),
    fe_only = sum(FE_sign & !Method_sign, na.rm = TRUE),
    method_only = sum(!FE_sign & Method_sign, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  gt() %>%
  tab_header(
    title = md("**Significance Agreement with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "Total",
    both_sig = "Both Sig",
    both_nonsig = "Both Non-sig",
    fe_only = "FE Only",
    method_only = "Method Only"
  ) %>%
  fmt_number(
    columns = c(n_total, both_sig, both_nonsig, fe_only, method_only),
    decimals = 0
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  )  %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

## Coherence of skewness signs (β, γ, AUCC ratio)

Agreement between different measures of skewness:

-   **β (beta)**: CI skewness
-   **γ (gamma)**: Data skewness\
-   **AUCC ratio**: Area Under Confidence Curve ratio

where $$\gamma = \frac{\sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^3}{\left( \sum_{i=1}^{n} w_i (\hat{\theta} - \bar{\theta})^2 \right)^{3/2} / \sqrt{\sum_{i=1}^{n} w_i}}$$ is Fisher’s weighted skewness coefficient.

And $$
\text{AUCC Ratio} = \frac {\text{AUCC}_{\text{upper}} - \text{AUCC}_{\text{lower}}} {AUCC}
$$.

*For more details see: Held, Leonhard, Felix Hofmann, and Samuel Pawel. "A comparison of combined p-value functions for meta-analysis." Research Synthesis Methods (2025)*

```{r}
#| label: sign-agreement-data

# Build sign agreement dataset
sign_agreement <- do.call(rbind, lapply(cis, function(x) {
  df <- x$ci 
  df <- subset(df, !is.na(ci_skewness) & !method %in% c("Fixed effect", "Hartung & Knapp", "Bayesmeta"))
  data.frame(
    MA = x$ma_id,
    method = df$method,
    data_skewness = x$data_skewness,
    ci_skewness = df$ci_skewness,
    aucc_ratio = df$aucc_ratio
  )
})) %>%
  mutate(
    same_sign_beta_gamma = sign(data_skewness) == sign(ci_skewness),
    same_sign_beta_aucc = sign(ci_skewness) == sign(aucc_ratio),
    same_sign_gamma_aucc = sign(data_skewness) == sign(aucc_ratio)
  )

```

```{r}
#| label: tbl-sign-coherence

# Calculate summary statistics
sign_summary <- sign_agreement %>%
  group_by(method) %>%
  summarise(
    n_MA = n(),
    agree_beta_gamma = sum(same_sign_beta_gamma, na.rm = TRUE),
    agree_beta_aucc = sum(same_sign_beta_aucc, na.rm = TRUE),
    agree_gamma_aucc = sum(same_sign_gamma_aucc, na.rm = TRUE),
    corr_beta_gamma = cor(ci_skewness, data_skewness, use = "complete.obs"),
    corr_beta_aucc = cor(ci_skewness, aucc_ratio, use = "complete.obs"),
    corr_gamma_aucc = cor(data_skewness, aucc_ratio, use = "complete.obs"),
    .groups = "drop"
  )

# Create gt table
sign_summary %>%
  gt() %>%
  tab_header(
    title = md("**Sign Agreement**"),
  ) %>%
  tab_spanner(
    label = md("**Sign Agreement (Count)**"),
    columns = c(agree_beta_gamma, agree_beta_aucc, agree_gamma_aucc)
  ) %>%
  tab_spanner(
    label = md("**Pearson Correlation**"),
    columns = c(corr_beta_gamma, corr_beta_aucc, corr_gamma_aucc)
  ) %>%
  cols_label(
    method = "Method",
    n_MA = md("*MA*"),
    agree_beta_gamma = "β vs γ",
    agree_beta_aucc = "β vs AUCC",
    agree_gamma_aucc = "γ vs AUCC",
    corr_beta_gamma = "β vs γ",
    corr_beta_aucc = "β vs AUCC",
    corr_gamma_aucc = "γ vs AUCC"
  ) %>%
  fmt_number(
    columns = starts_with("corr"),
    decimals = 3
  ) %>%
  fmt_number(
    columns = c(n_MA, starts_with("agree")),
    decimals = 0
  ) %>%
  sub_missing(
    columns = everything(),
    missing_text = "–"
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

{{< pagebreak >}}

## Sign agreement among β, γ, and AUCC ratio

```{r}
#| label: fig-sign-agreement-bar
#| fig-height: 6

sign_long <- sign_agreement %>%
  filter(!method %in% c("Fixed effect", "Hartung & Knapp")) %>%
  select(MA, method, starts_with("same_sign")) %>%
  pivot_longer(
    cols = starts_with("same_sign"),
    names_to = "comparison",
    values_to = "agreement"
  ) %>%
  filter(!is.na(agreement))
  
# Create bar plot showing proportions
sign_long %>%
  mutate(
    comparison = case_when(
      comparison == "same_sign_beta_gamma" ~ "β vs γ",
      comparison == "same_sign_beta_aucc" ~ "β vs AUCC",
      comparison == "same_sign_gamma_aucc" ~ "γ vs AUCC"
    )
  ) %>%
  group_by(method, comparison) %>%
  summarise(
    prop_agree = mean(agreement, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = method, y = prop_agree, fill = comparison)) +
  geom_col(position = position_dodge(0.8), width = 0.7, alpha = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey30") +
  scale_y_continuous(
    labels = scales::percent_format(),
    limits = c(0, 1),
    breaks = seq(0, 1, 0.1)
  ) +
  scale_fill_manual(
    values = c("#1f77b4", "#ff7f0e", "#2ca02c")
  ) +
  labs(
    x = "Method",
    y = "Proportion of Agreement",
    fill = "Comparison"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold", size = 12),
    legend.position = "bottom"
  )
```

{{< pagebreak >}}

## Heterogeneity summary
**Cochran's** $Q$
$$Q = \sum_{i=1}^{k} w_i (\hat{\theta}_i - \hat{\theta}_{FE})^2$$where $k$ is the number of studies (here $k=2$), and $w_i = 1/\sigma_i^2$. Under the null hypothesis of homogeneity, $Q$ follows a $\chi^2$ distribution with $k-1$ degrees of freedom.

**Inconsistency Index** ($I^2$)
Describes the percentage of variation across studies that is due to heterogeneity rather than chance.$$I^2 = \max\left(0, \frac{Q - (k - 1)}{Q}\right) \times 100\%$$

**Between-study Variance** ($\tau^2$) estimated using REML
```{r}


hetero_df <- bind_rows(
  lapply(cis, function(x) {
    x$heterogeneity %>%  
      as.data.frame() %>% 
      mutate (
        MA = x$ma_id,
        Tau2 = Tau^2
      ) }
  )
) 

summary_hetero <- hetero_df %>% 
  summarise(
    total = n(),
    avg_Q = mean(Q),
    avg_pQ = mean(p_Q),
    avg_I2 = mean(I2),
    avg_tau = mean(Tau2),
    .groups = "drop"
  )

summary_hetero %>%
  gt() %>%
  cols_label(
    total = "Total MA",
    avg_Q = "Average Q value",
    avg_pQ = "Average value of  p value of Q",
    avg_I2 = "Average I2 value",
    avg_tau = "Average tau value"
  ) %>%
  fmt_number(
    columns = c(avg_Q,avg_pQ,  avg_I2, avg_tau),
    decimals = 2
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")
```



## Stratified Q statistics

```{r}
#| label: prep-q-all-methods

# Create a long dataframe with one row per Method per MA
q_all_methods <- bind_rows(lapply(cis, function(x) {
  
  # 1. Extract heterogeneity (Common to all methods in this MA)
  Q_val <- x$heterogeneity$Q             # -->   would it be better to use I2???
  
  # 2. Extract Method-specific metrics
  methods <- names(x$width)
  
  data.frame(
    MA = x$ma_id,
    measure = x$measure,
    Q = Q_val,
    method = methods,
    width = as.numeric(x$width),
    p_value = as.numeric(x$p_0), # P-value for H0: theta=0
    stringsAsFactors = FALSE
  )
}))

```


### Q stratified by measure
Full graph on the above, with a zoomed-in view below due to the outliers that ruin the visualisation. 
Data are reported on the square root scale.
```{r}
#| label: fig-q-dist-measure
#| fig-height: 6

# We take unique MA rows to avoid counting Q multiple times
q_unique <- q_all_methods %>% 
  distinct(MA, .keep_all = TRUE)


#use the ggforce package --> facet_zoom allow us to zoom in


ggplot(q_unique, aes(x = Q, y = measure, color = measure)) +
  geom_boxplot(outlier.shape = NA, width = 0.5, alpha = 0.5) + 
  geom_jitter(height = 0.1, alpha = 0.5, size = 1) +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 500), 
    name = "Cochran's Q (Scale sqrt)"
  ) +
  
  # Note: xlim in facet_zoom must use original Q values 
  facet_zoom(xlim = c(0, 20), zoom.size = 1) + 
  
  theme_bw() +
  theme(
    legend.position = "none" 
  )

```

**IMPORTANT NOTE**, there is a huge outlier, with id "*A20-93 || Abb. 12 || HbA1c*". It has a Q-statistics = 570.95.
This value prevented the graphs do display correctly and detect pattern, so for the next two plots where Q is involved, I decided to remove it.

### Q vs CI Width

On the right, all methods are plotted together, on the right, to better inspection, the HK method is removed. 

```{r}
#| label: fig-q-vs-width-combined
#| fig-height: 15
#| fig-width: 20

q_all_methods_no_out <- q_all_methods %>% 
    filter (MA != "A20-93 || Abb. 12 || HbA1c")


data_with_HK <- q_all_methods_no_out %>% 
  mutate(panel = "All Methods (Including HK)")

data_no_HK <- q_all_methods_no_out %>% 
  filter(method != "Hartung & Knapp") %>% 
  mutate(panel = "Excluding Hartung & Knapp")

#merge
plot_data_combined <- bind_rows(data_with_HK, data_no_HK) %>% 
  mutate(panel = factor(panel, levels = c("All Methods (Including HK)", "Excluding Hartung & Knapp")))

ggplot(plot_data_combined, aes(x = Q, y = width, color = method)) +
  geom_smooth(formula = y ~ x,  method = "loess", se = FALSE) +
  facet_wrap(~panel, scales = "free_y") + #important free_y
  labs(
    x = "Cochran's Q",
    y = "CI Width",
    color = "Method"
  ) +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 30),
    axis.title.x = element_text(size = 25), 
    axis.title.y = element_text(size = 25), 
    axis.text.x  = element_text(size = 20), 
    axis.text.y  = element_text(size = 20),     
    legend.position = "bottom",
    legend.title = element_text(size = 25), 
    legend.text  = element_text(size = 25)  
    )

```

### Summary table
To compute the correlations, I am also using the MA with extreme Q-value for completeness 

```{r}
#| label: tbl-q-correlations

q_all_methods %>%
  group_by(method) %>%
  summarise(
    Cor_Q_Width = cor(Q, width, method = "spearman", use = "complete.obs"),
    Cor_Q_Width_p = cor(Q, width, method = "pearson", use = "complete.obs"),
    Cor_Q_Pval  = cor(Q, p_value, method = "spearman", use = "complete.obs"),
    Cor_Q_Pval_p  = cor(Q, p_value, method = "pearson", use = "complete.obs"),
    .groups = "drop"
  ) %>%
  arrange(desc(Cor_Q_Width)) %>%
  gt() %>%
  tab_header(
    title = md("**Correlation with Heterogeneity**"),
  ) %>%
  cols_label(
    method = "Method",
    Cor_Q_Width = "Corr(Q, Width) Spearman",
    Cor_Q_Width_p = "Corr(Q, Width) Pearson", 
    Cor_Q_Pval = "Corr(Q, P-val) Spearman", 
    Cor_Q_Pval_p  = "Corr(Q, P-val) Pearson"
  ) %>%
  fmt_number(columns = starts_with("Cor"), decimals = 3) %>%
  data_color(
    columns = starts_with("Cor"),
    method = "numeric",
    palette = "RdBu", # Red-Blue diverging palette
    domain = c(-1, 1)
  ) %>%
  opt_stylize(style = 6, color = "blue")
```

## Relative Width Increase

$$
\frac{\text{Width}_{\text{Method}} - \text{Width}_{\text{FE}}}{\text{Width}_{\text{FE}}} 
$$

Basically, we are measuring how much the interval is increasing with respect to FE. 


We plot this increase against Cochran's $Q$ to show how methods respond to heterogeneity. A positive value indicates a penalty for heterogeneity: as the study estimates diverge, the method produces a wider (more conservative) interval.




```{r}
fe_widths <- q_all_methods %>%
  filter(method == "Fixed effect") %>%
  select(MA, width_fe = width)

q_relative <- q_all_methods %>%
  left_join(fe_widths, by = "MA") %>%
  mutate(
    rel_width_increase = (width - width_fe) / width_fe
  )
```


```{r}
#| label: fig-q-vs-rel-width-combined
#| fig-height: 15
#| fig-width: 20

#add the panels
data_all <- q_relative %>% 
  mutate(panel = "All Methods (Including HK)")

data_no_hk <- q_relative %>% 
  filter(method != "Hartung & Knapp") %>% 
  mutate(panel = "Excluding Hartung & Knapp")

#same as above
combined_data <- bind_rows(data_all, data_no_hk) %>%
  mutate(panel = factor(panel, levels = c("All Methods (Including HK)", "Excluding Hartung & Knapp")))

ggplot(combined_data, aes(x = Q, y = rel_width_increase, color = method)) +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) + 
  
  facet_wrap(~panel, scales = "free_y") +
  
  scale_x_sqrt(
    breaks = c(0, 10, 50, 100, 200, 400),
    name = "Cochran's Q (sqrt scale)"
  ) +
  scale_y_continuous(labels = scales::percent, name = "Relative Width Increase vs FE") +
  theme_bw() +
  theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 30),
    axis.title.x = element_text(size = 25), 
    axis.title.y = element_text(size = 25), 
    axis.text.x  = element_text(size = 20), 
    axis.text.y  = element_text(size = 20),     
    legend.position = "bottom",
    legend.title = element_text(size = 25), 
    legend.text  = element_text(size = 25)  
  )
```


```{r}
#| label: tbl-relative-width-summary


summary_tab <- q_relative %>%
 filter (method != "Fixed effect") %>% 
  group_by(method) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase, na.rm = TRUE),
    Median_Increase = median(rel_width_increase, na.rm = TRUE),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  arrange(desc(Mean_Increase))

summary_tab %>%
  gt() %>%
  tab_header(
    title = md("**Performance vs Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Increase)[Spearman]", 
    cor_pear = "Corr(Q, Increase)[Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 3
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue")
```


### Relative width increase stratified by measure
Now let's do the same inspection stratified by measure. 
*Note*: we don't plot ROM, RD and IDR since respectively only 1,6,8 Meta-analysis use those measures. 

Also here we exclude the HK method. 
```{r}

#remove ROM, RD and IDR 
data_clean <- data_no_hk %>% 
  filter(measure != "ROM" & measure != "RD" & measure != "IDR")

ggplot(data_clean, aes(x = Q, y = rel_width_increase, color = method)) +
    geom_smooth(formula = y ~ x, method = "loess", se = FALSE, linewidth = 1.2) +
    facet_wrap(~ measure, scales = "free") +
    scale_x_sqrt(name = "Cochran's Q (sqrt scale)") +
    scale_y_continuous(labels = scales::percent, name = "Relative Width Increase vs FE") +
    labs(title = "Sensitivity to Heterogeneity by Measure") +
    theme_bw() +
    theme(
    strip.background = element_rect(fill = "lightblue"),
    strip.text = element_text(face = "bold", size = 10),
    axis.title.x = element_text(size = 10), 
    axis.title.y = element_text(size = 10), 
    axis.text.x  = element_text(size = 5), 
    axis.text.y  = element_text(size = 5),     
    legend.position = "bottom",
    legend.title = element_text(size = 10), 
    legend.text  = element_text(size = 10) 
    )
```

```{r}
summary_tab_strat <- q_relative %>%
  filter(method != "Fixed effect") %>%
  group_by(measure) %>%
  mutate(n_ma = n_distinct(MA)) %>% 
  group_by(measure, method, n_ma) %>%
  summarise(
    Mean_Increase = mean(rel_width_increase),
    Median_Increase = median(rel_width_increase),
    cor_spear = cor(Q, rel_width_increase, method = "spearman"),
    cor_pear = cor(Q, rel_width_increase, method = "pearson"),
    .groups = "drop"
  ) %>%
  mutate(measure = paste0(measure, " (N = ", n_ma, ")"))  %>%
  arrange(measure, desc(Mean_Increase))

summary_tab_strat %>%
  gt(groupname_col = "measure") %>%
  tab_header(
    title = md("**Sensitivity to Heterogeneity by Measure**")
  ) %>%
  cols_label(
    method = "Method",
    Mean_Increase = "Mean % Increase",
    Median_Increase = "Median % Increase",
    cor_spear = "Corr(Q, Incr) [Spearman]", 
    cor_pear = "Corr(Q, Incr) [Pearson]"
  ) %>%
  fmt_percent(
    columns = c(Mean_Increase, Median_Increase),
    decimals = 1
  ) %>%
  fmt_number(
    columns = c(cor_spear, cor_pear),
    decimals = 3
  ) %>%
  data_color(
    columns = Mean_Increase,
    method = "numeric",
    palette = "Reds" 
  ) %>%
  data_color(
    columns = c(cor_spear, cor_pear),
    method = "numeric",
    palette = "Blues",
    domain = c(0, 1) 
  ) %>%
  opt_stylize(style = 6, color = "blue") %>%
  tab_options(row_group.background.color = "#E6E6E6")
```


## Checking Estimates

Inspecting the Estimates (the point which has the highest p-value) provided by each method

```{r}

estimates_df <- extract_ma_metric(cis, "estimates")

# Calculate summary statistics
estimate_summary <- estimates_df %>%
  group_by(method) %>%
  summarise(
    mean = mean(estimates, na.rm = TRUE),
    sd = sd(estimates, na.rm = TRUE),
    median = median(estimates, na.rm = TRUE),
    q25 = quantile(estimates, 0.25, na.rm = TRUE),
    q75 = quantile(estimates, 0.75, na.rm = TRUE),
    min = min(estimates, na.rm = TRUE),
    max = max(estimates, na.rm = TRUE),
    .groups = "drop"
  )


estimate_summary %>%
  gt() %>%
  tab_header(
    title = md("**Meta Analytic estimates**"),
  ) %>%
  cols_label(
    method = "Method",
    mean = "Mean",
    sd = "SD",
    median = "Median",
    q25 = "Q25",
    q75 = "Q75",
    min = "Min",
    max = "Max"
  ) %>%
  fmt_number(
    columns = c(sd, median, q25, q75, min, max),
    decimals = 2
  ) %>%
    fmt_number(
    columns = c(mean),
    decimals = 4 
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 
```

Here, delta(Δ) = designated method - Fixed effect

```{r}

# wide: one row per MA, one column per method
wide_estimates <- estimates_df %>%
  select(MA, method, estimates) %>%
  pivot_wider(names_from = method, values_from = estimates)

# long comparison vs FE
long_vs_FE <- wide_estimates %>%
  pivot_longer(
    cols = -c(MA, `Fixed effect`),#keep FE as column so that I can compare
    names_to = "method",
    values_to = "method_est"
  ) %>%
  mutate(
    fe = `Fixed effect`,
    delta = method_est - fe,
    method_bigger_FE = method_est > `Fixed effect`,
    method_smaller_FE = method_est < `Fixed effect`,
    equal = method_est == `Fixed effect`
  ) %>% 
  select (-`Fixed effect`)



# summary table
summary_vs_FE <- long_vs_FE %>%
  group_by(method) %>%
  summarise(
    n = n(),
    pct_method_bigger_FE = mean(method_bigger_FE) * 100,
    pct_method_smaller_FE = mean(method_smaller_FE) * 100,
    pct_equal = mean(equal) * 100,
    mean_delta = mean(delta),
    mean_abs_delta = mean(abs(delta)),
    .groups = "drop"
  )


summary_vs_FE %>% 
  gt() %>%
  tab_header(
    title = md("**Meta Analytic delta-estimates**"),
  ) %>%
  cols_label(
    method = "Method",
    n = "N studies",
    pct_method_bigger_FE = "Bigger than FE (%)",
    pct_method_smaller_FE = "Smaller than FE (%)",
    pct_equal = "Equal (%)",
    mean_delta = "Mean Δ",
    mean_abs_delta = "Mean Abs Δ"
  ) %>%
  fmt_number(
    columns = c(pct_method_bigger_FE, pct_method_smaller_FE),
    decimals = 2
  ) %>%
    fmt_number(
    columns = c(mean_delta, mean_abs_delta),
    decimals = 4 
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 

```

```{r}

#COMMENTED OUT:
# Not sure if we want to inspect those studies where we have the same meta analytic estimate

#same_estimate <- wide_estimates %>% 
#  filter (Edgington == `Fixed effect`)

#list <- same_estimate$MA


#un <- unlist(list)
#same_est <- cis[un]

#extract only inputs from each elemnt of the list
#lapply(same_est, "[[" , "inputs")

```

### Agreement of sign of Estimates

Comparison of the **direction** of each method’s estimate with the **Fixed Effect** one.

For each method, we summarise:

The proportion of meta-analyses for which at least one of the FE or method estimates has no well-defined sign (\|estimate\| ≤ tol).

The proportion of meta-analyses where the method has the same sign as FE and the proportion with a flipped sign relative to FE, **among all meta-analyses**

```{r}

# Sign treshold ->  If a value is smaller than this, it is not considered as a sign. VALUE CAN BE CHANGED 
tol <- 1e-6

#Better sign function w tolerance
sign3 <- function(x, tol) {
       ifelse(abs(x) <= tol, 0L, ifelse(x > 0, 1L, -1L)) 
}

long_vs_FE <- long_vs_FE %>%
  mutate(
    sign_fe = sign3(fe, tol),
    sign_method = sign3(method_est, tol),
    sign_defined = (sign_fe != 0L) & (sign_method != 0L), #both different from 0
    sign_agree = ifelse(sign_defined, sign_fe == sign_method, NA), # same sign, if sign not defined put NA
    sign_flip = ifelse(sign_defined, sign_fe != sign_method, NA)
  )




summ_long_vs_FE <- long_vs_FE %>%
  group_by(method) %>%
  summarise(
    n_total = n(),
    prop_sign_non.defined = mean(!sign_defined, na.rm = TRUE) *100,
    prop_same_sign = sum(sign_agree, na.rm = TRUE) / n_total *100,
    prop_flip_sign = sum(sign_flip, na.rm = TRUE) / n_total * 100,
    .groups = "drop"
  )


summ_long_vs_FE %>%
  gt() %>%
  tab_header(
    title = md("**Agreement of sign with Fixed Effect**")
  ) %>%
  cols_label(
    method = "Method",
    n_total = "N studies",
    prop_sign_non.defined = "Sign not defined (%)",
    prop_same_sign = "Same sign as FE (%)",
    prop_flip_sign = "Sign flip vs FE (%)",
  ) %>%
  fmt_number(
    columns = c(prop_sign_non.defined, prop_same_sign, prop_flip_sign),
    decimals = 2
  ) %>%
  fmt_number(
    columns = c(n_total),
    decimals = 0,
    use_seps = TRUE
  ) %>%
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue")


```

### Shrinkage ratio

$SR = \frac {\hat{\mu}_{\text{method}}}  {\hat{\mu}_{\text{FE}}}$

Negative values of SR indicate a **sign flip** relative to FE, while positive values indicate agreement in direction. Among positive ratios, values between 0 and 1 correspond to **shrinkage toward zero** relative to FE, and values larger than 1 indicate **more extreme estimates** than FE. 
Finally, value equal to 1 indicate that the two methods provide the same estimate.

For each method, we report how often estimates flip sign, how often they are shrunken toward zero, how often they are more extreme than FE.

```{r}

# ---- SHRINKAGE RATIO SR = method / FE ----

ratio_df <- long_vs_FE %>%
  mutate(
    SR = method_est / fe,                     #  shrinkage ratio
    SR_class = case_when(
      SR < 0 ~ "Flip sign (SR<0)",
      SR >= 0 & SR < 1 ~ "Shrink toward 0 (0<=SR<1)",
      SR == 1 ~ "Equal (SR=1)",
      SR > 1 ~ "More extreme (SR>1)"
      )
  )

ratio_summary <- ratio_df %>%
  group_by(method) %>%
  summarise(
    n = n(),
    prop_flip = mean(SR < 0, na.rm = TRUE) *100,
    prop_shrink = mean(SR >= 0 & SR < 1, na.rm = TRUE) *100,
    prop_more_extreme = mean(SR > 1, na.rm = TRUE) * 100,
    prop_exactly_same = mean(SR == 1, na.rm = TRUE) * 100,
    .groups = "drop"
  )

ratio_summary %>% 
  gt() %>%
  tab_header(
    title = md("**Shrinkage ratio summary**"),
  ) %>%
  cols_label(
    method = "Method",
    n = "N studies",
    prop_flip = "Flipped signs (%)",
    prop_shrink = "Shrinked estimates(%)",
    prop_more_extreme = "More extreme estimates (%)",
    prop_exactly_same = "Exactly same estimates (%)"
    ) %>%
  fmt_number(
    columns = c(prop_flip, prop_shrink, prop_more_extreme, prop_exactly_same),
    decimals = 2
  ) %>% 
  cols_align(
    align = "center",
    columns = -method
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) %>%
  opt_stylize(style = 6, color = "blue") 

```

# Lemma 4.3.1

::: {.callout-note icon="false"}
## Lemma 4.3.1 (Two-study case)

Consider a meta-analysis with exactly two studies, with point estimates $\hat\theta_1, \hat\theta_2$ and standard errors $\sigma_1, \sigma_2 > 0$. Let the confidence level be $1-\alpha$ with $0 < \alpha < 1/4$, and define:

$$
z_{\alpha/2} := \Phi^{-1}(\alpha/2), \qquad z_{\sqrt{\alpha/4}} := \Phi^{-1}\!\left(\sqrt{\alpha/4}\right)
$$

where $\Phi^{-1}$ denotes the quantile function of the standard normal distribution. Let $\text{width}_E$ be the width of the $(1-\alpha)$-confidence interval based on the unweighted Edgington method, and $\text{width}_{FE}$ the width of the fixed-effect confidence interval.

Set $\Delta\theta := |\hat\theta_1 - \hat\theta_2|$. If at least one of the following two conditions holds, then $\text{width}_E < \text{width}_{FE}$.

**Condition 1:** $$
\frac{\max(\sigma_1,\sigma_2)}{\min(\sigma_1,\sigma_2)} \;<\; \sqrt{\left(\frac{z_{\alpha/2}}{z_{\sqrt{\alpha/4}}}\right)^2 - 1}
$$ and $$
\Delta\theta \;<\; -\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big|
$$

**Condition 2:** $$
-\,z_{\sqrt{\alpha/4}}\,\big|\sigma_2 - \sigma_1\big| \;\le\; \Delta\theta \;<\; \frac{\sigma_1 \sigma_2}{\sqrt{\sigma_1^2 + \sigma_2^2}}\, \big(-2 z_{\alpha/2}\big) \;+\; (\sigma_1 + \sigma_2)\,z_{\sqrt{\alpha/4}}
$$
:::

```{r}
#| label: fig-lemma
#| fig-height: 8

alpha <- 0.05

# Check lemma conditions
cis_lemma <- check_lemma_conditions(cis, alpha = alpha)

# Extract and summarize results
lemma_detailed <- extract_lemma_results(cis_lemma)


lemma_applicable <- lemma_detailed %>%
  filter(applicable == TRUE)

# Plot results
plot_lemma_results(cis_lemma)
```

```{r}
#| label: tbl-lemma-summary

# Create simple summary table
lemma_summary <- lemma_applicable %>%
  summarise(
    "Total 2-Study MAs" = n(),
    "Condition 1 Holds" = sum(condition1_holds, na.rm = TRUE),
    "Condition 2 Holds" = sum(condition2_holds, na.rm = TRUE),
    "Both Conditions" = sum(condition1_holds & condition2_holds, na.rm = TRUE),
    "Neither Condition" = sum(!condition1_holds & !condition2_holds, na.rm = TRUE),
    "Edgington < FE" = sum(edgington_smaller, na.rm = TRUE),
    "Lemma Verified" = sum(lemma_verified == TRUE, na.rm = TRUE),
    "Lemma Failed" = sum(lemma_verified == FALSE, na.rm = TRUE)
  ) %>%
  pivot_longer(
    everything(),
    names_to = "Metric",
    values_to = "Count"
  ) %>%
  mutate(
    Percentage = ifelse(
      Metric == "Total 2-Study MAs",
      NA_real_,
      Count / first(Count) * 100
    )
  )

# Create table
lemma_summary %>%
  gt() %>%
  tab_header(
    title = md("**Lemma 4.3.1 Summary Statistics**"),
  ) %>%
  fmt_number(
    columns = Count,
    decimals = 0,
    use_seps = TRUE
  ) %>%
  fmt_number(
    columns = Percentage,
    decimals = 1,
    suffix = "%"
  ) %>%
  sub_missing(
    columns = Percentage,
    missing_text = "—"
  ) %>%
  cols_align(
    align = "left",
    columns = Metric
  ) %>%
  cols_align(
    align = "center",
    columns = c(Count, Percentage)
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  tab_style(
    style = cell_fill(color = "#f0f0f0"),
    locations = cells_body(rows = Metric == "Total 2-Study MAs")
  ) %>%
  opt_stylize(style = 6, color = "blue")

```

{{< pagebreak >}}

